### 2024-11-12 _IDS702Projectproposal_template.qmd
adding files

---
title: "Vignette Title"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r load-packages, message = FALSE, warning = FALSE}
library(tidyverse)
```
# Dataset 1 (top choice)

**Data source:**

**Brief description:**

**Research question 1:**

-   Outcome variable (include the name/description and type of variable):

**Research question 2:**

-   Outcome variable (include the name/description and type of variable):

**Load the data and provide a `glimpse()`**:

```{r data-1, warning=FALSE}

```

**Exploratory Plots:**

```{r plots-1}

```

# Dataset 2

**Data source:**

**Brief description:**

**Research question 1:**

-   Outcome variable (include the name/description and type of variable):

**Research question 2:**

-   Outcome variable (include the name/description and type of variable):

**Load the data and provide a `glimpse()`**:

```{r data-2, warning=FALSE}

```

**Exploratory Plots:**

```{r plots-2}

```

#

# Dataset 3 (optional)

**Data source:**

**Brief description:**

**Research question 1:**

-   Outcome variable (include the name/description and type of variable):

**Research question 2:**

-   Outcome variable (include the name/description and type of variable):

**Load the data and provide a `glimpse()`**:

```{r data-3, warning=FALSE}

```

**Exploratory Plots:**

```{r plots-3}

```

#

# Team Charter

#### When will you meet as a team to work on the project components? Will these meetings be held in person or virtually?

#### What is your group policy on missing team meetings (e.g., how much advance notice should be provided)?

#### How will your team communicate (email, Slack, text messages)? What is your policy on appropriate response time (within a certain number of hours? Nights/weekends?)?

####

### 2024-11-15 _IDS702Projectproposal_template.qmd

@@
 ---
-title: "Vignette Title"
-author: "Vignette Author"
-date: "`r Sys.Date()`"
-output: rmarkdown::html_vignette
-vignette: >
-  %\VignetteIndexEntry{Vignette Title}
-  %\VignetteEngine{knitr::rmarkdown}
-  %\VignetteEncoding{UTF-8}
+title: "Project Proposal"
+author: "ALL group member names here"
+subtitle: "Due November 17 at 11:59pm"
+format: pdf
 ---
 
-```{r setup, include = FALSE}
-knitr::opts_chunk$set(
-  collapse = TRUE,
-  comment = "#>"
-)
-```
+#### Load Packages
 
@@
 library(tidyverse)
+library(dplyr)
 ```
+
 # Dataset 1 (top choice)
@@
 **Data source:**
+County Health Rankings & Roadmaps (NC 2024 County Data: https://www.countyhealthrankings.org/health-data/north-carolina/data-and-resources)
 
 **Brief description:**
+This dataset contains health factor and health outcome data for each county in NC in 2024.
+It is provided as part of a program by the University of Wisconsin Population Health Institute () and combines data from other sources, such as the Behavioral Risk Factor Surveillance System & National Center for Health Statistics.
 
 **Research question 1:**
+What is the relationship between residential segregation (continuous variable) and juvenile arrests (continuous variable)?
+How is this relationship affected by the county's median household income (categorical variable, where 0=below national average, 1=above national average)?
 
--   Outcome variable (include the name/description and type of variable):
+Residential segregation = Index of dissimilarity where higher values indicate greater residential segregation between Black and white county residents (2018-2022 data).
+
+Juvenile arrests = Rate of delinquency cases per 1000 juveniles (2021 data).
+
+NY = 52
+TX = 254
+NC = 100
+MN = 87
+CO = 64
+WA = 39
+AL = 67
 
@@
 
--   Outcome variable (include the name/description and type of variable):
+-   How does school segregation level (continuous predictor) relate to school funding adequacy (categorical outcome)?
+
+Qs:
+-Is it okay if the variables' data come from different years? 
+state in report and consider it a limitation
+-Is it better to combine multiple years of data for the same set of counties or include 5 different states' counties (but same year)?
+maybe look at year as interaction
+-Is it okay if original dataset only has numerical data and we turn some of the variables into categorical?
 
@@
 ```{r data-1, warning=FALSE}
+texas_full <- read.csv('https://github.com/rrandev03/Practical-Data-Science/raw/refs/heads/main/Texas.csv', skip=1)
+#colnames(texas_full)
+```
 
+```{r}
+texas <- texas_full[c('FIPS', 'State', 'County', "Segregation.Index", "Spending.per.Pupil", "School.Funding.Adequacy", "Segregation.Index.1", "Juvenile.Arrest.Rate")]
+head(texas)
+```
+```{r}
+texas <- texas |> 
+  rename(
+    Residential.Segregation.Index = Segregation.Index.1)
+
+head(texas)
 ```
@@
 #### When will you meet as a team to work on the project components? Will these meetings be held in person or virtually?
+We will meet virtually on Tuesday between 8 and 9 pm.
 
 #### What is your group policy on missing team meetings (e.g., how much advance notice should be provided)?
+If a team member will be missing a meeting, they should provide around 10-12 hours of advance notice (the morning of).
 
 #### How will your team communicate (email, Slack, text messages)? What is your policy on appropriate response time (within a certain number of hours? Nights/weekends?)?
+We will be using WhatsApp primarily for communication, with Slack as our secondary chat.
+We will send Zoom invites via email to reserve our weekly meeting time (Tuesdays 8-9 pm).
+Ideally, team members should respond to or acknowledge messages within 1-3 hours on weekdays and 12 hours on weekends/during the night.
 

### 2024-11-15 CURJ_IDS702Projectproposal.qmd
County data initial load & transform

--- 2024-11-15 _IDS702Projectproposal_template.qmd
@@
 ```{r data-1, warning=FALSE}
-texas_full <- read.csv('https://github.com/rrandev03/Practical-Data-Science/raw/refs/heads/main/Texas.csv', skip=1)
+alabama_full <- read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/county_data/Alabama.csv', skip=1)
+texas_full <- read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/county_data/Texas.csv', skip=1)
+nc_full <- read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/county_data/North%20Carolina.csv', skip=1)
+colorado_full <- read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/county_data/Colorado.csv', skip=1)
+washington_full <- read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/county_data/Washington.csv', skip=1)
+ny_full <- read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/county_data/New%20York.csv', skip=1)
+minnesota_full <-  read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/county_data/Minnesota.csv', skip=1)
 #colnames(texas_full)
@@
 ```{r}
-texas <- texas_full[c('FIPS', 'State', 'County', "Segregation.Index", "Spending.per.Pupil", "School.Funding.Adequacy", "Segregation.Index.1", "Juvenile.Arrest.Rate")]
-head(texas)
+head(alabama_full)
 ```
+
+```{r}
+desired_cols <- c('FIPS', 'State', 'County', "Segregation.Index", "Spending.per.Pupil", "School.Funding.Adequacy", "Segregation.Index.1", "Juvenile.Arrest.Rate")
+
+alabama <- alabama_full[desired_cols]
+texas <- texas_full[desired_cols]
+nc <- nc_full[desired_cols]
+ny <- ny_full[desired_cols]
+washington <- washington_full[desired_cols]
+colorado <- colorado_full[desired_cols]
+minnesota <- minnesota_full[desired_cols]
+
+```
+
 ```{r}
@@
   rename(
+    School.Segregation.Index = Segregation.Index,
     Residential.Segregation.Index = Segregation.Index.1)
 
-head(texas)
+alabama <- alabama |> 
+  rename(
+    School.Segregation.Index = Segregation.Index,
+    Residential.Segregation.Index = Segregation.Index.1)
+
+nc <- nc |> 
+  rename(
+    School.Segregation.Index = Segregation.Index,
+    Residential.Segregation.Index = Segregation.Index.1)
+
+ny <- ny |> 
+  rename(
+    School.Segregation.Index = Segregation.Index,
+    Residential.Segregation.Index = Segregation.Index.1)
+
+washington <- washington |> 
+  rename(
+    School.Segregation.Index = Segregation.Index,
+    Residential.Segregation.Index = Segregation.Index.1)
+
+colorado <- colorado |> 
+  rename(
+    School.Segregation.Index = Segregation.Index,
+    Residential.Segregation.Index = Segregation.Index.1)
+
+minnesota <- minnesota |> 
+  rename(
+    School.Segregation.Index = Segregation.Index,
+    Residential.Segregation.Index = Segregation.Index.1)
+
+head(minnesota)
+```
+
+```{r}
+all_states <- rbind(alabama, texas, nc, ny, washington, colorado, minnesota)
+all_states[sample(nrow(all_states), 10), ]
+```
+
+```{r}
+all_states_final <- all_states |>
+  mutate(Region = factor(case_when(
+    State == 'Texas' ~ 'West South Central', # use NA to set a value to missing
+    State == 'Alabama' ~ 'East South Central',
+    State == 'North Carolina' ~ 'South Atlantic',
+    State == 'New York' ~ 'Middle Atlantic',
+    State == 'Washington' ~ 'Pacific',
+    State == 'Colorado' ~ 'Mountain',
+    State == 'Minnesota' ~ 'West North Central',
+    TRUE ~ State #this says to keep all other values the same
+  )), School.Funding.Cat = factor(case_when(
+    School.Funding.Adequacy < 0 ~ 'Inadequate', # use NA to set a value to missing
+    School.Funding.Adequacy > 0  ~ 'Adequate',
+  TRUE ~ NA
+  )))
+
+all_states_final[sample(nrow(all_states_final), 10), c('State', 'County', 'School.Funding.Adequacy', 'Region', 'School.Funding.Cat')]
+```
+```{r}
+
 ```

### 2024-11-15 CURJ_IDS702Projectproposal.qmd
Fixing research question 1 for county dataset
EDA for Q2 county data

@@
 format: pdf
+editor: 
+  markdown: 
+    wrap: 72
 ---
@@
 library(dplyr)
+library(ggplot2)
 ```
@@
 
-**Data source:**
-County Health Rankings & Roadmaps (NC 2024 County Data: https://www.countyhealthrankings.org/health-data/north-carolina/data-and-resources)
+**Data source:** County Health Rankings & Roadmaps (NC 2024 County Data:
+https://www.countyhealthrankings.org/health-data/north-carolina/data-and-resources)
 
-**Brief description:**
-This dataset contains health factor and health outcome data for each county in NC in 2024.
-It is provided as part of a program by the University of Wisconsin Population Health Institute () and combines data from other sources, such as the Behavioral Risk Factor Surveillance System & National Center for Health Statistics.
+**Brief description:** This dataset contains health factor and health
+outcome data for each county in NC in 2024. It is provided as part of a
+program by the University of Wisconsin Population Health Institute ()
+and combines data from other sources, such as the Behavioral Risk Factor
+Surveillance System & National Center for Health Statistics.
 
-**Research question 1:**
-What is the relationship between residential segregation (continuous variable) and juvenile arrests (continuous variable)?
-How is this relationship affected by the county's median household income (categorical variable, where 0=below national average, 1=above national average)?
+**Research question 1:** What is the relationship between residential
+segregation (continuous variable) and juvenile arrests (continuous
+variable)? How is this relationship affected by the region of the US
+that the county / state is in (categorical variable called Region)?
 
-Residential segregation = Index of dissimilarity where higher values indicate greater residential segregation between Black and white county residents (2018-2022 data).
+Residential segregation = Index of dissimilarity where higher values
+indicate greater residential segregation between Black and white county
+residents (2018-2022 data).
 
-Juvenile arrests = Rate of delinquency cases per 1000 juveniles (2021 data).
+Juvenile arrests = Rate of delinquency cases per 1000 juveniles (2021
+data).
 
-NY = 52
-TX = 254
-NC = 100
-MN = 87
-CO = 64
-WA = 39
-AL = 67
+NY = 52 TX = 254 NC = 100 MN = 87 CO = 64 WA = 39 AL = 67
 
@@
 
--   How does school segregation level (continuous predictor) relate to school funding adequacy (categorical outcome)?
+-   How does school segregation level (continuous predictor) relate to
+    school funding adequacy (categorical outcome)?
 
-Qs:
--Is it okay if the variables' data come from different years? 
-state in report and consider it a limitation
--Is it better to combine multiple years of data for the same set of counties or include 5 different states' counties (but same year)?
-maybe look at year as interaction
--Is it okay if original dataset only has numerical data and we turn some of the variables into categorical?
+Qs: -Is it okay if the variables' data come from different years? state
+in report and consider it a limitation -Is it better to combine multiple
+years of data for the same set of counties or include 5 different
+states' counties (but same year)? maybe look at year as interaction -Is
+it okay if original dataset only has numerical data and we turn some of
+the variables into categorical?
 
@@
 ```{r}
-head(alabama_full)
-```
-
-```{r}
-desired_cols <- c('FIPS', 'State', 'County', "Segregation.Index", "Spending.per.Pupil", "School.Funding.Adequacy", "Segregation.Index.1", "Juvenile.Arrest.Rate")
+desired_cols <- c('FIPS', 'State', 'County', 
+"Segregation.Index",
+"Spending.per.Pupil", "School.Funding.Adequacy", 
+"Segregation.Index.1", "Juvenile.Arrest.Rate")
 
@@
 
-head(minnesota)
+#head(minnesota)
 ```
@@
 ```{r}
-all_states <- rbind(alabama, texas, nc, ny, washington, colorado, minnesota)
+all_states <- rbind(alabama, texas, nc, ny, washington, 
+                    colorado, minnesota)
 all_states[sample(nrow(all_states), 10), ]
@@
   mutate(Region = factor(case_when(
-    State == 'Texas' ~ 'West South Central', # use NA to set a value to missing
+    State == 'Texas' ~ 'West South Central', 
     State == 'Alabama' ~ 'East South Central',
@@
   )), School.Funding.Cat = factor(case_when(
-    School.Funding.Adequacy < 0 ~ 'Inadequate', # use NA to set a value to missing
+    School.Funding.Adequacy < 0 ~ 'Inadequate', 
     School.Funding.Adequacy > 0  ~ 'Adequate',
   TRUE ~ NA
-  )))
+  ), levels=c('Inadequate', 'Adequate'), labels=c(0, 1)))
 
-all_states_final[sample(nrow(all_states_final), 10), c('State', 'County', 'School.Funding.Adequacy', 'Region', 'School.Funding.Cat')]
+all_states_final[sample(nrow(all_states_final), 10), 
+c('State', 'County', 'School.Funding.Adequacy', 'Region', 
+  'School.Funding.Cat')]
 ```
-```{r}
+
+**Exploratory Plots:** 
+Q1 Outcome Variable (Juvenile Arrests):
+```{r plots-1}
 
@@
 
-**Exploratory Plots:**
+Q1 Relationship of Interest (Residential Segregation vs. Juvenile Arrests):
+```{r plots-2}
 
-```{r plots-1}
+```
 
+Q2 Outcome Variable (School Funding Adequacy):
+
+```{r}
+ggplot(all_states_final, aes(x = State, fill=School.Funding.Cat)) + 
+  geom_bar() + labs(x="State", y="Count of Counties", 
+  fill='School Funding Adequacy', title = "School Funding Adequacy 
+  in Seven US States") + theme(axis.text.x = 
+  element_text(angle = 45, vjust = 1, hjust=1)) +
+  scale_fill_discrete(labels=c('Inadequate Funding', 
+  'Adequate Funding', 'No Data Available'))
+```
+
+Q2 Relationship of Interest (School Segregation & School Funding
+Adequacy):
+
+```{r plots-4}
+ggplot(all_states_final, aes(x= School.Segregation.Index, y= School.Funding.Cat))+
+  geom_jitter()+ 
+  labs(title = "School Segregation vs. School Funding Adequacy 
+       in Counties Across 7 US States",
+       x = "School Segregation Index",
+       y = "School Funding Adequacy") 
 ```
@@
 
--   Outcome variable (include the name/description and type of variable):
+-   Outcome variable (include the name/description and type of
+    variable):
 
@@
 
-```{r plots-2}
+```{r plots-5}
 
@@
 #### When will you meet as a team to work on the project components? Will these meetings be held in person or virtually?
+
 We will meet virtually on Tuesday between 8 and 9 pm.
@@
 #### What is your group policy on missing team meetings (e.g., how much advance notice should be provided)?
-If a team member will be missing a meeting, they should provide around 10-12 hours of advance notice (the morning of).
+
+If a team member will be missing a meeting, they should provide around
+10-12 hours of advance notice (the morning of).
 
 #### How will your team communicate (email, Slack, text messages)? What is your policy on appropriate response time (within a certain number of hours? Nights/weekends?)?
-We will be using WhatsApp primarily for communication, with Slack as our secondary chat.
-We will send Zoom invites via email to reserve our weekly meeting time (Tuesdays 8-9 pm).
-Ideally, team members should respond to or acknowledge messages within 1-3 hours on weekdays and 12 hours on weekends/during the night.
+
+We will be using WhatsApp primarily for communication, with Slack as our
+secondary chat. We will send Zoom invites via email to reserve our
+weekly meeting time (Tuesdays 8-9 pm). Ideally, team members should
+respond to or acknowledge messages within 1-3 hours on weekdays and 12
+hours on weekends/during the night.
 

### 2024-11-17 CURJ_IDS702Projectproposal.qmd
finished cleaning merged data

--- 2024-11-15 CURJ_IDS702Projectproposal.qmd
@@
 What is the relationship between residential segregation (continuous variable) and juvenile arrests (continuous variable)?
-How is this relationship affected by the county's median household income (categorical variable, where 0=below national average, 1=above national average)?
+How is this relationship affected by the region of the US that the county / state is in (categorical variable called Region)?
 
@@
 ```{r}
-
+acs_proptax <- read.csv('https://raw.githubusercontent.com/wujenny214/F24_STATS_FINAL/refs/heads/main/county_data/acs_proptax.csv')
+merged_data_full <- merge(acs_proptax, all_states_final, by = "FIPS")
+merged_data_full <- select(merged_data_full, -c(6, 7))
 ```

### 2024-11-17 CURJ_IDS702Projectproposal.qmd
Completed EDA for county

--- 2024-11-15 CURJ_IDS702Projectproposal.qmd
@@
 library(dplyr)
+library(ggplot2)
+library(GGally)
+library(patchwork)
 ```
@@
 
-**Data source:**
-County Health Rankings & Roadmaps (NC 2024 County Data: https://www.countyhealthrankings.org/health-data/north-carolina/data-and-resources)
+**Data source:** County Health Rankings & Roadmaps
+https://www.countyhealthrankings.org/health-data/north-carolina/data-and-resources)
 
-**Brief description:**
-This dataset contains health factor and health outcome data for each county in NC in 2024.
-It is provided as part of a program by the University of Wisconsin Population Health Institute () and combines data from other sources, such as the Behavioral Risk Factor Surveillance System & National Center for Health Statistics.
+Secondary dataset: https://data.census.gov/table/ACSDT1Y2018.B25103?q=property
 
-**Research question 1:**
-What is the relationship between residential segregation (continuous variable) and juvenile arrests (continuous variable)?
-How is this relationship affected by the county's median household income (categorical variable, where 0=below national average, 1=above national average)?
+**Brief description:** The source of the datasets includes health
+factor and health outcome data for each county in the US by state in
+2024. These datasets are provided as part of a program by the University
+of Wisconsin Population Health Institute and combines data from other
+sources, such as the Behavioral Risk Factor Surveillance System &
+National Center for Health Statistics. We are compiling the datasets for
+seven states in seven different regions of the US to create one dataset
+(North Carolina, Texas, Alabama, Colorado, Minnesota, Washington, & New
+York). We will also be merging in a second American Community Survey dataset to obtain property tax data for our second research question.
 
-Residential segregation = Index of dissimilarity where higher values indicate greater residential segregation between Black and white county residents (2018-2022 data).
+**Research question 1:** What is the relationship between residential
+segregation (continuous variable), percentage of the county population
+with limited access to healthy foods (continuous variable), and
+percentage of adults who are uninsured (continuous variable), and
+average life expectancy (continuous variable)? How is the relationship
+between residential segregation and life expectancy affected by the
+region of the US that the county / state is in (categorical variable
+that we will create called Region)? The outcome variable is average life
+expectancy. The predictors are residential segregation, percentage with
+limited access to healthy foods, and percentage of adults who are
+uninsured.
 
-Juvenile arrests = Rate of delinquency cases per 1000 juveniles (2021 data).
+Average life expectancy = average number of years a person is expected
+to live (data from National Center for Health Statistics - Natality and
+Mortality Files; Census Population Estimates Program, 2019-2021).
 
-NY = 52
-TX = 254
-NC = 100
-MN = 87
-CO = 64
-WA = 39
-AL = 67
+Residential segregation = Index of dissimilarity where higher values
+indicate greater residential segregation between Black and white county
+residents (data from American Community Survey 5-year estimate,
+2018-2022).
 
-**Research question 2:**
+Percentage limited access to healthy foods = Percentage of population
+who are low-income and do not live close to a grocery store (data from
+USDA Food Environment Atlas, 2019).
 
--   How does school segregation level (continuous predictor) relate to school funding adequacy (categorical outcome)?
+Percentage adults uninsured = Percentage of adults under age 65 without
+health insurance (data from Small Area Health Insurance Estimates,
+2021).
 
-Qs:
--Is it okay if the variables' data come from different years? 
-state in report and consider it a limitation
--Is it better to combine multiple years of data for the same set of counties or include 5 different states' counties (but same year)?
-maybe look at year as interaction
--Is it okay if original dataset only has numerical data and we turn some of the variables into categorical?
+**Research question 2:** What is the relationship between residential
+segregation (continuous variable), property tax (continuous variable),
+and median household income (continuous variable), on school funding
+adequacy (categorical variable that we will create)?
 
-**Load the data and provide a `glimpse()`**:
+The outcome variable is school funding adequacy, with 0=inadequate
+funding and 1=adequate funding. The predictors are residential
+segregation, property tax, and median household income.
+
+School funding adequacy = The average gap in dollars between actual and
+required spending per pupil among public school districts. Required
+spending is an estimate of dollars needed to achieve U.S. average test
+scores in each district. (data from School Finance Indicators Database,
+2021).
+
+Median property tax = Median property tax value for each county between
+2018-2022 (data from IPUMS).
+
+Median household income = The income where half of households in a
+county earn more and half of households earn less (data from Small Area
+Income and Poverty Estimates; American Community Survey, 5-year
+estimates, 2022 & 2018-2022).
+
+**Load the datasets and prepare final dataset`**:
 
@@
 minnesota_full <-  read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/county_data/Minnesota.csv', skip=1)
-#colnames(texas_full)
+#colnames(alabama_full)
 ```
@@
 ```{r}
-head(alabama_full)
-```
-
-```{r}
-desired_cols <- c('FIPS', 'State', 'County', "Segregation.Index", "Spending.per.Pupil", "School.Funding.Adequacy", "Segregation.Index.1", "Juvenile.Arrest.Rate")
+desired_cols <- c('FIPS', 'State', 'County', 
+"Life.Expectancy", 
+"Segregation.Index.1", "X..Limited.Access.to.Healthy.Foods.1", 
+"X..Uninsured.Adults.1", 
+"Spending.per.Pupil", "School.Funding.Adequacy", 
+"Median.Household.Income" 
+)
 
@@
 minnesota <- minnesota_full[desired_cols]
-
+#head(alabama)
 ```
@@
 ```{r}
+
 texas <- texas |> 
   rename(
-    School.Segregation.Index = Segregation.Index,
+    Percent.Uninsured.Adults = X..Uninsured.Adults.1, 
+    Percent.Limited.Access.Healthy.Foods = X..Limited.Access.to.Healthy.Foods.1,
     Residential.Segregation.Index = Segregation.Index.1)
@@
   rename(
-    School.Segregation.Index = Segregation.Index,
+    Percent.Uninsured.Adults = X..Uninsured.Adults.1, 
+    Percent.Limited.Access.Healthy.Foods = X..Limited.Access.to.Healthy.Foods.1,
     Residential.Segregation.Index = Segregation.Index.1)
@@
   rename(
-    School.Segregation.Index = Segregation.Index,
+    Percent.Uninsured.Adults = X..Uninsured.Adults.1, 
+    Percent.Limited.Access.Healthy.Foods = X..Limited.Access.to.Healthy.Foods.1,
     Residential.Segregation.Index = Segregation.Index.1)
 
-head(minnesota)
+#head(minnesota)
 ```
@@
 ```{r}
-all_states <- rbind(alabama, texas, nc, ny, washington, colorado, minnesota)
-all_states[sample(nrow(all_states), 10), ]
+#Remove the first row from each dataset before concatenating 
+#because it contains statewide data instead of a specific county
+all_states <- rbind(alabama[-1,], texas[-1,], nc[-1,], ny[-1,], 
+                    washington[-1,], colorado[-1,], minnesota[-1,]) 
+#all_states[sample(nrow(all_states), 10), ]
 ```
@@
 all_states_final <- all_states |>
-  mutate(Region = factor(case_when(
-    State == 'Texas' ~ 'West South Central', # use NA to set a value to missing
+  mutate(
+    Region = factor(case_when(
+    State == 'Texas' ~ 'West South Central', 
     State == 'Alabama' ~ 'East South Central',
@@
   )), School.Funding.Cat = factor(case_when(
-    School.Funding.Adequacy < 0 ~ 'Inadequate', # use NA to set a value to missing
+    School.Funding.Adequacy < 0 ~ 'Inadequate', 
     School.Funding.Adequacy > 0  ~ 'Adequate',
   TRUE ~ NA
-  )))
-
-all_states_final[sample(nrow(all_states_final), 10), c('State', 'County', 'School.Funding.Adequacy', 'Region', 'School.Funding.Cat')]
-```
-```{r}
+  ), levels=c('Inadequate', 'Adequate'), labels=c(0, 1)))
 
@@
 
-**Exploratory Plots:**
+```{r}
+prop_tax <- read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/county_data/acs_proptax.csv')
+```
 
-```{r plots-1}
+```{r}
+merged_df <- merge(all_states_final, prop_tax, by = "FIPS", 
+              all.x = TRUE, all.y = TRUE)
+
+#merged_df[sample(nrow(all_states_final), 10), ]
+#colSums(is.na(final_df))
 
 ```
+
+```{r}
+final_df <- merged_df |> 
+  mutate(
+    Median.Prop.Tax = as.numeric(med_re_.taxes_tot))
+final_df = subset(final_df, select = -c(med_re_.taxes_tot, geo_id, county_name, state_name) )
+
+```
+
+**Provide a `glimpse():**
+
+```{r}
+glimpse(final_df)
+```
+
+**Exploratory Plots:** Q1 Outcome Variable (Life Expectancy):
+
+```{r plots-1, warning=FALSE}
+ggplot(final_df, aes(x = Life.Expectancy)) +
+  geom_histogram(binwidth = 0.5, fill = "skyblue") +
+  facet_wrap(~ Region) +
+  labs(title = "Life Expectancy Distribution", x = "Life Expectancy", y = "Count")
+```
+
+Q1 Relationship of Interest:
+
+```{r plots-2, warning=FALSE}
+plot1 <- ggplot(final_df, aes(x = Residential.Segregation.Index, y = Life.Expectancy)) +
+  geom_point(color='purple') +
+  labs(x = "Residential Segregation", y = "Life Expectancy")+ 
+  theme(plot.margin = margin(t = 20, r = 30, b = 20, l = 10),
+axis.title.x = element_text(size = 8),
+axis.title.y = element_text(size = 8))
+
+plot2 <- ggplot(final_df, aes(x = Percent.Limited.Access.Healthy.Foods
+, y = Life.Expectancy)) +
+  geom_point(color = "green") +
+  labs( x = "% Population With Limited Healthy Food Access") +
+  theme(plot.margin = margin(t = 20, r = 20, b = 20, l = 20),
+        axis.title.x = element_text(size = 8),
+        axis.title.y = element_text(size = 8))
+
+plot3 <- ggplot(final_df, aes(x = Percent.Uninsured.Adults, y = Life.Expectancy)) +
+  geom_point(color = "blue") +
+  labs(x = "% Adult Population Without Health Insurance", y = "Life Expectancy")+ 
+  theme(plot.margin = margin(t = 20, r = 20, b = 20, l = 20),
+        axis.title.x = element_text(size = 8),
+        axis.title.y = element_text(size = 8))
+
+plot1 + plot2 + plot3 + plot_layout(ncol = 2, nrow=2) + 
+  plot_annotation(title = "Relationships Between Predictors and 
+                  Life Expectancy Outcome",
+                  theme = theme(
+      plot.title = element_text(hjust=0.5)))
+```
+
+Q2 Outcome Variable (School Funding Adequacy):
+
+```{r}
+ggplot(final_df, aes(x = State, fill=School.Funding.Cat)) + 
+  geom_bar() + labs(x="State", y="Count of Counties", 
+  fill='School Funding Adequacy', title = "School Funding Adequacy 
+  in Seven US States") + theme(axis.text.x = 
+  element_text(angle = 45, vjust = 1, hjust=1)) +
+  scale_fill_discrete(labels=c('Inadequate Funding', 
+  'Adequate Funding', 'No Data Available'))
+```
+
+Q2 Relationship of Interest:
+
+```{r plots-4, warning=FALSE}
+ggpairs(final_df, columns = c("Residential.Segregation.Index", 
+                              "Median.Household.Income",
+                              "Median.Prop.Tax"), 
+        aes(color = School.Funding.Cat)) + 
+        labs(title = "Pairwise Plots of Predictors Colored by 
+             School Funding Adequacy",
+              x = "Predictor 1",
+              y = "Predictor 2",
+              color = "School Funding Adequacy") +
+         theme(plot.title = element_text(size=15, hjust=0.5), 
+               strip.text.x = element_text(size = 5),
+               strip.text.y = element_text(size = 4))
+
+```
+
+The above figure is a pairwise scatterplot for each pair of predictor
+variables (residential segregation, property tax, and household income),
+with the points being colored by the school funding adequacy outcome.
+Along the diagonal are histograms showing the distribution of each
+individual predictor, separated by each level of the school funding
+adequacy outcome variable (red = 0, or inadequate funding; blue = 1, or
+adequate funding; gray = NA values). The upper half shows the
+correlation coefficient between each pair of predictors, and also the
+specific correlation coefficients when the data is separated out by each
+level of the outcome.
 

### 2024-11-17 CURJ_IDS702Projectproposal.qmd

@@
 ```{r}
+
 merged_df <- merge(all_states_final, prop_tax, by = "FIPS", 
@@
 #colSums(is.na(final_df))
-
 ```

### 2024-11-17 NFL_IDS702Projectproposal.qmd

---
title: "Project Proposal"
author: "ALL group member names here"
subtitle: "Due November 17 at 11:59pm"
format: pdf
editor: 
  markdown: 
    wrap: 72
---

#### Load Packages

```{r load-packages, message = FALSE, warning = FALSE}
library(dplyr)
```

# NFL Dataset

**Data source:** NFL Kaggle 2025 Play Dataset
https://www.kaggle.com/competitions/nfl-big-data-bowl-2025/data?select=player_play.csv

**Brief description:** The Play Data Dataset provides detailed
play-by-play information about football games, capturing various aspects
of gameplay, team dynamics, and player actions. This dataset includes
variables that tell us more about what influences game play within a
football game.

**Research question 1:** Question: Does the distance that a quarterback
drops back after the snap affect the probability of a pass completion?
By focusing on quarterback actions and pass outcomes, it could help
understand if dropping back type was effective or not.

Outcome Variable: Pass Result (passResult) - This is a binary variable,
with categories such as "Complete" (C) and "Incomplete" (I), indicating
whether the pass was successfully completed.

Independent Variables: **dropbackType**: The type of drop back after the
snap by the QB (Traditional, Designed Rollout, Scramble, Scramble
Rollout, Designed Rollout Left, Designed Rollout Right, Scramble Rollout
Left, Scramble Rollout Right, Designed Run, QB Draw, Rollout, text)
(categorical) \
**dropbackDistance**: The distance the QB dropped back (yards) behind
the center after the snap (numeric) \
**timeToSack**: Time before the QB is sacked; this indicates the level
of pass protection (numeric) **rushLocationType**: Indicates where
defensive pressure might be coming from; outside pressure can force
quicker throws. (categorical)

Possible other variables: \
**pff_passCoverage**: Defensive coverage scheme (e.g., man vs. zone);
some schemes are harder to throw against.(categorica) \
**pff_manZone**: Specific indicator of man or zone coverage.
(categorical) \
**unblockedPressure**: Whether there was pressure from an unblocked
defender; unblocked pressure typically lowers completion probability.
(binary) \
**passTippedAtLine**: Whether the pass was tipped; this directly impacts
completion probability.(binary)

**Research question 2:** Question: How does the game situation affect
win probability change?

Outcome Variable: The outcome variable would be the
homeTeamWinProbabilityAdded

Independent Variable: **quarter**: the quarter the play is in may impact
the delta of the teams probability of winning as any slip up in the
latter half of the game may be more determental and harder to come back
from. \
**down**: how many downs are left in the play impacts when the team has
the ball\
**yardsToGo:** how many yards they have left to increase scoring
opportunities \
**gameClock:** how much time they have left to increase scoring
opportunities \
**preSnapHomeScore**: how far behind the are may impact their likelihood
of coming back or their hold on the game\
**preSnapVisitorScore**: how far ahead the opponent may be may impact
their likelihood of the team's coming back or hold on the game \
\
We can add an interaction term for quarter and down as that will test
whether the effect of down and distance on win probability change is
stronger or weaker in certain quarters.

**Load the data and provide a `glimpse()`**:

```{r data-1, warning=FALSE}
nfldata <- read.csv("https://raw.githubusercontent.com/wujenny214/F24_STATS_FINAL/refs/heads/main/data/NFL/plays.csv") #replace with you file path name 
glimpse(nfldata) 
```

**Exploratory Plots:** \
\
RQ1: Does the distance that a quarterback drops back after the snap
affect the probability of a pass completion?

```{r}
#Build a histogram of the win probably delta, "homeTeamWinProbabilityAdded"

ggplot(nfldata, aes(x = as.factor(quarter), y = preSnapHomeTeamWinProbability)) +
  geom_line() +
  labs(title = "Win Probability Before Play by Quarter (Home Team)", x = "Quarter", y = "Home Team Win Probability")
```

RQ2: How does the game situation affect win probability change?

```{r}
#Create a plot that demonstrates the relationship between the game's characteristics and the probability of winning. maybe a graph on gameClock and homeTeamWinProbabilityAdded

```

### 2024-11-17 CURJ_IDS702Projectproposal.qmd
commit change to branch

@@
 
-Secondary dataset: https://data.census.gov/table/ACSDT1Y2018.B25103?q=property
+Secondary dataset:
+https://data.census.gov/table/ACSDT1Y2018.B25103?q=property
 
-**Brief description:** The source of the datasets includes health
-factor and health outcome data for each county in the US by state in
-2024. These datasets are provided as part of a program by the University
-of Wisconsin Population Health Institute and combines data from other
-sources, such as the Behavioral Risk Factor Surveillance System &
-National Center for Health Statistics. We are compiling the datasets for
-seven states in seven different regions of the US to create one dataset
-(North Carolina, Texas, Alabama, Colorado, Minnesota, Washington, & New
-York). We will also be merging in a second American Community Survey dataset to obtain property tax data for our second research question.
+**Brief description:** This dataset contains health factor and health
+outcome data for each county in NC in 2024. It is provided as part of a
+program by the University of Wisconsin Population Health Institute ()
+and combines data from other sources, such as the Behavioral Risk Factor
+Surveillance System & National Center for Health Statistics.
 
 **Research question 1:** What is the relationship between residential
-segregation (continuous variable), percentage of the county population
-with limited access to healthy foods (continuous variable), and
-percentage of adults who are uninsured (continuous variable), and
-average life expectancy (continuous variable)? How is the relationship
-between residential segregation and life expectancy affected by the
-region of the US that the county / state is in (categorical variable
-that we will create called Region)? The outcome variable is average life
-expectancy. The predictors are residential segregation, percentage with
-limited access to healthy foods, and percentage of adults who are
-uninsured.
+segregation (continuous variable) and juvenile arrests (continuous
+variable)? How is this relationship affected by the region of the US
+that the county / state is in (categorical variable called Region)?
 
-Average life expectancy = average number of years a person is expected
-to live (data from National Center for Health Statistics - Natality and
-Mortality Files; Census Population Estimates Program, 2019-2021).
+-   **Outcome (Dependent) Variable** $Juvenile.Arrest.Rate$
 
-Residential segregation = Index of dissimilarity where higher values
-indicate greater residential segregation between Black and white county
-residents (data from American Community Survey 5-year estimate,
-2018-2022).
+    -   Represents the rate of delinquency cases per 1,000 juveniles in
+        a county. This variable quantifies juvenile arrests as a measure
+        of youth criminal activity.
+    -   **Source**: Data from Juvenile Justice Databases, 2021.
 
-Percentage limited access to healthy foods = Percentage of population
-who are low-income and do not live close to a grocery store (data from
-USDA Food Environment Atlas, 2019).
+-   **Primary Independent Variable** $Residential.Segregation.Index$
 
-Percentage adults uninsured = Percentage of adults under age 65 without
-health insurance (data from Small Area Health Insurance Estimates,
-2021).
+    -   The index of dissimilarity measures residential segregation
+        between Black and white residents in a county. Higher values
+        indicate greater segregation. This metric captures the extent to
+        which these populations reside in separate geographic areas
+        within the same county.
+    -   **Source:** American Community Survey 5-year estimates,
+        2018-2022.
+
+-   **Interaction Term** $Residential.Segregation.Index * Region$
+
+    -   Examines how the relationship between residential segregation
+        and juvenile arrests varies by region (`Region` categorical
+        variable).
+
+-   **Control Variables** Adjusts for other factors that might confound
+    the relationship between segregation and arrests.
+
+    -   $Spending.per.Pupil$ –The average educational expenditure per
+        student in a district, which reflects investment in education
+        and resources that may impact juvenile arrest rates.
+
+        -   **Source:** State or National Education Finance Databases,
+            2021
+
+    -   $School.Funding.Adequacy$ –A continuous measure indicating
+        whether school funding meets the needs of the district.
+        Inadequate funding may correlate with higher juvenile
+        delinquency rates.
+
+        -   **Source:** School Finance Indicators Database, 2021
+
+    -   $School.Segregation.Index$ –A measure of racial or ethnic
+        segregation within schools, potentially influencing both
+        educational outcomes and juvenile arrest rates.
+
+        -   **Source**: Education Statistics Databases, 2021
+
+    -   $Median.Prop.Tax$ –Median residential property taxes, serving as
+        a proxy for socioeconomic conditions and wealth distribution
+        within a county.
+
+        -   **Source**: IPUMS, 2018-2022.
 
@@
 
-The outcome variable is school funding adequacy, with 0=inadequate
-funding and 1=adequate funding. The predictors are residential
-segregation, property tax, and median household income.
+-   **Outcome (Dependent) Variable** $School.Funding.Cat$
 
-School funding adequacy = The average gap in dollars between actual and
-required spending per pupil among public school districts. Required
-spending is an estimate of dollars needed to achieve U.S. average test
-scores in each district. (data from School Finance Indicators Database,
-2021).
+    -   Categorical variable indicating school funding adequacy. Values:
+        "Adequate" or "Inadequate" (derived from
+        `School.Funding.Adequacy`).
+    -   **Source:** Derived from the School Finance Indicators
+        Database, 2021. The adequacy metric reflects the average dollar
+        gap between actual and required per-pupil spending needed to
+        achieve U.S. average test scores.
 
-Median property tax = Median property tax value for each county between
-2018-2022 (data from IPUMS).
+-   **Primary Independent Variable:**
 
-Median household income = The income where half of households in a
-county earn more and half of households earn less (data from Small Area
-Income and Poverty Estimates; American Community Survey, 5-year
-estimates, 2022 & 2018-2022).
+    -   $Residential.Segregation.Index$: A continuous variable measuring
+        the level of residential segregation between Black and white
+        residents in a county. Higher values indicate greater
+        segregation.
 
-**Load the datasets and prepare final dataset`**:
+        -   **Source**: American Community Survey 5-year estimates,
+            2018-2022.
+
+    -   $Median.Prop.Tax$ :A continuous variable representing the median
+        property tax value in each county. It serves as a proxy for
+        socioeconomic resources, as higher property taxes often
+        correlate with better-funded schools.
+
+        -   **Source**: Data from IPUMS, 2018-2022.
+
+    -   $Median.Household.Income$ : Median income for households in the
+        county (proxy needed if missing; likely not in your dataset).
+
+        -   **Source**: Small Area Income and Poverty Estimates;
+            American Community Survey, 5-year estimates, 2018-2022.
+
+-   **Interaction Term** $Residential.Segregation.Index * Region$
+
+    -   Examines how residential segregation’s relationship with school
+        funding adequacy varies across regions.
+    -   Allows the model to capture region-specific dynamics in the
+        influence of segregation on school funding adequacy.
+
+-   **Control Variables**
+
+    -   $Spending.per.Pupil$ –Average spending per student in a
+        district.
+
+    -   $School.Segregation.Index$ –A measure of racial/ethnic
+        segregation within schools.
+
+**Load the datasets and prepare final dataset\`**:
 
@@
 
-**Provide a `glimpse():**
+**Provide a \`glimpse():**
 

### 2024-11-17 NFL_IDS702Projectproposal.qmd
NFL data set question 2 EDA

@@
 Left, Scramble Rollout Right, Designed Run, QB Draw, Rollout, text)
-(categorical) \
+(categorical)\
 **dropbackDistance**: The distance the QB dropped back (yards) behind
-the center after the snap (numeric) \
+the center after the snap (numeric)\
 **timeToSack**: Time before the QB is sacked; this indicates the level
@@
 
-Possible other variables: \
+Possible other variables:\
 **pff_passCoverage**: Defensive coverage scheme (e.g., man vs. zone);
-some schemes are harder to throw against.(categorica) \
+some schemes are harder to throw against.(categorica)\
 **pff_manZone**: Specific indicator of man or zone coverage.
-(categorical) \
+(categorical)\
 **unblockedPressure**: Whether there was pressure from an unblocked
 defender; unblocked pressure typically lowers completion probability.
-(binary) \
+(binary)\
 **passTippedAtLine**: Whether the pass was tipped; this directly impacts
@@
 latter half of the game may be more determental and harder to come back
-from. \
+from.\
 **down**: how many downs are left in the play impacts when the team has
@@
 **yardsToGo:** how many yards they have left to increase scoring
-opportunities \
+opportunities\
 **gameClock:** how much time they have left to increase scoring
-opportunities \
+opportunities\
 **preSnapHomeScore**: how far behind the are may impact their likelihood
@@
 **preSnapVisitorScore**: how far ahead the opponent may be may impact
-their likelihood of the team's coming back or hold on the game \
+their likelihood of the team's coming back or hold on the game\
 \
@@
 
-**Exploratory Plots:** \
+**Exploratory Plots:**\
 \
@@
 #Build a histogram of the win probably delta, "homeTeamWinProbabilityAdded" 
+library(ggplot2)
 
 ggplot(nfldata, aes(x = as.factor(quarter), y = preSnapHomeTeamWinProbability)) +
-  geom_line() +
+  geom_histogram() +
   labs(title = "Win Probability Before Play by Quarter (Home Team)", x = "Quarter", y = "Home Team Win Probability")
@@
 #Create a plot that demonstrates the relationship between the game's characteristics and the probability of winning. maybe a graph on gameClock and homeTeamWinProbabilityAdded
+ggplot(nfldata, aes(x = homeTeamWinProbabilityAdded)) +
+  geom_histogram(binwidth = 0.05, fill = "skyblue") +  # Adjust binwidth as needed
+  labs(title = "Home Team Win Probability Added Distribution",
+       x = "Home Team Win Probability Added",
+       y = "Count") +
+  theme_minimal()  # Minimal theme for a clean look
+```
+
+```{r}
+#Create a plot that demonstrates the relationship between the game's characteristics and the probability of winning. maybe a graph on gameClock and homeTeamWinProbabilityAdded
+# Load the necessary packages
+library(ggplot2)
+library(patchwork)
+
+# Create the individual plots
+plot1 <- ggplot(nfldata, aes(x = factor(down), y = homeTeamWinProbabilityAdded)) +
+  geom_boxplot(fill = 'purple') +
+  labs(x = "Down", y = "Home Team Win Probability Added") + 
+  theme(plot.margin = margin(t = 20, r = 30, b = 20, l = 10),
+        axis.title.x = element_text(size = 8),
+        axis.title.y = element_text(size = 8))
+
+plot2 <- ggplot(nfldata, aes(x = factor(quarter), y = homeTeamWinProbabilityAdded)) +
+  geom_boxplot(fill = "green") +
+  labs(x = "Quarter", y = "Win Probability") +
+  theme(plot.margin = margin(t = 20, r = 20, b = 20, l = 20),
+        axis.title.x = element_text(size = 8),
+        axis.title.y = element_text(size = 8))
+
+plot3 <- ggplot(nfldata, aes(x = preSnapVisitorScore, y = homeTeamWinProbabilityAdded)) +
+  geom_point(color = "red") +
+  labs(x = "Pre-Snap Visitor Score", y = "Win Probability") +
+  theme(plot.margin = margin(t = 20, r = 20, b = 20, l = 20),
+        axis.title.x = element_text(size = 8),
+        axis.title.y = element_text(size = 8))
+
+# Combine the plots using patchwork
+combined_plot <- plot1 + plot2 + plot3 + 
+  plot_layout(ncol = 2, nrow = 2) + 
+  plot_annotation(title = "Relationships Between Predictors and Win Probability Outcome",
+                  theme = theme(plot.title = element_text(hjust = 0.5)))
+
+# Print the combined plot
+print(combined_plot)
 

### 2024-11-17 NFL_IDS702Projectproposal.qmd
box plots

@@
 plot1 <- ggplot(nfldata, aes(x = factor(down), y = homeTeamWinProbabilityAdded)) +
-  geom_boxplot(fill = 'purple') +
+  geom_boxplot() +
   labs(x = "Down", y = "Home Team Win Probability Added") + 
@@
 plot2 <- ggplot(nfldata, aes(x = factor(quarter), y = homeTeamWinProbabilityAdded)) +
-  geom_boxplot(fill = "green") +
+  geom_boxplot() +
   labs(x = "Quarter", y = "Win Probability") +

### 2024-11-17 NFL_IDS702Projectproposal.qmd
d2q1 plot

--- 2024-11-17 NFL_IDS702Projectproposal.qmd
@@
 
-Independent Variables: **dropbackType**: The type of drop back after the
-snap by the QB (Traditional, Designed Rollout, Scramble, Scramble
-Rollout, Designed Rollout Left, Designed Rollout Right, Scramble Rollout
-Left, Scramble Rollout Right, Designed Run, QB Draw, Rollout, text)
-(categorical) \
+Independent Variables:
+
+**dropbackType**: The type of drop back after the snap by the QB
+(Traditional, Designed Rollout, Scramble, Scramble Rollout, Designed
+Rollout Left, Designed Rollout Right, Scramble Rollout Left, Scramble
+Rollout Right, Designed Run, QB Draw, Rollout, text) (categorical)\
 **dropbackDistance**: The distance the QB dropped back (yards) behind
-the center after the snap (numeric) \
+the center after the snap (numeric)\
 **timeToSack**: Time before the QB is sacked; this indicates the level
@@
 
-Independent Variable: **quarter**: the quarter the play is in may impact
-the delta of the teams probability of winning as any slip up in the
-latter half of the game may be more determental and harder to come back
-from. \
+Independent Variable:
+
+**quarter**: the quarter the play is in may impact the delta of the
+teams probability of winning as any slip up in the latter half of the
+game may be more determental and harder to come back from.\
 **down**: how many downs are left in the play impacts when the team has
@@
 glimpse(nfldata)
+
 ```
 
-**Exploratory Plots:** \
+**Exploratory Plots:**\
 \
@@
 ```{r}
-#Build a histogram of the win probably delta, "homeTeamWinProbabilityAdded"
+nfldata %>%
+  distinct(timeToSack)
+```
 
-ggplot(nfldata, aes(x = as.factor(quarter), y = preSnapHomeTeamWinProbability)) +
-  geom_line() +
-  labs(title = "Win Probability Before Play by Quarter (Home Team)", x = "Quarter", y = "Home Team Win Probability")
+```{r}
+nfldata_viz <- subset(
+  nfldata,
+  !is.na(dropbackType) & dropbackType != "UNKNOWN" & 
+    passResult %in% c("C", "I")
+)
+
+ggplot(nfldata_viz, aes(x = dropbackType, fill=passResult)) + 
+  geom_bar() + labs(x="dropbackType", y="dropbackDistance", 
+  fill='passResult', title = "passResult 
+  in different dropbackType") + theme(axis.text.x = 
+  element_text(angle = 45, vjust = 1, hjust=1)) +
+  scale_fill_discrete(labels=c('Complete', 
+  'Incomplete'))
+```
+
+Q1 Relationship of Interest:
+
+```{r}
+nfldata_viz <- nfldata_viz %>%
+  mutate(
+    dropbackType = ifelse(dropbackType %in% c("Traditional", "Rollout"), dropbackType, "Other")
+  )
+
+ggpairs(
+  nfldata_viz,
+  columns = c("dropbackType", "dropbackDistance" ),
+  aes(color = as.factor(passResult)), # Treat passResult as a factor
+  upper = list(continuous = "points"),
+  lower = list(continuous = "points")
+) +
+  labs(
+    title = "Pairwise Plots of Predictors Colored by Pass Result",
+    color = "Pass Result"
+  ) +
+  theme_minimal() +
+  theme(
+    plot.title = element_text(size = 16, hjust = 0.5, face = "bold"),
+    legend.title = element_text(size = 12),
+    legend.text = element_text(size = 10)
+  )
+
 ```
@@
 ```{r}
-#Create a plot that demonstrates the relationship between the game's characteristics and the probability of winning. maybe a graph on gameClock and homeTeamWinProbabilityAdded
+
+ggplot(nfldata, aes(x = preSnapHomeTeamWinProbability)) +
+  geom_histogram(binwidth = 0.5, fill = "skyblue") +
+  facet_grid(~  as.factor(quarter))
+  labs(title = "Win Probability Before Play by Quarter (Home Team)")
 

### 2024-11-17 NFL_IDS702Projectproposal.qmd
change the variable of nfl Q1

@@
 
-Outcome Variable: Pass Result (passResult) - This is a binary variable,
-with categories such as "Complete" (C) and "Incomplete" (I), indicating
-whether the pass was successfully completed.
+**Outcome Variable**: `passResult` (Binary: Complete/Incomplete),
+indicating whether the pass was successfully completed.
 
-Independent Variables:
+**Primary Predictors**:
 
-**dropbackType**: The type of drop back after the snap by the QB
-(Traditional, Designed Rollout, Scramble, Scramble Rollout, Designed
-Rollout Left, Designed Rollout Right, Scramble Rollout Left, Scramble
-Rollout Right, Designed Run, QB Draw, Rollout, text) (categorical)\
-**dropbackDistance**: The distance the QB dropped back (yards) behind
-the center after the snap (numeric)\
-**timeToSack**: Time before the QB is sacked; this indicates the level
-of pass protection (numeric) **rushLocationType**: Indicates where
-defensive pressure might be coming from; outside pressure can force
-quicker throws. (categorical)
+-   `dropbackDistance` (Continuous).The distance the QB dropped back
+    (yards) behind the center after the snap (numeric)
 
-Possible other variables:\
-**pff_passCoverage**: Defensive coverage scheme (e.g., man vs. zone);
-some schemes are harder to throw against.(categorica)\
-**pff_manZone**: Specific indicator of man or zone coverage.
-(categorical)\
-**unblockedPressure**: Whether there was pressure from an unblocked
-defender; unblocked pressure typically lowers completion probability.
-(binary)\
-**passTippedAtLine**: Whether the pass was tipped; this directly impacts
-completion probability.(binary)
+-   `dropbackType` (Categorical).The type of drop back after the snap by
+    the QB (Traditional, Designed Rollout, Scramble, Scramble Rollout,
+    Designed Rollout Left, Designed Rollout Right, Scramble Rollout
+    Left, Scramble Rollout Right, Designed Run, QB Draw, Rollout, text)
+    (categorical)
+
+**Contextual Variables:**
+
+-   `timeToThrow` (Continuous). The time (secs) elapsed between snap and
+    pass (numeric)
+
+-   `passLength` (Continuous). The distance beyond the LOS that the ball
+    traveled not including yards into the endzone. If thrown behind LOS,
+    the value is negative. (numeric)
+
+**Potential Interaction Terms:**
+
+-   `dropbackDistance* dropbackType` : Different types of dropbacks
+    (e.g., "Rollout" vs. "Traditional") may influence the effectiveness
+    of a specific dropback distance.
 
@@
 ```{r}
-nfldata %>%
-  distinct(timeToSack)
+# Clean and transform dataset
+nfldata_1 <- nfldata %>%
+  # Select relevant variables
+  select(dropbackDistance, dropbackType, timeToThrow, passLength, passResult) %>%
+  # Filter for valid pass results
+  filter(
+    passResult %in% c("C", "I"),
+    !is.na(dropbackType) & dropbackType != "UNKNOWN",
+    !is.na(dropbackDistance),
+    !is.na(timeToThrow),
+    !is.na(passLength)
+  ) %>%
+  # Convert variables to factors
+  mutate(
+    dropbackType = as.factor(dropbackType),
+    passResult = as.factor(passResult)
+  )
+glimpse(nfldata_1)
+
 ```
@@
 ```{r}
-nfldata_viz <- subset(
-  nfldata,
-  !is.na(dropbackType) & dropbackType != "UNKNOWN" & 
-    passResult %in% c("C", "I")
-)
-
-ggplot(nfldata_viz, aes(x = dropbackType, fill=passResult)) + 
+ggplot(nfldata_1, aes(x = dropbackType, fill=passResult)) + 
   geom_bar() + labs(x="dropbackType", y="dropbackDistance", 
   fill='passResult', title = "passResult
-  in different dropbackType") + theme(axis.text.x = 
+  in dropbackType") + theme(axis.text.x = 
   element_text(angle = 45, vjust = 1, hjust=1)) +
-  scale_fill_discrete(labels=c('Complete', 
-  'Incomplete'))
+  scale_fill_discrete(labels=c("Complete", "Incomplete"))
 ```
@@
 ```{r}
-nfldata_viz <- nfldata_viz %>%
+table(nfldata_1$dropbackType)
+```
+
+```{r}
+nfldata_1_viz <-nfldata_1 %>%
   mutate(
-    dropbackType = ifelse(dropbackType %in% c("Traditional", "Rollout"), dropbackType, "Other")
+    dropbackCategory = case_when(
+      dropbackType == "TRADITIONAL" ~ "Traditional",
+      dropbackType %in% c("DESIGNED_ROLLOUT_LEFT", "DESIGNED_ROLLOUT_RIGHT") ~ "Designed Rollout",
+      dropbackType %in% c("SCRAMBLE", "SCRAMBLE_ROLLOUT_LEFT", "SCRAMBLE_ROLLOUT_RIGHT") ~ "Scramble",
+      TRUE ~ NA_character_  # Handle unexpected values
+    )
   )
@@
 ggpairs(
-  nfldata_viz,
-  columns = c("dropbackType", "dropbackDistance" ),
+  nfldata_1_viz,
+  columns = c("dropbackCategory", "dropbackDistance" ),
   aes(color = as.factor(passResult)), # Treat passResult as a factor
@@
 ```{r}
-
 ggplot(nfldata, aes(x = preSnapHomeTeamWinProbability)) +

### 2024-11-17 NFL_IDS702Projectproposal.qmd
update nfl Q1

@@
   nfldata_1_viz,
-  columns = c("dropbackCategory", "dropbackDistance" ),
+  columns = c("dropbackCategory", "dropbackDistance","timeToThrow", "passLength" ),
   aes(color = as.factor(passResult)), # Treat passResult as a factor
@@
     title = "Pairwise Plots of Predictors Colored by Pass Result",
+    x = "Predictor 1",
+    y = "Predictor 2",
     color = "Pass Result"

### 2024-11-18 NFL_IDS702Projectproposal.qmd
change plot size

--- 2024-11-17 NFL_IDS702Projectproposal.qmd
@@
 library(dplyr)
+library(ggplot2)
+library(GGally)
 ```
@@
   nfldata_1_viz,
-  columns = c("dropbackCategory", "dropbackDistance" ),
+  columns = c("dropbackCategory", "dropbackDistance", "timeToThrow", "passLength"),
   aes(color = as.factor(passResult)), # Treat passResult as a factor
-  upper = list(continuous = "points"),
-  lower = list(continuous = "points")
+  upper = list(continuous =  wrap("points", size = 0.3)),
+  lower = list(continuous =  wrap("points", size = 0.3))
 ) +
@@
     legend.title = element_text(size = 12),
-    legend.text = element_text(size = 10)
+    legend.text = element_text(size = 10),
+    axis.text.x = element_text(size = 3.5),  # Adjust x-axis text size
+    axis.text.y = element_text(size = 8),  # Adjust y-axis text size
+    axis.ticks.x = element_blank(),       # Remove excess x-axis ticks
+    axis.ticks.y = element_blank()        # Remove excess y-axis ticks
   )

### 2024-11-27 project_proposal/Final_Proposal_CURJ_IDS702.qmd
deleted nfl data and added a project_proposal folder

--- 2024-11-17 CURJ_IDS702Projectproposal.qmd
@@
 title: "Project Proposal"
-author: "ALL group member names here"
+author: "Jenny Wu, Rishika Randev, Shiyue Zhou, Uzoma Uwazurike"
 subtitle: "Due November 17 at 11:59pm"
@@
     wrap: 72
+echo: FALSE
+warning: FALSE
 ---
@@
 
-**Load the datasets and prepare final dataset`**:
+**Load the datasets and prepare final dataset (code is hidden to reduce proposal length)**:
 
@@
 
-**Exploratory Plots:** Q1 Outcome Variable (Life Expectancy):
+**Exploratory Plots:** 
+Q1 Outcome Variable (Life Expectancy):
 
@@
                strip.text.x = element_text(size = 5),
-               strip.text.y = element_text(size = 4))
+               strip.text.y = element_text(size = 3))
 
@@
 
-**Data source:**
+**Data source:** NFL Kaggle 2025 Play Dataset
+https://www.kaggle.com/competitions/nfl-big-data-bowl-2025/data?select=player_play.csv
 
-**Brief description:**
+**Brief description:** The Play Data Dataset provides detailed
+play-by-play information about football games, capturing various aspects
+of gameplay, team dynamics, and player actions. This dataset includes
+variables that tell us more about what influences game play within a
+football game.
 
-**Research question 1:**
+**Research question 1:** Does the distance that a quarterback
+drops back after the snap and the dropback type affect the probability of a pass completion?
 
--   Outcome variable (include the name/description and type of
-    variable):
+**Outcome Variable**: `passResult` (Binary: Complete/Incomplete),
+indicating whether the pass was successfully completed.
 
-**Research question 2:**
+**Primary Predictors**:
 
--   Outcome variable (include the name/description and type of
-    variable):
+-   `dropbackDistance` (Continuous).The distance the QB dropped back
+    (yards) behind the center after the snap (numeric)
+
+-   `dropbackType` (Categorical).The type of drop back after the snap by
+    the QB (Traditional, Designed Rollout, Scramble, Scramble Rollout,
+    Designed Rollout Left, Designed Rollout Right, Scramble Rollout
+    Left, Scramble Rollout Right, Designed Run, QB Draw, Rollout, text)
+    (categorical)
+
+**Potential Interaction Terms:**
+
+-   `dropbackDistance* dropbackType` : Different types of dropbacks
+    (e.g., "Rollout" vs. "Traditional") may influence the effectiveness
+    of a specific dropback distance.
+
+**Research question 2:** How does the game situation (down, quarter, and opponent score) affect
+home team's win probability change after a play?
+
+**Outcome Variable**: `homeTeamWinProbabilityAdded` (Continuous),
+indicating how much the home team's win probability changes after a given play.
+
+**Primary Predictors**:
+
+-   `quarter` (Discrete), indicating whether the play is in the first, second, third, or fourth      quarter, or during overtime.
+
+-   `down` (Discrete), indicating whether the play is on a first, second, third, or fourth down.
+
+-   `preSnapVisitorScore` (Discrete), indicating the score of the oponent team prior to the          current play.
 
@@
 
-```{r data-2, warning=FALSE}
+```{r, warning=FALSE}
+nfldata <- read.csv("https://raw.githubusercontent.com/wujenny214/F24_STATS_FINAL/refs/heads/main/data/NFL/plays.csv") #replace with you file path name 
+glimpse(nfldata) 
+```
+
+**Exploratory Plots:**
+RQ1: Does the distance that a quarterback drops back after the snap
+affect the probability of a pass completion?
+
+Q1 Outcome Plot:
+
+```{r}
+# Clean and transform dataset
+nfldata_1 <- nfldata %>%
+  # Select relevant variables
+  select(dropbackDistance, dropbackType, timeToThrow, passLength, passResult) %>%
+  # Filter for valid pass results
+  filter(
+    passResult %in% c("C", "I"),
+    !is.na(dropbackType) & dropbackType != "UNKNOWN",
+    !is.na(dropbackDistance),
+    !is.na(timeToThrow),
+    !is.na(passLength)
+  ) %>%
+  # Convert variables to factors
+  mutate(
+    dropbackType = as.factor(dropbackType),
+    passResult = as.factor(passResult)
+  )
+glimpse(nfldata_1)
 
@@
 
-**Exploratory Plots:**
+```{r}
+ggplot(nfldata_1, aes(x = dropbackType, fill=passResult)) + 
+  geom_bar() + labs(x="dropbackType", y="dropbackDistance", 
+  fill='passResult', title = "Pass Result
+  by Dropback Type") + theme(axis.text.x = 
+  element_text(angle = 45, vjust = 1, hjust=1)) +
+  scale_fill_discrete(labels=c("Complete", "Incomplete"))
+```
 
-```{r plots-5}
+Q1 Relationship of Interest:
+
+```{r}
+table(nfldata_1$dropbackType)
+```
+
+```{r}
+nfldata_1_viz <-nfldata_1 %>%
+  mutate(
+    dropbackCategory = case_when(
+      dropbackType == "TRADITIONAL" ~ "Traditional",
+      dropbackType %in% c("DESIGNED_ROLLOUT_LEFT", "DESIGNED_ROLLOUT_RIGHT") ~ "Designed Rollout",
+      dropbackType %in% c("SCRAMBLE", "SCRAMBLE_ROLLOUT_LEFT", "SCRAMBLE_ROLLOUT_RIGHT") ~ "Scramble",
+      TRUE ~ NA_character_  # Handle unexpected values
+    )
+  )
+
+ggpairs(
+  nfldata_1_viz,
+  columns = c("dropbackCategory", "dropbackDistance", "timeToThrow", "passLength"),
+  aes(color = as.factor(passResult)), # Treat passResult as a factor
+  upper = list(continuous =  wrap("points", size = 0.3)),
+  lower = list(continuous =  wrap("points", size = 0.3))
+) +
+  labs(
+    title = "Pairwise Plots of Predictors Colored by Pass Result",
+    x = "Predictor 1",
+    y = "Predictor 2",
+    color = "Pass Result"
+  ) +
+  theme_minimal() +
+  theme(
+    plot.title = element_text(size = 14, hjust = 0.5, face = "bold"),
+    legend.title = element_text(size = 12),
+    legend.text = element_text(size = 10),
+    axis.text.x = element_text(size = 3.5),  # Adjust x-axis text size
+    axis.text.y = element_text(size = 8),  # Adjust y-axis text size
+    axis.ticks.x = element_blank(),       # Remove excess x-axis ticks
+    axis.ticks.y = element_blank(),     # Remove excess y-axis ticks
+    strip.text.y = element_text(size = 3)
+  )
+```
+
+RQ2: How does the game situation affect win probability change?
+
+Q2 Outcome Plot:
+
+```{r}
+ggplot(nfldata, aes(x = homeTeamWinProbabilityAdded)) +
+  geom_histogram(binwidth = 0.05, fill = "skyblue") + 
+  labs(title = "Home Team Win Probability Added Distribution",
+       x = "Home Team Win Probability Added",
+       y = "Count") +
+  theme_minimal()  
+```
+
+Q2 Relationship of Interest:
+```{r}
+#Create a plot that demonstrates the relationship between the game's characteristics and the probability of winning. 
+# Load the necessary packages
+library(ggplot2)
+library(patchwork)
+
+# Create the individual plots
+plot1 <- ggplot(nfldata, aes(x = factor(down), y = homeTeamWinProbabilityAdded)) +
+  geom_boxplot() +
+  labs(x = "Down", y = "Win Probability Change") + 
+  theme(plot.margin = margin(t = 20, r = 30, b = 20, l = 10),
+        axis.title.x = element_text(size = 8),
+        axis.title.y = element_text(size = 8))
+
+plot2 <- ggplot(nfldata, aes(x = factor(quarter), y = homeTeamWinProbabilityAdded)) +
+  geom_boxplot() +
+  labs(x = "Quarter", y = "Win Probability Change") +
+  theme(plot.margin = margin(t = 20, r = 20, b = 20, l = 20),
+        axis.title.x = element_text(size = 8),
+        axis.title.y = element_text(size = 8))
+
+plot3 <- ggplot(nfldata, aes(x = preSnapVisitorScore, y = homeTeamWinProbabilityAdded)) +
+  geom_point(color = "red") +
+  labs(x = "Pre-Snap Visitor Score", y = "Win Probability Change") +
+  theme(plot.margin = margin(t = 20, r = 20, b = 20, l = 20),
+        axis.title.x = element_text(size = 8),
+        axis.title.y = element_text(size = 8))
+
+# Combine the plots using patchwork
+combined_plot <- plot1 + plot2 + plot3 + 
+  plot_layout(ncol = 2, nrow = 2) + 
+  plot_annotation(title = "Relationships Between Predictors and Home Team Win Probability Change",
+                  theme = theme(plot.title = element_text(size=8, hjust = 0.5)))
+
+# Print the combined plot
+print(combined_plot)
 
 ```
-
-#
-
-# Dataset 3 (optional)
-
-**Data source:**
-
-**Brief description:**
-
-**Research question 1:**
-
--   Outcome variable (include the name/description and type of
-    variable):
-
-**Research question 2:**
-
--   Outcome variable (include the name/description and type of
-    variable):
-
-**Load the data and provide a `glimpse()`**:
-
-```{r data-3, warning=FALSE}
-
-```
-
-**Exploratory Plots:**
-
-```{r plots-3}
-
-```
-
-#
 

### 2024-11-28 county_data/conversion.py
changing the csvs into xlsx

@@ -1,0 +1,27 @@
+import pandas as pd
+import os
+
+# Directory containing CSV files
+input_directory = "/Users/jw/F24_STATS_FINAL/county_data"  # Replace with your CSV directory path
+output_directory = "/Users/jw/F24_STATS_FINAL/county_data"  # Replace with your output directory path
+
+
+# Create output directory if it doesn't exist
+os.makedirs(output_directory, exist_ok=True)
+
+# Loop through all files in the input directory
+for filename in os.listdir(input_directory):
+    if filename.endswith('.csv'):
+        csv_path = os.path.join(input_directory, filename)
+        xlsx_path = os.path.join(output_directory, f"{os.path.splitext(filename)[0]}.xlsx")
+        
+        # Read the CSV with the second row as the header
+        df = pd.read_csv(csv_path, header=1)
+        
+        # Save to Excel with the specified sheet name
+        with pd.ExcelWriter(xlsx_path, engine='openpyxl') as writer:
+            df.to_excel(writer, sheet_name="Additional Measure Data", index=False)
+        
+        print(f"Converted: {csv_path} -> {xlsx_path} with sheet title 'Additional Measure Data' and second row as header")
+
+print("All CSV files have been converted to XLSX format with the specified sheet name.")

### 2024-11-29 missing.qmd
Final dataset file, cleaning file, and missing file

---
title: "cleaning.qmd"
format: html
editor: visual
---

```{r}
library(dplyr)
library(tidyverse)
```

```{r}
combined <- read.csv('F24_STATS_FINAL/combined.csv')
```

```{r}
combined = subset(combined, select = -c(X))
head(combined)

```

```{r}
dim(combined)
```

```{r}
combined |> group_by(State) %>% summarise(total_count = n())
```

```{r}
sum(duplicated(combined))
```

```{r}
sum(duplicated(combined$FIPS))
```

```{r}
summary(combined)
```

```{r}
combined %>%
           group_by(State) %>%
           summarise(total_count = n(), Income_NA_count = sum(is.na(Median.Household.Income)), Uninsured_NA_count = sum(is.na(Percent.Uninsured.Adults)), Tax_NA_count = sum(is.na(Median.Prop.Tax)), Segregation_NA_count = sum(is.na(Residential.Segregation.Index)), Foods_NA_count = sum(is.na(Percent.Limited.Access.Healthy.Foods)), Expectancy_NA_count = sum(is.na(Life.Expectancy)), Adequacy_NA_count = sum(is.na(School.Funding.Adequacy)))
```

```{r}
na.patterns <- naclus(combined)
naplot(na.patterns, "na per var")
```

```{r}
plot(na.patterns)
```

```{r}
combined <- combined |>
  mutate(expectancy.miss = ifelse(is.na(Life.Expectancy),1,0))

expectancy_miss_mod <- glm(expectancy.miss ~ Residential.Segregation.Index+Percent.Limited.Access.Healthy.Foods+Percent.Uninsured.Adults, 
                    data=combined, family="binomial")

summary(expectancy_miss_mod)
```

```{r}
combined <- combined |>
  mutate(foods.miss = ifelse(is.na(Percent.Limited.Access.Healthy.Foods),1,0))

foods_miss_mod <- glm(foods.miss ~ Residential.Segregation.Index+Life.Expectancy+Percent.Uninsured.Adults, 
                    data=combined, family="binomial")

summary(foods_miss_mod)
```

```{r}
combined <- combined |>
  mutate(funding.miss = ifelse(is.na(School.Funding.Adequacy),1,0))

funding_miss_mod <- glm(funding.miss ~ Residential.Segregation.Index+Median.Prop.Tax+Median.Household.Income, 
                    data=combined, family="binomial")

summary(funding_miss_mod)
```

```{r}
library(mice)
library(sjlabelled)
library(tidyverse)

sub <- combined |> select(c("Life.Expectancy","Residential.Segregation.Index","Percent.Limited.Access.Healthy.Foods","Percent.Uninsured.Adults","Median.Household.Income", "School.Funding.Cat", "Median.Prop.Tax"))
sub <- unlabel(sub) #unlabel the data (labels cause problem for the mice function)
sub.imp <- mice(sub, m=5, method="pmm", print=FALSE)
```

```{r}
sub.mods <- with(sub.imp, #replace with. model
)
summary(pool(sub.mods))

```

### 2024-11-29 stats_dataset_cleaning.R

@@ -20,7 +20,7 @@ desired_cols <- c('FIPS', 'State',
                   "% Limited Access to Healthy Foods", 
                   "% Uninsured Adults", 
                   "Spending per Pupil", "School Funding Adequacy", 
-                  "Median Household Income" ) # Update with actual column names
+                  "Median Household Income" ) 
 
 # Base URL
 base_path <- "F24_STATS_FINAL/county_data/"
@@ -55,5 +55,51 @@ for (state in states) {
 # Combine all processed data frames into one
 combined <- do.call(rbind, all_data)
 
-# The final data frame `combined` contains all processed state data.
+#Create categorical variable for school funding adequacy
+combined_pre_merge <- combined |>
+  mutate(
+    School.Funding.Cat = factor(case_when(
+      School.Funding.Adequacy < 0 ~ 'Inadequate', 
+      School.Funding.Adequacy > 0  ~ 'Adequate',
+      TRUE ~ NA
+    ), levels=c('Inadequate', 'Adequate'), labels=c(0, 1)))
 
+#Merge with ACS (property tax) data
+acs <- read.csv("https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/county_data/acs_prop_2.csv")
+merged_df <- full_join(combined_pre_merge, acs, by = c("FIPS" = "fips")) %>%
+  mutate(
+    indicator = case_when(
+      is.na(State) & !is.na(state) ~ "Right Only",  # Row from acs only
+      !is.na(State) & is.na(state) ~ "Left Only",   # Row from combined_pre_merge only
+      !is.na(State) & !is.na(state) ~ "Both"        # Row present in both dfs
+    )
+  )
+
+print(merged_df[merged_df$indicator == "Right Only", ])
+print(merged_df[merged_df$indicator == "Left Only", ])
+#Difference of 8 rows, indicating that there are 8 counties with health data but no property tax data. 
+#Alaska and South Dakota are going to be dropped anyways due to missing residential segregation index data.
+#Looking into the VA county with missing property tax data (51515), it turns out that this county does not have any
+#health data either in the original spreadsheet, but is still recorded as a county. Upon further analysis
+#it was found that this FIPS code corresponded to Bedford City, which was merged into 51519 in 2010. Therefore it was safe to drop.
+
+merged_df <- merged_df[merged_df$FIPS != 51515, ]
[…]

### 2024-11-30 missing.qmd
Removing certain states

@@
 ```{r}
-combined <- read.csv('F24_STATS_FINAL/combined.csv')
+combined <- read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/combined.csv')
 ```
@@
 summary(pool(sub.mods))
+```
 
+```{r}
+head(combined.subset)
 ```
+
+```{r}
+combined.subset %>%
+           group_by(State) %>%
+           summarise(total_count = n(), Income_NA_count = sum(is.na(Median.Household.Income)), Uninsured_NA_count = sum(is.na(Percent.Uninsured.Adults)), Tax_NA_count = sum(is.na(Median.Prop.Tax)), Segregation_NA_count = sum(is.na(Residential.Segregation.Index)), Foods_NA_count = sum(is.na(Percent.Limited.Access.Healthy.Foods)), Expectancy_NA_count = sum(is.na(Life.Expectancy)), Adequacy_NA_count = sum(is.na(School.Funding.Adequacy)))
+```
+
+```{r}
+summary(combined.subset)
+```

### 2024-11-30 stats_dataset_cleaning.R

@@ -42,7 +42,8 @@ for (state in states) {
   # Rename columns
   renamed_data <- subset_data %>%
     rename(
-      "Residential.Segregation.Index" = "Segregation Index...200"
+      "Residential.Segregation.Index" = "Segregation Index...200",
+      "School.Funding.Adequacy" = "School Funding Adequacy"
     )
   
   # Drop the first row
@@ -66,7 +67,9 @@ combined_pre_merge <- combined |>
 
 #Merge with ACS (property tax) data
 acs <- read.csv("https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/county_data/acs_prop_2.csv")
-merged_df <- full_join(combined_pre_merge, acs, by = c("FIPS" = "fips")) %>%
+combined_pre_merge$FIPS <- as.numeric(combined_pre_merge$FIPS)
+
+merged_df <- full_join(combined_pre_merge, acs, join_by("FIPS" == "fips")) %>%
   mutate(
     indicator = case_when(
       is.na(State) & !is.na(state) ~ "Right Only",  # Row from acs only
@@ -95,11 +98,11 @@ regions <- read.csv("https://github.com/cphalpert/census-regions/raw/refs/heads/
 final_merged <- left_join(intermed_merged, regions, by = c("State" = "State"))
 final_merged = subset(final_merged, select = -c(State.Code))
 
-#Rename columns
-final_merged <- final_merged |> 
-  rename(
-    Percent.Uninsured.Adults = X..Uninsured.Adults, 
-    Percent.Limited.Access.Healthy.Foods = X..Limited.Access.to.Healthy.Foods)
+
+#Due to substantial missing data in certain states, remove them from the final dataframe
+states_to_exclude <- c('Alaska', 'Colorado', 'Idaho', 'Iowa', 'Kansas', 'Minnesota', 'Montana', 'Nebraska', 'North Dakota', 'South Dakota', 'Utah', 'Wyoming', 'Vermont')
+final <- final_merged %>%
+  filter(!State %in% states_to_exclude)
 
 # The final data frame `combined` contains all processed state data.
-write.csv(final_merged, '~/Documents/R\ Course\ Code/F24_STATS_FINAL/combined.csv')
+write.csv(final, '~/Documents/R\ Course\ Code/F24_STATS_FINAL/combined.csv')

### 2024-11-30 analysis.qmd
Adding analysis code file

---
title: "analysis"
format: html
editor: visual
---

## Load Data & Create Factors

```{r}
combined <- read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/combined.csv')
```

```{r}
head(combined)
```

```{r}
combined <- combined |> mutate(State_fac = factor(combined$State), Region_fac = factor(combined$Region, levels=c('South', 'West', 'Northeast', 'Midwest'), labels=c(0, 1, 2, 3)))
```

```{r}
combined <- combined %>% select(-X, -Spending.per.Pupil)
```

```{r}
head(combined)
```

## Imputation

```{r}

library(mice)
library(sjlabelled)
library(tidyverse)

```

```{r}
sub <- combined |> select(c("Life.Expectancy","Residential.Segregation.Index","X..Limited.Access.to.Healthy.Foods","X..Uninsured.Adults","Median.Household.Income", "School.Funding.Adequacy", "Median.Prop.Tax", "Region_fac"))
sub <- unlabel(sub) #unlabel the data (labels cause problem for the mice function)
sub.imp <- mice(sub, m=5, method="pmm", print=FALSE)
```

```{r}
#Need to convert school funding adequacy imputed values to school funding categorical variable before using in models
```

## Q1

```{r}
#MLR
#Outcome = life expectancy
#Predictors = residential segregation, limited access to healthy foods, uninsured adults, region_fac as interaction with residential segregation (or any other variable)
```

## Q2

```{r}
#Logistic 
#Outcome = School.Funding.Cat
#Predictors = residential segregation, median.household.income, median.prop.tax
```

### 2024-11-30 missing.qmd
Adding analysis code file

--- 2024-11-29 missing.qmd
@@
 library(tidyverse)
+library(Hmisc)
 ```
@@
 ```{r}
+combined <- combined %>%
+    rename(
+      "Perc.Uninsured.Adults" = "X..Uninsured.Adults",
+      "Perc.Limited.Access.to.Healthy.Foods" = "X..Limited.Access.to.Healthy.Foods"
+    )
+```
+
+```{r}
 combined %>%
            group_by(State) %>%
-           summarise(total_count = n(), Income_NA_count = sum(is.na(Median.Household.Income)), Uninsured_NA_count = sum(is.na(Percent.Uninsured.Adults)), Tax_NA_count = sum(is.na(Median.Prop.Tax)), Segregation_NA_count = sum(is.na(Residential.Segregation.Index)), Foods_NA_count = sum(is.na(Percent.Limited.Access.Healthy.Foods)), Expectancy_NA_count = sum(is.na(Life.Expectancy)), Adequacy_NA_count = sum(is.na(School.Funding.Adequacy)))
+           summarise(total_count = n(), Income_NA_count = sum(is.na(Median.Household.Income)), Uninsured_NA_count = sum(is.na(Perc.Uninsured.Adults)), Tax_NA_count = sum(is.na(Median.Prop.Tax)), Segregation_NA_count = sum(is.na(Residential.Segregation.Index)), Foods_NA_count = sum(is.na(Perc.Limited.Access.to.Healthy.Foods)), Expectancy_NA_count = sum(is.na(Life.Expectancy)), Adequacy_NA_count = sum(is.na(School.Funding.Adequacy)))
 ```
@@
 ```{r}
+colnames(combined)
+```
+
+```{r}
 library(mice)
@@
 
-sub <- combined |> select(c("Life.Expectancy","Residential.Segregation.Index","Percent.Limited.Access.Healthy.Foods","Percent.Uninsured.Adults","Median.Household.Income", "School.Funding.Cat", "Median.Prop.Tax"))
+sub <- combined |> select(c("Life.Expectancy","Residential.Segregation.Index","X..Limited.Access.to.Healthy.Foods","X..Uninsured.Adults","Median.Household.Income", "School.Funding.Cat", "Median.Prop.Tax"))
 sub <- unlabel(sub) #unlabel the data (labels cause problem for the mice function)
@@
 summary(pool(sub.mods))
-
 ```

### 2024-12-04 analysis.qmd
jw_push for mlr

@@
 
-```{r}
+```{r, echo= FALSE}
+library("dplyr")
+library(mice)
+library(sjlabelled)
+library(tidyverse)
+
 combined <- read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/combined.csv')
@@
 ```{r}
-combined <- combined |> mutate(State_fac = factor(combined$State), Region_fac = factor(combined$Region, levels=c('South', 'West', 'Northeast', 'Midwest'), labels=c(0, 1, 2, 3)))
+combined <- combined |> mutate(State_fac = factor(combined$State), Region_fac = factor(combined$Region, levels=c('South', 'West', 'Northeast', 'Midwest')))
 ```
@@
 ```{r}
-
-library(mice)
-library(sjlabelled)
-library(tidyverse)
-
-```
-
-```{r}
 sub <- combined |> select(c("Life.Expectancy","Residential.Segregation.Index","X..Limited.Access.to.Healthy.Foods","X..Uninsured.Adults","Median.Household.Income", "School.Funding.Adequacy", "Median.Prop.Tax", "Region_fac"))
@@
 #Need to convert school funding adequacy imputed values to school funding categorical variable before using in models
+# imputated data has 5 dataframes within a list, need to run a loop for all five datasets to create a categorical variable for each set. 
 ```
@@
 #MLR
-#Outcome = life expectancy
-#Predictors = residential segregation, limited access to healthy foods, uninsured adults, region_fac as interaction with residential segregation (or any other variable)
+# Outcome = life expectancy #Predictors = residential segregation, limited access to healthy foods, uninsured adults, region_fac as interaction with residential segregation (or any other variable)
+
+mlrmodel <- lm(
+  Life.Expectancy ~ Residential.Segregation.Index +
+    X..Limited.Access.to.Healthy.Foods +
+    X..Uninsured.Adults,
+  data = combined)
+summary(mlrmodel)
+
 ```
+
+```{r}
+mlr.imp.mod <- with(sub.imp, lm(Life.Expectancy ~ Residential.Segregation.Index +
+    X..Limited.Access.to.Healthy.Foods +
+    X..Uninsured.Adults + 
+    Region_fac))
+
+summary(pool(mlr.imp.mod))
+```
+
+```{r}
+mlr.imp.mod2 <- with(sub.imp, lm(Life.Expectancy ~ Residential.Segregation.Index +
+    X..Limited.Access.to.Healthy.Foods +
+    X..Uninsured.Adults + 
+    Region_fac + 
+    Region_fac * Residential.Segregation.Index))
+
+summary(pool(mlr.imp.mod2))
+```
+
+@ UZO: Add plots to check assumptions: QQ plot, nested F-test, write p-value interpretations, VIF, Cooks distance
 
@@
 #Logistic 
+lmod <- glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax,
+              data=combined,
+              family="binomial")
+summary(lmod)
+
+with(sub.imp, glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax, family="binomial"))
+
 #Outcome = School.Funding.Cat

### 2024-12-06 analysis.qmd
created the school.funding.cat variable

@@
 # imputated data has 5 dataframes within a list, need to run a loop for all five datasets to create a categorical variable for each set.
+
+# Create a function to classify school funding adequacy into categories
+# Create a function to classify school funding adequacy into categories
+categorize_school_funding <- function(x) {
+  ifelse(x < 50, "Low",
+         ifelse(x < 75, "Medium", "High"))
+}
+
+# Create an empty list to store the categorical variables
+sub.imp$imp$School.Funding.Cat <- list()
+
+# Loop over the 5 imputations
+for (i in 1:5) {
+  # Access the i-th set of imputed values for School.Funding.Adequacy
+  adequacy_values <- sub.imp$imp$School.Funding.Adequacy[, i]
+
+  # Apply the categorization function
+  sub.imp$imp$School.Funding.Cat[[i]] <- categorize_school_funding(adequacy_values)
+}
+
+# Verify the new categorical variables
+str(sub.imp$imp$School.Funding.Cat)
+
 ```
@@
 
-with(sub.imp, glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax, family="binomial"))
+```
+
+```{r}
+with(sub.imp, lmod, family="binomial")
 

### 2024-12-06 analysis.qmd
Diagnostics for Question 1

--- 2024-12-04 analysis.qmd
@@
 
-@ UZO: Add plots to check assumptions: QQ plot, nested F-test, write p-value interpretations, VIF, Cooks distance
+\@ UZO: Add plots to check assumptions: QQ plot, nested F-test, write p-value interpretations, VIF, Cooks distance
+
+```{r}
+#QQPlot
+qqnorm(residuals(mlrmodel))
+qqline(residuals(mlrmodel), col = "red") 
+```
+
+```{r}
+##Nested F-Test
+#Assuming we will drop %Uninsured Adults Predictor
+
+#print(colnames(combined))
+reduced_model <- lm(Life.Expectancy ~ Residential.Segregation.Index + X..Limited.Access.to.Healthy.Foods, data = combined)
+
+# Nested F-Test
+anova(reduced_model, mlrmodel)
+```
+
+```{r}
+#Cooks Distance 
+cooksd <- cooks.distance(mlrmodel)
+
+# Plot Cook's Distance
+plot(cooksd, main = "Cook's Distance", ylab = "Cook's Distance")
+abline(h = 4/(nrow(data) - length(coef(mlrmodel))), col = "red")  # Threshold line
+
+# Identify Influential Points
+influential <- which(cooksd > 4/(nrow(combined) - length(coef(mlrmodel))))
+print(influential)
+```
+
+```{r}
+#Variance Inflaction Factor (VIF)
+library(car)
+vif(mlrmodel)
+
+```
+
+Results are all slightly greater than one meaning some correlation between factors but not much. Percentage of uninsured adults having the greatest correlation
+
+```{r}
+#Generate 4 Diagnostic Plots at once
+par(mfrow = c(2, 2), mar = c(4,4,2,2))
+plot(mlrmodel)
+```
 

### 2024-12-07 analysis.qmd
diagnostics for q1

--- 2024-12-06 analysis.qmd
@@
 
+```{r}
+#QQPlot
+qqnorm(residuals(mlrmodel))
+qqline(residuals(mlrmodel), col = "red") 
+```
+
+```{r}
+##Nested F-Test
+#Assuming we will drop %Uninsured Adults Predictor
+
+#print(colnames(combined))
+reduced_model <- lm(Life.Expectancy ~ Residential.Segregation.Index + X..Limited.Access.to.Healthy.Foods, data = combined)
+
+# Nested F-Test
+anova(reduced_model, mlrmodel)
+```
+
+```{r}
+#Cooks Distance 
+cooksd <- cooks.distance(mlrmodel)
+
+# Plot Cook's Distance
+plot(cooksd, main = "Cook's Distance", ylab = "Cook's Distance")
+abline(h = 4/(nrow(data) - length(coef(mlrmodel))), col = "red")  # Threshold line
+
+# Identify Influential Points
+influential <- which(cooksd > 4/(nrow(combined) - length(coef(mlrmodel))))
+print(influential)
+```
+
+```{r}
+#Variance Inflaction Factor (VIF)
+library(car)
+vif(mlrmodel)
+
+```
+
+Results are all slightly greater than one meaning some correlation between factors but not much. Percentage of uninsured adults having the greatest correlation
+
+```{r}
+#Generate 4 Diagnostic Plots at once
+par(mfrow = c(2, 2), mar = c(4,4,2,2))
+plot(mlrmodel)
+```
+
 ## Q2

### 2024-12-08 analysis.qmd
added q2 code

--- 2024-12-06 analysis.qmd
@@
 
-```{r, echo= FALSE}
+```{r, echo= FALSE, warning = FALSE}
 library("dplyr")
@@
 #Need to convert school funding adequacy imputed values to school funding categorical variable before using in models
-# imputated data has 5 dataframes within a list, need to run a loop for all five datasets to create a categorical variable for each set. 
+# Imputated data has 5 dataframes within a list, need to run a loop for all five datasets to create a categorical variable for each set.
+
+sub <- combined |> select(c("Life.Expectancy","Residential.Segregation.Index","X..Limited.Access.to.Healthy.Foods","X..Uninsured.Adults","Median.Household.Income", "School.Funding.Adequacy", "Median.Prop.Tax", "Region_fac"))
+sub <- unlabel(sub) #unlabel the data (labels cause problem for the mice function)
+
+# Perform the imputation
+sub$School.Funding.Cat <- ifelse(sub$School.Funding.Adequacy > 0, 1, 0)  # Add the new variable to the dataset
+sub.imp <- mice(sub, m = 5, method = "pmm", print = FALSE)
+
+```
+
+```{r}
+# To create a School.Funding.Categorical variable, create a function and run the loop
+
+# Define the categorization function
+categorize_school_funding <- function(x) {
+  ifelse(x > 0, 1, 0)  # Binary classification
+}
+
+# Loop over the imputations
+for (i in 1:5) {
+  # Extract the imputed values for the i-th dataset
+  imputed_values <- sub.imp$imp$School.Funding.Adequacy[, i]
+
+  # Categorize the imputed values
+  categorized_values <- categorize_school_funding(imputed_values)
+  complete_cat <- rep(NA, nrow(sub))  # Initialize with NAs
+
+  # Get the indices where imputation occurred
+  imp_indices <- which(!is.na(categorized_values))  # Only for imputed values
+
+  # Assign the categorized values to the rows that were imputed
+  complete_cat[imp_indices] <- categorized_values[imp_indices]
+
+  # Save into the `imp` list of the `sub.imp` object
+  sub.imp$imp$School.Funding.Cat[[i]] <- complete_cat[imp_indices]
+}
+```
+
+```{r}
+long_data <- complete(sub.imp, action = "long", include = TRUE)  # Include original data
+
+# Check the structure
+str(long_data)
+
+# Ensure that `School.Funding.Cat` is correctly populated
+long_data$School.Funding.Cat <- ifelse(long_data$School.Funding.Adequacy > 0, 1, 0)
+
+if (!".imp" %in% colnames(long_data) || !".id" %in% colnames(long_data)) {
+  stop("`.imp` or `.id` columns are missing in the long data!")
+}
 ```
@@
 vif(mlrmodel)
-
 ```
@@
 ```{r}
-#Logistic 
+## Checking to see if the imputated data has a large effect on the outcome
+
+# Logistic Regression without Imputated Data
+# Outcome = School.Funding.Cat
+# Predictors = residential segregation, median.household.income, median.prop.tax
+
 lmod <- glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax,
@@
               family="binomial")
-summary(lmod)
 
-with(sub.imp, glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax, family="binomial"))
+library("modelsummary")
+modelsummary(lmod,
+             fmt = fmt_significant(2),
+             shape = term ~ model + statistic,
+             statistic = c("std.error","conf.int","p.value"),
+             exponentiate = FALSE,
+             gof_map=NA)
 
-#Outcome = School.Funding.Cat
-#Predictors = residential segregation, median.household.income, median.prop.tax
+modelsummary(lmod,
+             fmt = fmt_significant(2),
+             shape = term ~ model + statistic,
+             statistic = c("std.error","conf.int","p.value"),
+             exponentiate = TRUE,
+             gof_map=NA)
+
 ```
+
+```{r}
+imp_mods <- with(sub_imp, glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax), family="binomial")
+
+modelsummary(pool(imp_mods),
+             fmt = fmt_significant(2),
+             shape = term ~ model + statistic,
+             statistic = c("std.error","conf.int","p.value"),
+             exponentiate = FALSE,
+             gof_map=NA)
+
+modelsummary(pool(imp_mods),
+             fmt = fmt_significant(2),
+             shape = term ~ model + statistic,
+             statistic = c("std.error","conf.int","p.value"),
+             exponentiate = TRUE,
+             gof_map=NA)
+```
+
+```{r}
+# Creating the AUC curve
+
+library(caret)
+library(pROC)
+
+set.seed(42)
+train_idx <- sample(1:nrow(long_data), size = 0.7 * nrow(long_data))
+train_data <- long_data[train_idx, ]
+test_data <- long_data[-train_idx, ]
+
+glm_model <- glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax, data = train_data, family = binomial)
+
+# Step 4: Make predictions on the test set
+test_probs <- predict(glm_model, newdata = test_data, type = "response")
+test_pred <- ifelse(test_probs > 0.5, 1, 0)  # Classify based on threshold 0.5
+
+# Step 5: Create a confusion matrix
+conf_matrix <- confusionMatrix(factor(test_pred), factor(test_data$School.Funding.Cat))
+print(conf_matrix)
+```

### 2024-12-08 analysis.qmd
reformatted qmd

@@
 ---
-title: "analysis"
+title: "Understanding the Impact of Socioeconomic and Regional Factors on Health and Educational Outcomes"
+author: "Rishika Randev, Jenny Wu, Uzoma Uwazurike Jr., Shiyue Zhou"
+instructor: "Andrea Lane"
+institution: "Duke University, IDS 702 Data Modeling and Presentation"
 format: html
 editor: visual
+execute:
+  echo: false
+  warning: false
 ---
 
-## Load Data & Create Factors
+## **Abstract**
 
-```{r, echo= FALSE, warning = FALSE}
+## **Introduction**
+
+### Background
+
+### Research Question
+
+### Data Overview
+
+## **Methods**
+
+### Variable Selection
+
+### Exploratory Data Analysis & Data Preprocessing
+
+### Model Fitting & Assessment
+
+```{r, echo= FALSE, warning = FALSE, message=FALSE}
 library("dplyr")
@@
 library(tidyverse)
+library("modelsummary")
+```
 
+```{r,echo= FALSE, warning = FALSE, message=FALSE}
 combined <- read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/combined.csv')
@@
 
-```{r}
-head(combined)
-```
-
-```{r}
+```{r, echo= FALSE, warning = FALSE}
+# Making the state_fac variable a factor variable for future use
 combined <- combined |> mutate(State_fac = factor(combined$State), Region_fac = factor(combined$Region, levels=c('South', 'West', 'Northeast', 'Midwest')))
@@
 
-```{r}
-combined <- combined %>% select(-X, -Spending.per.Pupil)
-```
-
-```{r}
-head(combined)
-```
-
-## Imputation
-
-```{r}
-sub <- combined |> select(c("Life.Expectancy","Residential.Segregation.Index","X..Limited.Access.to.Healthy.Foods","X..Uninsured.Adults","Median.Household.Income", "School.Funding.Adequacy", "Median.Prop.Tax", "Region_fac"))
-sub <- unlabel(sub) #unlabel the data (labels cause problem for the mice function)
-sub.imp <- mice(sub, m=5, method="pmm", print=FALSE)
-```
-
-```{r}
-#Need to convert school funding adequacy imputed values to school funding categorical variable before using in models
-# Imputated data has 5 dataframes within a list, need to run a loop for all five datasets to create a categorical variable for each set.
+```{r, echo= FALSE, warning = FALSE}
+## Imputation Code 
+# In this block, we are looking to use predictive mean matching to combing regression methods nad hot deck imputing to predicted value closest to the predicted value of the observation with missing data we also needed to convert school funding adequacy imputed values to school funding categorical variable before using in models
 
@@
 
+# Add School.Funding.Cat to the orginal dataset.
+Since we do not want to imputate on the categorical outcome variable,  this creation of a variable exists to create the proper dataframe within the imputated data
+sub$School.Funding.Cat <- ifelse(sub$School.Funding.Adequacy > 0, 1, 0)
+
 # Perform the imputation
-sub$School.Funding.Cat <- ifelse(sub$School.Funding.Adequacy > 0, 1, 0)  # Add the new variable to the dataset
 sub.imp <- mice(sub, m = 5, method = "pmm", print = FALSE)
@@
 
-```{r}
-# To create a School.Funding.Categorical variable, create a function and run the loop
+```{r, echo= FALSE, warning = FALSE}
+# Since we should not imputate on the outcome, but need to imputate the School.Funding.Adequacy variable, we are feeding the School.Funding.Adequecy variable into a ifelse conditional that will replaced the impuated outcome variable.
 
@@
 
-```{r}
-long_data <- complete(sub.imp, action = "long", include = TRUE)  # Include original data
+```{r, echo= FALSE, warning = FALSE, include = FALSE }
+# To have a full dataset, we are taking the first imputated dataset R provided and using that as our full dataset sample. 
+completed_data <- complete(sub.imp, 1)
 
 # Check the structure
-str(long_data)
+str(completed_data)
 
-# Ensure that `School.Funding.Cat` is correctly populated
-long_data$School.Funding.Cat <- ifelse(long_data$School.Funding.Adequacy > 0, 1, 0)
+# Ensure that there are no missing data 
+colSums(is.na(completed_data))
 
-if (!".imp" %in% colnames(long_data) || !".id" %in% colnames(long_data)) {
-  stop("`.imp` or `.id` columns are missing in the long data!")
-}
 ```
 
-## Q1
+## Q1 Life Expectancy; MLR
 
-```{r}
-#MLR
-# Outcome = life expectancy #Predictors = residential segregation, limited access to healthy foods, uninsured adults, region_fac as interaction with residential segregation (or any other variable)
+### **Research Question 1: Life Expectancy**
+
+### Model Fitting
+
+```{r, echo= FALSE, warning = FALSE}
+# MLR
+# Outcome = life expectancy 
+# Predictors = residential segregation, limited access to healthy foods, uninsured adults, region_fac as interaction with residential segregation (or any other variable)
+
+# To get a general sense of what the data tells us, we use the orignial dataset with missing values.
 
@@
   data = combined)
-summary(mlrmodel)
+
+# Renaming the output names 
+renamed <- c(
+  "Residential.Segregation.Index" = "Residential Segregation Index (RSI)",
+  "X..Limited.Access.to.Healthy.Foods" = "Access to Healthy Foods",
+  "X..Uninsured.Adults" = "Pct. of Uninsured Adults")
+
+modelsummary(mlrmodel, fmt = fmt_significant(2),
+             coef_rename = renamed, 
+             shape = term ~ model + statistic,
+             statistic = c("std.error","conf.int","p.value"),
+             exponentiate = FALSE,
+             gof_map=NA,
+             )
 
@@
 
-```{r}
+### Regression Results
+
+```{r, echo= FALSE, warning = FALSE}
+# Running a regression with the imputated data to understand the impact of missing data
+
 mlr.imp.mod <- with(sub.imp, lm(Life.Expectancy ~ Residential.Segregation.Index +
@@
 
-summary(pool(mlr.imp.mod))
+renamed <- c(
+  "Residential.Segregation.Index" = "Residential Segregation Index (RSI)",
+  "X..Limited.Access.to.Healthy.Foods" = "Access to Healthy Foods",
+  "X..Uninsured.Adults" = "Pct. of Uninsured Adults",
+  "Region_facWest" = "Region.W", 
+  "Region_facNortheast" = "Region.NE",
+  "Region_facMidwest" = "Region.MidW"
+  )
+
+modelsummary(mlr.imp.mod, fmt = fmt_significant(2),
+             coef_rename = renamed, 
+             shape = term ~ model + statistic,
+             statistic = c("std.error","conf.int","p.value"),
+             exponentiate = FALSE,
+             gof_map=NA,
+             )
 ```
 
-```{r}
+```{r echo= FALSE, warning = FALSE}
+# Running a regression with the interaction term between residential segregation on the regions
+
 mlr.imp.mod2 <- with(sub.imp, lm(Life.Expectancy ~ Residential.Segregation.Index +
@@
 
-summary(pool(mlr.imp.mod2))
+renamed <- c(
+  "Residential.Segregation.Index" = "Residential Segregation Index (RSI)",
+  "X..Limited.Access.to.Healthy.Foods" = "Access to Healthy Foods",
+  "X..Uninsured.Adults" = "Pct. of Uninsured Adults",
+  "Region_facWest" = "Region.W", 
+  "Region_facNortheast" = "Region.NE",
+  "Region_facMidwest" = "Region.MidW", 
+  "Residential.Segregation.Index:Region_facWest" = "RSI * Region.W ", 
+  "Residential.Segregation.Index:Region_facNortheast" = "RSI * Region.NE",
+  "Residential.Segregation.Index:Region_facMidwest" = "RSI * Region.MidW"
+  )
+
+modelsummary(mlr.imp.mod2, fmt = fmt_significant(2),
+             coef_rename = renamed, 
+             shape = term ~ model + statistic,
+             statistic = c("std.error","conf.int","p.value"),
+             exponentiate = FALSE,
+             gof_map=NA,
+             )
+
 ```
 
-\@ UZO: Add plots to check assumptions: QQ plot, nested F-test, write p-value interpretations, VIF, Cooks distance
+### Evaluation & Assessment of Model
 
@@
 anova(reduced_model, mlrmodel)
+
 ```
@@
 
+```
+
+```{r include = FALSE}
 # Identify Influential Points
@@
 
-## Q2
+## Q2 **School Funding Adequacy; GLM**
+
+### **Research Question 2: School Funding Adequacy**
+
+### Model Fitting
 
@@
 library("modelsummary")
+
+#log-odds scale
 modelsummary(lmod,
@@
 
+### Regression Results
+
 ```{r}
-imp_mods <- with(sub_imp, glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax), family="binomial")
+imp_mods <- with(sub.imp, glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax), family="binomial")
 
+#log-odds scale
 modelsummary(pool(imp_mods),
@@
 
+### Evaluation and Assessment of Model
+
 ```{r}
-# Creating the AUC curve
+# Creating the confusion matrix
 
@@
 
-set.seed(42)
-train_idx <- sample(1:nrow(long_data), size = 0.7 * nrow(long_data))
-train_data <- long_data[train_idx, ]
-test_data <- long_data[-train_idx, ]
+# set.seed(42)
+# train_idx <- sample(1:nrow(long_data), size = 0.5 * nrow(long_data))
+# train_data <- long_data[train_idx, ]
+# test_data <- long_data[-train_idx, ]
 
-glm_model <- glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax, data = train_data, family = binomial)
+t_glm_model <- glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax, data = completed_data, family = binomial)
 
-# Step 4: Make predictions on the test set
-test_probs <- predict(glm_model, newdata = test_data, type = "response")
+# Make predictions on the test set
+test_probs <- predict(t_glm_model, newdata = completed_data, type = "response")
 test_pred <- ifelse(test_probs > 0.5, 1, 0)  # Classify based on threshold 0.5
+```
 
-# Step 5: Create a confusion matrix
-conf_matrix <- confusionMatrix(factor(test_pred), factor(test_data$School.Funding.Cat))
-print(conf_matrix)
+```{r}
+# Create a confusion matrix
+library("caret")
+conf_matrix <- confusionMatrix(factor(test_pred), factor(completed_data$School.Funding.Cat))
+#print(conf_matrix)
+
+library(pander)
+
+# Format and display confusion matrix as a Markdown table
+pander(conf_matrix, caption = "Confusion Matrix")
+
 ```
+
+```{r}
+#Creating the AUC curves
+roc_result <- roc(
+  completed_data$School.Funding.Cat,
+  predict(t_glm_model, type = "response"),
+  legacy.axes = TRUE,
+  plot = TRUE,
+  print.thres = 0.5,
+  print.auc = FALSE  # Disable automatic AUC
+)
+
+# Add AUC manually
+text(
+  x = 1.2,  # Adjust X-coordinate
+  y = 1,  # Adjust Y-coordinate
+  labels = sprintf("AUC = %.3f", auc(roc_result)),  # Add AUC value
+  cex = 1,  # Label size
+  col = "red"  # Label color
+)
+```
+
+```{r}
+
+```

### 2024-12-08 analysis.qmd

@@
 library("modelsummary")
+library(car)
 ```
@@
 
-#print(colnames(combined))
 reduced_model <- lm(Life.Expectancy ~ Residential.Segregation.Index + X..Limited.Access.to.Healthy.Foods, data = combined)
@@
 #Variance Inflaction Factor (VIF)
-library(car)
+
 vif(mlrmodel)

### 2024-12-08 analysis.qmd

@@
 ```{r, echo= FALSE, warning = FALSE}
-# Since we should not imputate on the outcome, but need to imputate the School.Funding.Adequacy variable, we are feeding the School.Funding.Adequecy variable into a ifelse conditional that will replaced the impuated outcome variable.
+# Since we should not imputate on the outcome, but need to imputate the School.Funding.Adequacy variable, we are feeding the School.Funding.Adequacy variable into a ifelse conditional that will replaced the impuated outcome variable.
 
@@
   X..Limited.Access.to.Healthy.Foods +
-    X..Uninsured.Adults,
+  X..Uninsured.Adults +
+  Region_fac,
   data = combined)
@@
   "X..Limited.Access.to.Healthy.Foods" = "Access to Healthy Foods",
-  "X..Uninsured.Adults" = "Pct. of Uninsured Adults")
+  "X..Uninsured.Adults" = "Pct. of Uninsured Adults",
+  "Region_facWest" = "Region.W", 
+  "Region_facNortheast" = "Region.NE",
+  "Region_facMidwest" = "Region.MidW")
 

### 2024-12-09 code.qmd
code and report qmds

--- 2024-12-08 analysis.qmd
@@
 
-## **Abstract**
+### Exploratory Data Analysis & Data Preprocessing
 
-## **Introduction**
-
-### Background
-
-### Research Question
-
-### Data Overview
-
-## **Methods**
-
-### Variable Selection
-
-### Exploratory Data Analysis & Data Preprocessing
+(Rishika will add here)
 
@@
 
-```{r, echo= FALSE, warning = FALSE}
+```{r, warning = FALSE}
 # Making the state_fac variable a factor variable for future use
@@
 
-```{r, echo= FALSE, warning = FALSE}
+```{r, warning = FALSE}
 ## Imputation Code 
-# In this block, we are looking to use predictive mean matching to combing regression methods nad hot deck imputing to predicted value closest to the predicted value of the observation with missing data we also needed to convert school funding adequacy imputed values to school funding categorical variable before using in models
+# In this block, we are looking to use predictive mean matching to combing regression methods and hot deck imputing to predicted value closest to the predicted value of the observation with missing data; we also needed to convert school funding adequacy imputed values to school funding categorical variable before using in models
 
@@
 # Perform the imputation
-sub.imp <- mice(sub, m = 5, method = "pmm", print = FALSE)
+sub.imp <- mice(sub, m = 1, method = "pmm", print = FALSE)
 
@@
 
-```{r, echo= FALSE, warning = FALSE}
-# Since we should not imputate on the outcome, but need to imputate the School.Funding.Adequacy variable, we are feeding the School.Funding.Adequacy variable into a ifelse conditional that will replaced the impuated outcome variable.
+```{r, warning = FALSE}
+# Since we should not imputate on the outcome, but need to imputate the School.Funding.Adequacy variable, we are feeding the School.Funding.Adequacy variable into a ifelse conditional that will replaced the imputed outcome variable.
 
@@
 # Loop over the imputations
-for (i in 1:5) {
+for (i in 1:1) {
   # Extract the imputed values for the i-th dataset
@@
 
-```{r, echo= FALSE, warning = FALSE, include = FALSE }
-# To have a full dataset, we are taking the first imputated dataset R provided and using that as our full dataset sample. 
+```{r, warning = FALSE}
+# To have a full dataset, we are taking the first imputed dataset R provided and using that as our full dataset sample. 
 completed_data <- complete(sub.imp, 1)
@@
 
-### Model Fitting
+### Model Fitting (Original Data)
 
-```{r, echo= FALSE, warning = FALSE}
+```{r, warning = FALSE}
 # MLR
@@
 mlrmodel <- lm(
-  Life.Expectancy ~ Residential.Segregation.Index +
+  Life.Expectancy ~ Residential.Segregation.Index * Region_fac +
   X..Limited.Access.to.Healthy.Foods +
-  X..Uninsured.Adults +
-  Region_fac,
+  X..Uninsured.Adults,
   data = combined)
@@
 
-modelsummary(mlrmodel, fmt = fmt_significant(2),
+summary(mlrmodel, fmt = fmt_significant(2),
              coef_rename = renamed, 
@@
 
-### Regression Results
+### Model Fitting (Imputed Data)
 
-```{r, echo= FALSE, warning = FALSE}
-# Running a regression with the imputated data to understand the impact of missing data
+test
 
-mlr.imp.mod <- with(sub.imp, lm(Life.Expectancy ~ Residential.Segregation.Index +
+```{r}
+# Running a regression with the imputated data (completed_data) to understand the impact of missing data
+
+mlr.imp <- lm(Life.Expectancy ~ Residential.Segregation.Index * Region_fac +
     X..Limited.Access.to.Healthy.Foods +
-    X..Uninsured.Adults + 
-    Region_fac))
+    X..Uninsured.Adults, data=completed_data)
 
@@
 
-modelsummary(mlr.imp.mod, fmt = fmt_significant(2),
+summary(mlr.imp, fmt = fmt_significant(2),
              coef_rename = renamed, 
@@
 ```{r echo= FALSE, warning = FALSE}
-# Running a regression with the interaction term between residential segregation on the regions
+# Running a regression without the interaction term between residential segregation on the regions
 
-mlr.imp.mod2 <- with(sub.imp, lm(Life.Expectancy ~ Residential.Segregation.Index +
+mlr.imp.reduced <- lm(Life.Expectancy ~ Residential.Segregation.Index +
     X..Limited.Access.to.Healthy.Foods +
     X..Uninsured.Adults + 
-    Region_fac + 
-    Region_fac * Residential.Segregation.Index))
+    Region_fac, data=completed_data)
 
@@
 
-modelsummary(mlr.imp.mod2, fmt = fmt_significant(2),
+summary(mlr.imp.reduced, fmt = fmt_significant(2),
              coef_rename = renamed, 
@@
 
+FINAL FULL MODEL: mlr.imp
+
+FINAL REDUCED MODEL: mlr.imp.reduced (without region \* segregation interactino)
+
 ### Evaluation & Assessment of Model
@@
 #QQPlot
-qqnorm(residuals(mlrmodel))
-qqline(residuals(mlrmodel), col = "red") 
+qqnorm(residuals(mlr.imp))
+qqline(residuals(mlr.imp), col = "red") 
 ```
@@
 ```{r}
-##Nested F-Test
-#Assuming we will drop %Uninsured Adults Predictor
-
-reduced_model <- lm(Life.Expectancy ~ Residential.Segregation.Index + X..Limited.Access.to.Healthy.Foods, data = combined)
-
-# Nested F-Test
-anova(reduced_model, mlrmodel)
+##Nested F-Test to see if interaction term makes a difference
+anova(mlr.imp, mlr.imp.reduced)
 
@@
 #Cooks Distance 
-cooksd <- cooks.distance(mlrmodel)
+cooksd <- cooks.distance(mlr.imp)
 
@@
 plot(cooksd, main = "Cook's Distance", ylab = "Cook's Distance")
-abline(h = 4/(nrow(data) - length(coef(mlrmodel))), col = "red")  # Threshold line
+abline(h = 4/(nrow(data) - length(coef(mlr.imp))), col = "red")  # Threshold line
 
@@
 # Identify Influential Points
-influential <- which(cooksd > 4/(nrow(combined) - length(coef(mlrmodel))))
+influential <- which(cooksd > 4/(nrow(combined) - length(coef(mlr.imp))))
 print(influential)
@@
 
-vif(mlrmodel)
+vif(mlr.imp)
 ```
 
-Results are all slightly greater than one meaning some correlation between factors but not much. Percentage of uninsured adults having the greatest correlation
+```{r}
+vif(mlr.imp.reduced)
+```
+
+Please check first one!!
 
@@
 par(mfrow = c(2, 2), mar = c(4,4,2,2))
-plot(mlrmodel)
+plot(mlr.imp)
 ```
@@
 #log-odds scale
-modelsummary(lmod,
+summary(lmod,
              fmt = fmt_significant(2),
@@
 
-modelsummary(lmod,
+summary(lmod,
              fmt = fmt_significant(2),
@@
 ```{r}
-imp_mods <- with(sub.imp, glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax), family="binomial")
+imp_mods <- glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax, data=completed_data, family="binomial")
 
 #log-odds scale
-modelsummary(pool(imp_mods),
+summary(imp_mods,
              fmt = fmt_significant(2),
@@
 
-modelsummary(pool(imp_mods),
+summary(imp_mods,
              fmt = fmt_significant(2),
@@
 
-t_glm_model <- glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax, data = completed_data, family = binomial)
+#t_glm_model <- glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax, data = completed_data, family = binomial)
 
 # Make predictions on the test set
-test_probs <- predict(t_glm_model, newdata = completed_data, type = "response")
+test_probs <- predict(imp_mods, newdata = completed_data, type = "response")
 test_pred <- ifelse(test_probs > 0.5, 1, 0)  # Classify based on threshold 0.5
@@
   completed_data$School.Funding.Cat,
-  predict(t_glm_model, type = "response"),
+  predict(imp_mods, type = "response"),
   legacy.axes = TRUE,

### 2024-12-09 paper.qmd
code and report qmds

--- 2024-12-08 analysis.qmd
@@
 
-### Research Question
+Life expectancy and school funding are important indicators of public well-being because they capture fundamental aspects of health, education, and social equity that are essential for the development of individuals and communities.
+Life expectancy is an important health indicator that reflects not only the overall health of a society, but also its stability, economic resilience, and levels of inequality.
+For example, differences in life expectancy between regions and socioeconomic groups often reveal systemic inequalities, with marginalized communities experiencing shorter life spans due to limited access to health care, healthy living environments, and economic opportunities.
+Similarly, adequately funded schools mean that students have equal access to learning opportunities and are an important investment in social mobility.
+Underfunded schools are often located in economically disadvantaged or segregated areas, exacerbating educational gaps and limiting students’ potential to break the cycle of poverty.
+These gaps are rooted in systemic inequalities, such as residential segregation, which concentrates poverty, limits access to quality education and health care, and perpetuates income and health disparities across generations.
+By addressing these inequalities, policymakers can foster fairer, more prosperous communities and ensure that everyone benefits from improved health and education systems.
+
+### Research Questions
+
+For this purpose, we focus on two key research questions to explore these relationships.  First, we investigate how socioeconomic factors—residential segregation, access to healthy foods, and health insurance coverage—impact average life expectancy.
+This analysis includes a comparison across different regions to identify variations in the relationship between residential segregation and life expectancy.
+Second, we evaluate how residential segregation, property taxes, and median household income influence school funding adequacy.
+This question allows us to assess educational investment disparities among regions with varying economic conditions.
+The observational unit for this analysis was the U.S.
+county.
 
 ### Data Overview
+
+The dataset used in this study integrates information from three primary sources.
+
+The first source was the County Health Rankings & Roadmaps program, associated with the University of Wisconsin Population Health Institute, which provided us with 2024 datasets for all U.S.
+states.
+This program consolidates the latest county-level measurements of various population health, economic, demographic, and social factors from the American Community Survey, the USDA Food Environment Atlas, Small Area Income and Poverty Estimates, and other reliable government sources into publicly available datasets every year.
+Our specific variables of interest were originally collected as part of:
+
+-   National Center for Health Statistics - Natality & Mortality Files, 2019-2021 (average life expectancy)
+
+-   American Community Survey 5-year estimates, 2018-2022 (residential segregation index & median household income)
+
+-   USDA Food Environment Atlas, 2019 (percentage of population with limited access to healthy foods)
+
+-   Small Area Health Insurance Estimates, 2021 (percentage of adults uninsured)
+
+-   School Finance Indicators Database, 2021 (school funding adequacy)
+
+The second source was IPUMS, which provided us with median poverty tax data for every county from 2018-2022.
+
+The third source was the U.S.
+Census Bureau’s Regions and Divisions of the United States, which maps every U.S.
+to one of our geographic regions, and also a division within the regions.
+We specifically included regions as a variable in this analysis. This data was loaded into R using a publicly available GitHub repository maintained by Halpert (2014), titled Census Regions.
+This repository provides the necessary mappings in CSV format, allowing for efficient integration of regional classification data.
 
@@
 
+2024 datasets for all U.S.
+states were combined and merged with median property tax and region mapping data, resulting in a final dataset with 3144 county observations and 9 different variables (state, average life expectancy, residential segregation index, % limited access to healthy foods, % uninsured adults, school funding adequacy, median household income, median property tax, and region).
+Because school funding adequacy was given as a numerical value, and we wanted to focus simply on better understanding the dichotomy between school funding adequacy and inadequacy, it was converted into a binary categorical variable (1 if the numerical value was positive, indicating the counties’ schools were adequately funded, and 0 if the numerical value was negative, indicating underfunding).
+
+Exploratory data analysis for the two outcome variables was conducted by creating  .
+During exploratory data analysis, we found that ⅓ of our observations were missing residential segregation index data, and in addition, certain states were missing nearly all of their segregation values.
+Because of this, we decided to drop all states missing over 50% of their values for any variable.
+Going forward, our analysis is limited to 33 states, and excludes the following: Alaska, Colorado, Idaho, Iowa, Kansas, Minnesota, Montana, Nebraska, North Dakota, South Dakota, Utah, Wyoming, and Vermont.
+We were left with 2379 total observations, and the missing percentage for residential segregation dropped to 22%.
+
+After this, remaining missing values in all variables were imputed using the mice package and predictive mean matching to generate one complete dataset.
+The original school funding adequacy numerical variable was first imputed, and then converted into binary values for subsequent analysis.
+
 ### Model Fitting & Assessment
+
+For our first research question, a multiple linear regression model was fit, regressing average life expectancy on residential segregation index, % limited access to healthy foods, % uninsured adults, and region.
+Linear regression assumptions were assessed using diagnostic plots, especially residual vs.
+fitted and quantile-quantile plots.
+The adjusted R-squared value was used to evaluate the fit of the model, and the effect of region as an interaction with residential segregation was evaluated using nested F tests.
+We also calculated VIF to understand multicollinearity and Cook’s distance to identify influential points.
+
+For our second research question, a binary logistic regression model was fit with school funding adequacy (categorical) as the outcome and residential segregation index, median property tax, and median household income as predictors.
+The model was assessed using a confusion matrix.
+ROC curve, and deviance.
+
+## Results
+
+### Data Overview
+
+### Question 1
+
+### Question 2
 
@@
 ## Imputation Code 
-# In this block, we are looking to use predictive mean matching to combing regression methods nad hot deck imputing to predicted value closest to the predicted value of the observation with missing data we also needed to convert school funding adequacy imputed values to school funding categorical variable before using in models
+# In this block, we are looking to use predictive mean matching to combing regression methods and hot deck imputing to predicted value closest to the predicted value of the observation with missing data; we also needed to convert school funding adequacy imputed values to school funding categorical variable before using in models
 
@@
 ```{r, echo= FALSE, warning = FALSE}
-# Since we should not imputate on the outcome, but need to imputate the School.Funding.Adequacy variable, we are feeding the School.Funding.Adequacy variable into a ifelse conditional that will replaced the impuated outcome variable.
+# Since we should not imputate on the outcome, but need to imputate the School.Funding.Adequacy variable, we are feeding the School.Funding.Adequacy variable into a ifelse conditional that will replaced the imputed outcome variable.
 
@@
 ```{r, echo= FALSE, warning = FALSE, include = FALSE }
-# To have a full dataset, we are taking the first imputated dataset R provided and using that as our full dataset sample. 
+# To have a full dataset, we are taking the first imputed dataset R provided and using that as our full dataset sample. 
 completed_data <- complete(sub.imp, 1)
@@
 
-### Model Fitting
+### Model Fitting (Original Data)
 
@@
 
-### Regression Results
+### Model Fitting (Imputed Data)
 
@@
 
-modelsummary(mlr.imp.mod, fmt = fmt_significant(2),
+modelsummary(pool(mlr.imp.mod), fmt = fmt_significant(2),
              coef_rename = renamed, 
@@
 
-modelsummary(mlr.imp.mod2, fmt = fmt_significant(2),
+modelsummary(pool(mlr.imp.mod2), fmt = fmt_significant(2),
              coef_rename = renamed, 

### 2024-12-09 code.qmd
adding anova for q2 deviance

@@
 
-lmod <- glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax,
+lmod <- glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax + Region_fac,
               data=combined,
@@
 ```{r}
-imp_mods <- glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax, data=completed_data, family="binomial")
+imp_mods <- glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax + Region_fac, data=completed_data, family="binomial")
 
@@
 ```{r}
+imp_mods_red <- glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax, data=completed_data, family="binomial")
+
+anova(imp_mods_red, imp_mods, test="Chisq")
 

### 2024-12-09 code.qmd
changing summary to modelsummary for q2

@@
 #log-odds scale
-summary(lmod,
+modelsummary(lmod,
              fmt = fmt_significant(2),
@@
 
-summary(lmod,
+modelsummary(lmod,
              fmt = fmt_significant(2),
@@
 #log-odds scale
-summary(imp_mods,
+modelsummary(imp_mods,
              fmt = fmt_significant(2),
@@
 
-summary(imp_mods,
+modelsummary(imp_mods,
              fmt = fmt_significant(2),

### 2024-12-09 code.qmd
draft

@@
 plot(cooksd, main = "Cook's Distance", ylab = "Cook's Distance")
-abline(h = 4/(nrow(data) - length(coef(mlr.imp))), col = "red")  # Threshold line
+abline(h = 4/(nrow(completed_data) - length(coef(mlr.imp))), col = "red")  # Threshold line
+4/(nrow(data) - length(coef(mlr.imp)))
 
@@
 
-```{r include = FALSE}
+```{r, include = FALSE}
 # Identify Influential Points
@@
 
-lmod <- glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax + Region_fac,
+lmod <- glm(factor(School.Funding.Cat) ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax + Region_fac,
               data=combined,

### 2024-12-09 paper.qmd
draft

@@
 institution: "Duke University, IDS 702 Data Modeling and Presentation"
-format: html
+format: pdf
 editor: visual
@@
 ---
-
-## **Abstract**
-
-## **Introduction**
-
-### Background
-
-Life expectancy and school funding are important indicators of public well-being because they capture fundamental aspects of health, education, and social equity that are essential for the development of individuals and communities.
-Life expectancy is an important health indicator that reflects not only the overall health of a society, but also its stability, economic resilience, and levels of inequality.
-For example, differences in life expectancy between regions and socioeconomic groups often reveal systemic inequalities, with marginalized communities experiencing shorter life spans due to limited access to health care, healthy living environments, and economic opportunities.
-Similarly, adequately funded schools mean that students have equal access to learning opportunities and are an important investment in social mobility.
-Underfunded schools are often located in economically disadvantaged or segregated areas, exacerbating educational gaps and limiting students’ potential to break the cycle of poverty.
-These gaps are rooted in systemic inequalities, such as residential segregation, which concentrates poverty, limits access to quality education and health care, and perpetuates income and health disparities across generations.
-By addressing these inequalities, policymakers can foster fairer, more prosperous communities and ensure that everyone benefits from improved health and education systems.
-
-### Research Questions
-
-For this purpose, we focus on two key research questions to explore these relationships.  First, we investigate how socioeconomic factors—residential segregation, access to healthy foods, and health insurance coverage—impact average life expectancy.
-This analysis includes a comparison across different regions to identify variations in the relationship between residential segregation and life expectancy.
-Second, we evaluate how residential segregation, property taxes, and median household income influence school funding adequacy.
-This question allows us to assess educational investment disparities among regions with varying economic conditions.
-The observational unit for this analysis was the U.S.
-county.
-
-### Data Overview
-
-The dataset used in this study integrates information from three primary sources.
-
-The first source was the County Health Rankings & Roadmaps program, associated with the University of Wisconsin Population Health Institute, which provided us with 2024 datasets for all U.S.
-states.
-This program consolidates the latest county-level measurements of various population health, economic, demographic, and social factors from the American Community Survey, the USDA Food Environment Atlas, Small Area Income and Poverty Estimates, and other reliable government sources into publicly available datasets every year.
-Our specific variables of interest were originally collected as part of:
-
--   National Center for Health Statistics - Natality & Mortality Files, 2019-2021 (average life expectancy)
-
--   American Community Survey 5-year estimates, 2018-2022 (residential segregation index & median household income)
-
--   USDA Food Environment Atlas, 2019 (percentage of population with limited access to healthy foods)
-
--   Small Area Health Insurance Estimates, 2021 (percentage of adults uninsured)
-
--   School Finance Indicators Database, 2021 (school funding adequacy)
-
-The second source was IPUMS, which provided us with median poverty tax data for every county from 2018-2022.
-
-The third source was the U.S.
-Census Bureau’s Regions and Divisions of the United States, which maps every U.S.
-to one of our geographic regions, and also a division within the regions.
-We specifically included regions as a variable in this analysis. This data was loaded into R using a publicly available GitHub repository maintained by Halpert (2014), titled Census Regions.
-This repository provides the necessary mappings in CSV format, allowing for efficient integration of regional classification data.
-
-## **Methods**
-
-### Variable Selection
-
-### Exploratory Data Analysis & Data Preprocessing
-
-2024 datasets for all U.S.
-states were combined and merged with median property tax and region mapping data, resulting in a final dataset with 3144 county observations and 9 different variables (state, average life expectancy, residential segregation index, % limited access to healthy foods, % uninsured adults, school funding adequacy, median household income, median property tax, and region).
-Because school funding adequacy was given as a numerical value, and we wanted to focus simply on better understanding the dichotomy between school funding adequacy and inadequacy, it was converted into a binary categorical variable (1 if the numerical value was positive, indicating the counties’ schools were adequately funded, and 0 if the numerical value was negative, indicating underfunding).
-
-Exploratory data analysis for the two outcome variables was conducted by creating  .
-During exploratory data analysis, we found that ⅓ of our observations were missing residential segregation index data, and in addition, certain states were missing nearly all of their segregation values.
-Because of this, we decided to drop all states missing over 50% of their values for any variable.
-Going forward, our analysis is limited to 33 states, and excludes the following: Alaska, Colorado, Idaho, Iowa, Kansas, Minnesota, Montana, Nebraska, North Dakota, South Dakota, Utah, Wyoming, and Vermont.
-We were left with 2379 total observations, and the missing percentage for residential segregation dropped to 22%.
-
-After this, remaining missing values in all variables were imputed using the mice package and predictive mean matching to generate one complete dataset.
-The original school funding adequacy numerical variable was first imputed, and then converted into binary values for subsequent analysis.
-
-### Model Fitting & Assessment
-
-For our first research question, a multiple linear regression model was fit, regressing average life expectancy on residential segregation index, % limited access to healthy foods, % uninsured adults, and region.
-Linear regression assumptions were assessed using diagnostic plots, especially residual vs.
-fitted and quantile-quantile plots.
-The adjusted R-squared value was used to evaluate the fit of the model, and the effect of region as an interaction with residential segregation was evaluated using nested F tests.
-We also calculated VIF to understand multicollinearity and Cook’s distance to identify influential points.
-
-For our second research question, a binary logistic regression model was fit with school funding adequacy (categorical) as the outcome and residential segregation index, median property tax, and median household income as predictors.
-The model was assessed using a confusion matrix.
-ROC curve, and deviance.
-
-## Results
-
-### Data Overview
-
-### Question 1
-
-### Question 2
 
@@
 combined <- read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/combined.csv')
+
 ```
@@
 combined <- combined |> mutate(State_fac = factor(combined$State), Region_fac = factor(combined$Region, levels=c('South', 'West', 'Northeast', 'Midwest')))
+
 ```
@@
 # Perform the imputation
-sub.imp <- mice(sub, m = 1, method = "pmm", print = FALSE)
+sub.imp <- mice(sub, m = 1, method = "pmm", print = FALSE, seed=3)
 
@@
 }
+
 ```
@@
 colSums(is.na(completed_data))
+```
+
+## **Abstract**
+
+This study examines the relationships between socioeconomic, health, and education indicators in U.S.
+counties using data from the American Community Survey, the USDA Food Environment Atlas, small area income and poverty estimates, and other sources.
+We will explore regional influences and predictors of life expectancy, and also identify socioeconomic factors associated with school funding adequacy.
+By analyzing differences in life expectancy and education funding across socioeconomic and regional contexts, this study aims to gain insight into public well-being and economic inequality across the U.S.
+Through exploratory data analysis and regression modeling with interaction terms, we identify significant regional disparities in both life expectancy and school funding adequacy.
+The findings highlight that regions, particularly the West and Northeast, tend to have better health and education outcomes, emphasizing the critical role of location and economic conditions in shaping societal well-being.
+
+## **Introduction**
+
+### Background
+
+Life expectancy and school funding are important indicators of public well-being because they capture fundamental aspects of health, education, and social equity that are essential for the development of individuals and communities.
+Life expectancy is an important health indicator that reflects not only the overall health of a society, but also its stability, economic resilience, and levels of inequality.
+For example, differences in life expectancy between regions and socioeconomic groups often reveal systemic inequalities, with marginalized communities experiencing shorter life spans due to limited access to health care, healthy living environments, and economic opportunities.
+Similarly, adequately funded schools mean that students have equal access to learning opportunities and are an important investment in social mobility.
+Underfunded schools are often located in economically disadvantaged or segregated areas, exacerbating educational gaps and limiting students’ potential to break the cycle of poverty.
+Because both of these outcomes are impacted by a complicated web of interactions between socioeconomic, political, and cultural factors, it is vital that we work towards better understanding the most influential out of these, so that policy and community intervention designed to improve life expectancy and school funding can take on a multifaceted, and thus a more effective approach.
+
+### Research Questions
+
+For this purpose, we focused our analysis on two key research questions.  First, we investigated how socioeconomic factors—residential segregation, access to healthy foods, and health insurance coverage—impact average life expectancy.
+This analysis included a comparison across different regions to identify variations in the relationship between residential segregation and life expectancy.
+Second, we evaluated how residential segregation, property taxes, and median household income influence school funding adequacy.
+This question allowed us to assess educational investment disparities among regions with varying economic conditions.
+The observational unit for this analysis was the U.S.
+county.
+
+### Data Overview
+
+The dataset used in this study combined information from three primary sources.
+
+The first source was the County Health Rankings & Roadmaps program, associated with the University of Wisconsin Population Health Institute, which provided us with 2024 datasets for all U.S.
+states.
+This program consolidates the latest county-level measurements of various population health, economic, demographic, and social factors from the American Community Survey, the USDA Food Environment Atlas, Small Area Income and Poverty Estimates, and other reliable government sources into publicly available datasets every year.
+Our specific variables of interest were originally collected as part of:
+
+-   National Center for Health Statistics - Natality & Mortality Files, 2019-2021 (average life expectancy)
+
+-   American Community Survey 5-year estimates, 2018-2022 (residential segregation index & median household income)
+
+-   USDA Food Environment Atlas, 2019 (percentage of population with limited access to healthy foods)
+
+-   Small Area Health Insurance Estimates, 2021 (percentage of adults uninsured)
+
+-   School Finance Indicators Database, 2021 (school funding adequacy)
+
+The second source was IPUMS, which provided us with median poverty tax data for every county from 2018-2022.
+
+The third source was the U.S.
+Census Bureau’s Regions and Divisions of the United States, which maps every U.S.
+to one of four geographic regions, and also a division within the regions.
+We specifically included regions as a variable in this analysis. This data was loaded into R using a [publicly available GitHub repository](https://github.com/cphalpert/census-regions) titled *Census Regions*.
+
+## **Methods**
+
+### Variable Selection
+
+Predictor variables were selected *a priori* based on their perceived relevance to the two outcomes we wanted to better understand: average life expectancy and school funding adequacy.
+Nutrition and access to quality healthcare are inherently linked to overall well-being.
+To analyze the extent to which each contributes to improving life expectancy, we included two key predictors: the percentage of the population with limited access to healthy foods and the percentage of uninsured adults in a county.
+The former is defined by the US Department of Agriculture as "the percentage of the \[county\] population that is low-income and does not live close to a grocery store", where "close" is defined differently for rural and nonrural areas.
+The latter variable represents the percentage of adults in a county younger than 65 who lack health insurance.
+We also were interested in seeing how residential segregation would impact life expectancy among these predictors, given existing studies on the negative health effects of segregation, and whether this relationship would exhibit any variations across different regions.
+This is why we chose to include residential segregation index values.
+The residential segregation index, used by the County Health Rankings & Roadmaps program, measures how evenly two groups, such as Black and white residents, are distributed across geographic areas, with values ranging from 0 (complete integration) to 100 (complete segregation).
+
+School funding adequacy is defined by the County Health Rankings program as "the average dollar gap between actual per-pupil spending and the amount needed for students to achieve national average test scores, considering districts' varying equity-based needs." Because a major source of public school funding is tax revenue, especially property tax, we chose to include this variable as a predictor, along with an indicator of the general economic conditions of a county—median household income.
+In addition, to explore how race and racial diversity interface with variations in school funding, we selected residential segregation to be a part of this regression as well.
+
+### Exploratory Data Analysis & Data Preprocessing
+
+2024 datasets for all U.S.
+states were combined and merged with median property tax and region mapping data, resulting in a final dataset with 3144 county observations and 9 different variables (state, average life expectancy, residential segregation index, % limited access to healthy foods, % uninsured adults, school funding adequacy, median household income, median property tax, and region).
+Because school funding adequacy was given as a numerical value, and we wanted to focus on better understanding the dichotomy between school funding adequacy and inadequacy, this variable was converted into a binary categorical variable (where the value was set to 1 if the numerical value was positive, indicating the counties’ schools were adequately funded, and set to 0 if the numerical value was negative, indicating underfunding).
+
+Exploratory data analysis for the two outcome variables was conducted by 1) creating scatterplots of each individual predictor against the outcome (for life expectancy) and 2) creating pairwise plots between variables (for school funding).
+During the initial stages of analysis, we found that ⅓ of our observations were missing residential segregation index data, and in addition, certain states were missing nearly all of their segregation values.
+Because of this, we decided to drop all states missing over 50% of their values for any variable.
+This led to our analysis being limited to 33 states, and excluding the following: Alaska, Colorado, Idaho, Iowa, Kansas, Minnesota, Montana, Nebraska, North Dakota, South Dakota, Utah, Wyoming, and Vermont.
+We were left with 2379 total observations, and the missing percentage for residential segregation dropped to 22%.
+
+After this, remaining missing values in all variables were imputed using the mice package and predictive mean matching to generate one complete dataset.
+The original school funding adequacy numerical variable was first imputed, and then converted into binary values for subsequent analysis.
+
+### Model Fitting & Assessment
+
+For our first research question, a multiple linear regression model was fit, regressing average life expectancy on residential segregation index, percentage with limited access to healthy foods, percentage uninsured adults, and region.
+Linear regression assumptions were assessed using diagnostic plots, especially residual vs.
+fitted and quantile-quantile plots.
+The adjusted R-squared value was used to evaluate the fit of the model, and the effect of region as an interaction with residential segregation was evaluated using nested F tests.
+We also calculated VIF to check for multicollinearity and Cook’s distance to identify influential points.
+
+For our second research question, a binary logistic regression model was fit with school funding adequacy (categorical) as the outcome and residential segregation index, median property tax, region, and median household income as predictors.
+The model was assessed using a confusion matrix, an ROC curve, and a comparison of deviance between the full model and a reduced model excluding region as a predictor.
+
+## Results
+
+### Data Overview
+
+The below tables display summary statistics on average life expectancy (1) and school funding adequacy (2).
+Both the median and mean life expectancy were 75 years.
+School funding adequacy counts show that around 66% of the counties in our analysis had under funded schools.
+
+|                 | **Average Life Expectancy (years)** |
+|-----------------|-------------------------------------|
+| Min             | 65                                  |
+| 25th percentile | 73                                  |
+| Median          | 75                                  |
+| 75th percentile | 77                                  |
+| Max             | 99                                  |
+| Mean            | 75                                  |
+
+|                      |                           |
+|----------------------|---------------------------|
+| **0 (Under Funded)** | **1 (Adequately Funded)** |
+| 1539                 | 806                       |
+
+Additionally, the number of counties per region in this analysis indicate an imbalance between categories, with the South region being over represented (holding 60% of all counties), and the East and Northeast holding around 9% of the counties each.
+
+|             |               |           |          |
+|-------------|---------------|-----------|----------|
+| **Midwest** | **Northeast** | **South** | **East** |
+| 552         | 203           | 1421      | 203      |
+
+As mentioned earlier, EDA for the first research question involved generating scatterplots plotting each numerical predictor plotted against life expectancy, as shown below.
+The residential segregation index vs.
+life expectancy plot was also faceted by region to explore whether each region demonstrated a different relationship in these two variables.
+Because the slopes were largely similar, this suggested that the interaction effect might not be that strong.
+In general, all three scatterplots demonstrated considerable spread and did not show a very substantial linear relationship between the numerical predictors and average life expectancy.
+
+```{r, echo=FALSE}
+library(ggplot2)
+library(GGally)
+library(patchwork)
+plot1 <- ggplot(completed_data, aes(x = Residential.Segregation.Index, y = Life.Expectancy, color=Region_fac)) +
+  geom_point(color='black') + geom_smooth(method="lm",se=F) + 
+  scale_color_brewer(palette = "Dark2") +
+  labs(x = "Residential Segregation Index", y = "Life Expectancy (yrs)", color='Region')+ 
+  theme(plot.margin = margin(t = 20, r = 30, b = 20, l = 10),
+axis.title.x = element_text(size = 8),
+axis.title.y = element_text(size = 8))
+
+plot2 <- ggplot(completed_data, aes(x = X..Limited.Access.to.Healthy.Foods
+, y = Life.Expectancy)) +
+  geom_point(color = "black") + geom_smooth(method="lm",se=F)+
+  labs( x = "% Population With Limited Healthy Food Access",y = "Life Expectancy (yrs)") +
+  theme(plot.margin = margin(t = 20, r = 20, b = 20, l = 20),
+        axis.title.x = element_text(size = 8),
+        axis.title.y = element_text(size = 8))
+
+plot3 <- ggplot(completed_data, aes(x = X..Uninsured.Adults, y = Life.Expectancy)) +
+  geom_point(color = "black") +geom_smooth(method="lm",se=F)+
+  labs(x = "% Adult Population Without Health Insurance", y = "Life Expectancy (yrs)")+ 
+  theme(plot.margin = margin(t = 20, r = 20, b = 20, l = 20),
+        axis.title.x = element_text(size = 8),
+        axis.title.y = element_text(size = 8))
+
+plot1 + plot2 + plot3 + plot_layout(ncol = 2, nrow=2) + 
+  plot_annotation(title = "Relationships Between Predictors and 
+                  Life Expectancy Outcome",
+                  theme = theme(
+      plot.title = element_text(hjust=0.5)))
+```
+
+For the second research question, pairwise plots were generated to show the relationship between predictors (as seen by the scatterplots below), and also to show the relationship between each numerical predictor and the school funding outcome variable (as seen by the density plots below).
+Points and plots colored in red represent counties where school funding is inadequate (=0), and points and plots colored in blue represent counties where funding is adequate (=1).
+
+```{r, echo=FALSE}
+ggpairs(completed_data, columns = c("Residential.Segregation.Index", 
+                              "Median.Household.Income",
+                              "Median.Prop.Tax"), 
+        aes(color = factor(School.Funding.Cat))) + 
+        labs(title = "Pairwise Plots of Predictors Colored by 
+             School Funding Adequacy",
+              x = "Predictor 1",
+              y = "Predictor 2",
+              color = "School Funding Adequacy") +
+         theme(plot.title = element_text(size=15, hjust=0.5), 
+               strip.text.x = element_text(size = 5),
+               strip.text.y = element_text(size = 3))
 
@@
 
-## Q1 Life Expectancy; MLR
+### Question 1
 
-### **Research Question 1: Life Expectancy**
+#### Model Results
 
-### Model Fitting (Original Data)
+Based on the multiple linear regression results, percentage of uninsured adults, percentage of population with limited access to healthy foods, and region all emerged with statistically significant coefficient values.
+The interaction term between region and residential segregation, as well as the individual term of residential segregation, were not statistically significant.
+While percentage of uninsured adults and percentage with limited access to healthy foods were statistically significant, the coefficients themselves indicate very minor impacts on life expectancy (a 1 unit increase in either of these percentages leads to average life expectancy decreasing by around 0.05 years).
+The impact of region on life expectancy, however, was notable: life expectancy is on average 4.17 years higher in the West region compared to the South, 6.7 years higher in the Northeast compared to the South, and 2.41 years higher in the Midwest compared to the South.
 
-```{r, echo= FALSE, warning = FALSE}
+```{r, echo= FALSE, warning = FALSE, include=FALSE}
 # MLR
@@
 renamed <- c(
-  "Residential.Segregation.Index" = "Residential Segregation Index (RSI)",
-  "X..Limited.Access.to.Healthy.Foods" = "Access to Healthy Foods",
-  "X..Uninsured.Adults" = "Pct. of Uninsured Adults",
-  "Region_facWest" = "Region.W", 
-  "Region_facNortheast" = "Region.NE",
-  "Region_facMidwest" = "Region.MidW")
+   "Residential Segregation Index (RSI)" = "Residential.Segregation.Index",
+  "Percentage With Limited Access to Healthy Foods" = "X..Limited.Access.to.Healthy.Foods",
+  "Percentage of Uninsured Adults"= "X..Uninsured.Adults",
+  "Region-West" = "Region_facWest", 
+  "Region-Northeast"= "Region_facNortheast",
+  "Region-Midwest"= "Region_facMidwest",
+  "RSI*Region-West" = "Residential.Segregation.Index:Region_facWest",
+  "RSI*Region-Northeast" = "Residential.Segregation.Index:Region_facNortheast",
+  "RSI*Region-Midwest" = "Residential.Segregation.Index:Region_facMidwest")
 
@@
 
-### Model Fitting (Imputed Data)
+```{r, echo= FALSE, warning = FALSE}
+# Running a regression with the imputated data (completed_data) to understand the impact of missing data
 
-```{r, echo= FALSE, warning = FALSE}
-# Running a regression with the imputated data to understand the impact of missing data
-
-mlr.imp.mod <- with(sub.imp, lm(Life.Expectancy ~ Residential.Segregation.Index +
+mlr.imp <- lm(Life.Expectancy ~ Residential.Segregation.Index * Region_fac +
     X..Limited.Access.to.Healthy.Foods +
-    X..Uninsured.Adults + 
-    Region_fac))
+    X..Uninsured.Adults, data=completed_data)
 
@@
     "Residential.Segregation.Index" = "Residential Segregation Index (RSI)",
-  "X..Limited.Access.to.Healthy.Foods" = "Access to Healthy Foods",
-  "X..Uninsured.Adults" = "Pct. of Uninsured Adults",
-  "Region_facWest" = "Region.W", 
-  "Region_facNortheast" = "Region.NE",
-  "Region_facMidwest" = "Region.MidW"
-  )
+  "X..Limited.Access.to.Healthy.Foods" = "Percentage With Access to Healthy Foods",
+  "X..Uninsured.Adults"= "Percentage of Uninsured Adults",
+  "Region_facWest"="Region-West", 
+  "Region_facNortheast"="Region-Northeast",
+  "Region_facMidwest"="Region-Midwest" ,
+  "Residential.Segregation.Index:Region_facWest" = "RSI*Region-West",
+  "Residential.Segregation.Index:Region_facNortheast"= "RSI*Region-Northeast",
+  "Residential.Segregation.Index:Region_facMidwest"="RSI*Region-Midwest")
 
-modelsummary(pool(mlr.imp.mod), fmt = fmt_significant(2),
+modelsummary(mlr.imp, fmt = fmt_significant(2),
              coef_rename = renamed, 
@@
 
-```{r echo= FALSE, warning = FALSE}
-# Running a regression with the interaction term between residential segregation on the regions
+```{r echo= FALSE, warning = FALSE, include=FALSE}
+# Running a regression without the interaction term between residential segregation on the regions
 
-mlr.imp.mod2 <- with(sub.imp, lm(Life.Expectancy ~ Residential.Segregation.Index +
+mlr.imp.reduced <- lm(Life.Expectancy ~ Residential.Segregation.Index +
     X..Limited.Access.to.Healthy.Foods +
     X..Uninsured.Adults + 
-    Region_fac + 
-    Region_fac * Residential.Segregation.Index))
+    Region_fac, data=completed_data)
 
 renamed <- c(
-  "Residential.Segregation.Index" = "Residential Segregation Index (RSI)",
-  "X..Limited.Access.to.Healthy.Foods" = "Access to Healthy Foods",
-  "X..Uninsured.Adults" = "Pct. of Uninsured Adults",
-  "Region_facWest" = "Region.W", 
-  "Region_facNortheast" = "Region.NE",
-  "Region_facMidwest" = "Region.MidW", 
-  "Residential.Segregation.Index:Region_facWest" = "RSI * Region.W ", 
-  "Residential.Segregation.Index:Region_facNortheast" = "RSI * Region.NE",
-  "Residential.Segregation.Index:Region_facMidwest" = "RSI * Region.MidW"
-  )
+   "Residential Segregation Index (RSI)" = "Residential.Segregation.Index",
+  "Percentage With Limited Access to Healthy Foods" = "X..Limited.Access.to.Healthy.Foods",
+  "Percentage of Uninsured Adults"= "X..Uninsured.Adults",
+  "Region-West" = "Region_facWest", 
+  "Region-Northeast"= "Region_facNortheast",
+  "Region-Midwest"= "Region_facMidwest")
 
-modelsummary(pool(mlr.imp.mod2), fmt = fmt_significant(2),
+modelsummary(mlr.imp.reduced, fmt = fmt_significant(2),
              coef_rename = renamed, 
@@
 
-### Evaluation & Assessment of Model
+#### Model Assessment
 
-```{r}
+The adjusted R-squared value for the model was around 0.2, indicating that only a small amount of the variability in life expectancy was explained by the predictors in the model.
+This points to the fact that are a plethora of other socioeconomic and environmental factors that have an effect on life expectancy, and expanding our model to include more of these would have likely yielded a better fit.
+
+In conducting diagnostics on the multilinear regression model, we first checked to see if the model meets our assumptions of linearity, independence, normality, and homoscedasticity.
+The residuals vs.
+fitted plot showed relatively equal variance and no discernible pattern, meaning that the linearity and homoscedasticity assumptions were reasonably met.
+In addition, because each observation in the dataset represented a unique county in the U.S., independence can be reasonably assumed.
+Assessing the Q-Q Plot for normality, the residual distribution was approximately normal, despite a slight divergence on both tails of the plot.
+Finally, through observing the Residuals vs.
+Leverage plot (not pictured), we observed only a few points falling near or beyond the Cook’s distance threshold, suggesting that there may be a few influential points that can be seen as significantly impacting the results of the regression.
+These points could affect the validity of the model.
+
+With variance inflation factors of roughly less than 2 for all predictors, there was little to no multicollinearity amongst the predictors.
+Through a nested-f test comparing a full model with a reduced model that excludes the interaction term between region and residential segregation, we concluded that a full model fit the data substantially better.
+The reduction in RSS in the full model compared to the reduced model was highly significant, indicating that the interaction between region and residential segregation added meaningful explanatory power to the model.
+
+```{r, echo=FALSE, include=FALSE}
 #QQPlot
-qqnorm(residuals(mlrmodel))
-qqline(residuals(mlrmodel), col = "red") 
+qqnorm(residuals(mlr.imp))
+qqline(residuals(mlr.imp), col = "red") 
 ```
 
-```{r}
-##Nested F-Test
-#Assuming we will drop %Uninsured Adults Predictor
-
-reduced_model <- lm(Life.Expectancy ~ Residential.Segregation.Index + X..Limited.Access.to.Healthy.Foods, data = combined)
-
-# Nested F-Test
-anova(reduced_model, mlrmodel)
+```{r, echo=FALSE, include=FALSE}
+##Nested F-Test to see if interaction term makes a difference
+anova(mlr.imp, mlr.imp.reduced)
 
@@
 
-```{r}
+```{r, include=FALSE, echo=FALSE}
 #Cooks Distance 
-cooksd <- cooks.distance(mlrmodel)
+cooksd <- cooks.distance(mlr.imp)
 
@@
 
-```{r include = FALSE}
+```{r, echo=FALSE, include = FALSE}
 # Identify Influential Points
-influential <- which(cooksd > 4/(nrow(combined) - length(coef(mlrmodel))))
+influential <- which(cooksd > 4/(nrow(combined) - length(coef(mlr.imp))))
 print(influential)
@@
 
-```{r}
+```{r, echo=FALSE, include=FALSE}
 #Variance Inflaction Factor (VIF)
 
-vif(mlrmodel)
+vif(mlr.imp)
+vif(mlr.imp.reduced)
 ```
 
-Results are all slightly greater than one meaning some correlation between factors but not much. Percentage of uninsured adults having the greatest correlation
-
-```{r}
+```{r, echo=FALSE}
 #Generate 4 Diagnostic Plots at once
 par(mfrow = c(2, 2), mar = c(4,4,2,2))
-plot(mlrmodel)
+plot(mlr.imp)
 ```
 
-## Q2 **School Funding Adequacy; GLM**
+## Question 2
 
-### **Research Question 2: School Funding Adequacy**
+Pivoting to the second research question, we sought to assess the impact residential segregation, median property taxes, the median household income, and the region of which we are assessing has on whether or not a school is adequately funded by using a logistic regression model.
 
-### Model Fitting
+#### Model Results
 
-```{r}
+In holding all other variables constant and using the South region as the baseline category for region, the table below displays the odds ratio results from the logistic regression model.
+
+While all estimates are statistically significant at the 95% confidence level, exponentiated regression coefficients from the logistic model demonstrate that the level of residential segregation, the median household income, and median property tax of the county have very minimal impacts on the odds of a school being adequately funded.
+
+Region, however, has a substantial impact on the odds of school funding adequacy.  Being located in the Northeast or the Midwest has the greatest impact on how well the school is funded, with the odds of adequate school funding being over 200 times higher in a Northeast county compared to a Southern county, and the odds of adequate school funding being over 3 times higher in a Northeast county compared to a Southern county, holding all other variables constant.
+Such disparities across the country are caused by a number of factors, including: (1) capacity - how well off a state is based on its economy and resources, and (2) effort - the states willingness to provide funding for education.
+Wealthier states with a high fiscal capacity, (typically those in the Northeast), have more funding available to spend on education than states with more limited resources (typically those in the South and the West).
+
+```{r, echo=FALSE, include=FALSE}
 ## Checking to see if the imputated data has a large effect on the outcome
@@
 
-lmod <- glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax,
+lmod <- glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax + Region_fac,
               data=combined,
               family="binomial")
-
-library("modelsummary")
 
@@
 
-### Regression Results
-
-```{r}
-imp_mods <- with(sub.imp, glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax), family="binomial")
+```{r, include=FALSE, echo=FALSE}
+imp_mods <- glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax + Region_fac, data=completed_data, family="binomial")
 
 #log-odds scale
-modelsummary(pool(imp_mods),
+modelsummary(imp_mods,
              fmt = fmt_significant(2),
@@
              gof_map=NA)
+```
 
-modelsummary(pool(imp_mods),
+```{r, echo=FALSE}
+
+renamed <- c(
+    "Residential.Segregation.Index" = "Residential Segregation Index (RSI)",
+    "Median.Household.Income"="Median Household Income, USD",
+    "Median.Prop.Tax" = "Median Property Tax, USD",
+  "Region_facWest"="Region-West", 
+  "Region_facNortheast"="Region-Northeast",
+  "Region_facMidwest"="Region-Midwest"
+)
+modelsummary(imp_mods,
              fmt = fmt_significant(2),
              shape = term ~ model + statistic,
+             coef_rename = renamed,
              statistic = c("std.error","conf.int","p.value"),
@@
 
-### Evaluation and Assessment of Model
+#### Model Assessment
 
-```{r}
+Evaluating our model, we found that the logistic regression model demonstrated a strong overall performance with an accuracy rate of 80.33%.
+The model had a high sensitivity of 91%, indicating that it accurately predicted 91% of actual positive cases of inadequate school funding when evaluated on our dataset.
+However, it had a 60% specificity rate, highlighting it correctly identified cases of adequate school funding only 60% of the time.
+Moreover, looking at the ROC curve below, the AUC is 0.858, indicating a moderately strong discriminate ability.
+
+In comparing our full model (with predictor variables for residential segregation, median household income, median property tax, and region) to a model with including all of these variables except for region, we found that there is statistical significance in the deviance between the residuals of the full model and reduced.
+As such, we can conclude that including region as a predictor refines and creates a better fit for our model.
+
+```{r, echo=FALSE, include=FALSE}
 # Creating the confusion matrix
@@
 
-```{r}
+```{r, echo=FALSE, include=FALSE}
 # Create a confusion matrix
@@
 
-```{r}
+```{r, echo=FALSE}
 #Creating the AUC curves
@@
 
-```{r}
+```{r, include=FALSE, echo=FALSE}
+imp_mods_red <- glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax, data=completed_data, family="binomial")
 
+anova(imp_mods_red, imp_mods, test="Chisq")
 ```
+
+## **Conclusion**
+
+By analyzing these two research questions, we observed the significant impact of region on both life expectancy (a health indicator) and school funding (an education indicator), both of which are key components of societal well-being.
+These indices tend to be higher in the West and Northeast, which are more developed regions.
+The findings provide empirical evidence supporting the significant disparities in life expectancy across regions with differing economic conditions.
+They also provide insights into social policy investment decisions, highlighting the importance of fair resource allocation to people's well-being.
+
+Our analysis faced limitations due to missing data, causing our dataset to be restricted to 33 states.
+Additionally, because The County Health Rankings program consolidates the latest from a variety of governmental sources, most of our variables were pulled from different years between 2018 and 2022; we assumed that trends remained largely consistent between these years, treating time as insignificant, but this may not actually be the case.
+
+Future studies could incorporate more social variables to enhance the analysis and investigate the causal relationships between region and outcomes.
+Expanding the dataset to include more states and addressing missing data would also provide a more comprehensive understanding.

### 2024-12-14 paper.qmd
fixing draft formatting
updating paper qmd with new q1 model and variable selection changes

@@
 
-### Background
-
 Life expectancy and school funding are important indicators of public well-being because they capture fundamental aspects of health, education, and social equity that are essential for the development of individuals and communities.
 Life expectancy is an important health indicator that reflects not only the overall health of a society, but also its stability, economic resilience, and levels of inequality.
-For example, differences in life expectancy between regions and socioeconomic groups often reveal systemic inequalities, with marginalized communities experiencing shorter life spans due to limited access to health care, healthy living environments, and economic opportunities.
-Similarly, adequately funded schools mean that students have equal access to learning opportunities and are an important investment in social mobility.
-Underfunded schools are often located in economically disadvantaged or segregated areas, exacerbating educational gaps and limiting students’ potential to break the cycle of poverty.
-Because both of these outcomes are impacted by a complicated web of interactions between socioeconomic, political, and cultural factors, it is vital that we work towards better understanding the most influential out of these, so that policy and community intervention designed to improve life expectancy and school funding can take on a multifaceted, and thus a more effective approach.
+According to "The Root Causes of Health Inequity - Communities in Action" report from the National Institutes of Health, we can infer that differences in life expectancy between regions and socioeconomic groups often reveal systemic inequalities, with marginalized communities experiencing shorter life spans due to limited access to health care, healthy living environments, and economic opportunities.^1^ Similarly, adequately funded schools ensure that students have equal access to learning opportunities, serving as a crucial investment in social mobility.
+The Stanford Center for Opportunity Policy in Education highlights that "the unequal allocation and inadequate levels of resources in schools and communities \[are\] at the heart of many gaps in student opportunity."^2^ This emphasizes the importance of equitable and adequate school funding in closing opportunity gaps and empowering students from disadvantaged backgrounds to achieve academic and economic success.
+Consequently, it can be inferred that underfunded schools, often situated in economically disadvantaged or segregated areas, exacerbate educational disparities and limit students' ability to break the cycle of poverty.
+These gaps are rooted in systemic inequalities, such as residential segregation, which concentrates poverty, limits access to quality education and healthcare, and perpetuates income and health disparities across generations.
+By addressing these inequalities, policymakers can foster fairer, more prosperous communities and ensure that everyone benefits from improved health and education systems.
 
-### Research Questions
-
-For this purpose, we focused our analysis on two key research questions.  First, we investigated how socioeconomic factors—residential segregation, access to healthy foods, and health insurance coverage—impact average life expectancy.
-This analysis included a comparison across different regions to identify variations in the relationship between residential segregation and life expectancy.
-Second, we evaluated how residential segregation, property taxes, and median household income influence school funding adequacy.
-This question allowed us to assess educational investment disparities among regions with varying economic conditions.
-The observational unit for this analysis was the U.S.
+For this purpose, we focus our analysis on two key research questions.  First, we investigate how socioeconomic factors—residential segregation, access to healthy foods, and health insurance coverage—impact average life expectancy.
+This analysis includes a comparison across different regions to identify variations in the relationship between residential segregation and life expectancy.
+Second, we evaluate how residential segregation and region influence school funding adequacy.
+This question allows us to assess how the racial makeup and geographic location—given that each region of the U.S.
+is characterized by differing political and cultural norms—affect a county's educational investment.
+The observational unit for this analysis is the U.S.
 county.
 
-### Data Overview
+To obtain data on the aforementioned variables, we used three primary sources.
 
-The dataset used in this study combined information from three primary sources.
-
-The first source was the County Health Rankings & Roadmaps program, associated with the University of Wisconsin Population Health Institute, which provided us with 2024 datasets for all U.S.
+The first source was the County Health Rankings & Roadmaps program,^3^ associated with the University of Wisconsin Population Health Institute, which provided us with 2024 datasets for all U.S.
 states.
@@
 
-The second source was IPUMS, which provided us with median poverty tax data for every county from 2018-2022.
+The second source was IPUMS, which provided us with median poverty tax data for every county from 2018-2022.^4^
 
@@
 to one of four geographic regions, and also a division within the regions.
-We specifically included regions as a variable in this analysis. This data was loaded into R using a [publicly available GitHub repository](https://github.com/cphalpert/census-regions) titled *Census Regions*.
+We specifically included regions as a variable in this analysis. This data was loaded into R using a [publicly available GitHub repository](https://github.com/cphalpert/census-regions) titled *Census Regions*.^5^
 
@@
 
-Predictor variables were selected *a priori* based on their perceived relevance to the two outcomes we wanted to better understand: average life expectancy and school funding adequacy.
-Nutrition and access to quality healthcare are inherently linked to overall well-being.
-To analyze the extent to which each contributes to improving life expectancy, we included two key predictors: the percentage of the population with limited access to healthy foods and the percentage of uninsured adults in a county.
-The former is defined by the US Department of Agriculture as "the percentage of the \[county\] population that is low-income and does not live close to a grocery store", where "close" is defined differently for rural and nonrural areas.
-The latter variable represents the percentage of adults in a county younger than 65 who lack health insurance.
+Variables of interest were selected *a priori* based on their perceived relevance to the two outcomes we wanted to better understand: average life expectancy and school funding adequacy.
+While nutrition and access to quality healthcare are inherently linked to general wellbeing, it is worthwhile to assess how influential the environmental and financial aspects of these factors are on our tangible outcome.
+Therefore, we selected two key predictors for our life expectancy model: the percentage of the population with limited access to healthy foods and the percentage of uninsured adults in a county.
+The former is defined by the US Department of Agriculture as "the percentage of the \[county\] population that is low-income and does not live close to a grocery store", where "close" is defined differently for rural and nonrural areas.^6^ The latter variable represents the percentage of adults in a county younger than 65 who lack health insurance.
 We also were interested in seeing how residential segregation would impact life expectancy among these predictors, given existing studies on the negative health effects of segregation, and whether this relationship would exhibit any variations across different regions.
 This is why we chose to include residential segregation index values.
-The residential segregation index, used by the County Health Rankings & Roadmaps program, measures how evenly two groups, such as Black and white residents, are distributed across geographic areas, with values ranging from 0 (complete integration) to 100 (complete segregation).
+The residential segregation index, used by the County Health Rankings & Roadmaps program, measures how evenly two groups, such as Black and white residents, are distributed across geographic areas, with values ranging from 0 (complete integration) to 100 (complete segregation).^7^
 
-School funding adequacy is defined by the County Health Rankings program as "the average dollar gap between actual per-pupil spending and the amount needed for students to achieve national average test scores, considering districts' varying equity-based needs." Because a major source of public school funding is tax revenue, especially property tax, we chose to include this variable as a predictor, along with an indicator of the general economic conditions of a county—median household income.
-In addition, to explore how race and racial diversity interface with variations in school funding, we selected residential segregation to be a part of this regression as well.
+*Income is known to have a strong relationship with.*^8^
+
+School funding adequacy is defined by the County Health Rankings program as "the average dollar gap between actual per-pupil spending and the amount needed for students to achieve national average test scores, considering districts' varying equity-based needs."^9^ To explore how racial diversity interface with variations in school funding, we selected residential segregation to be a part of this regression.
+In addition, given that state policy has a large influence on school administration and the determination of school needs, and that states in the same region tend to have similar political leanings and cultural beliefs, we wanted to analyze how much of a relationship exists between region and actual school funding levels.
+
+Because a major source of public school funding is tax revenue, median property tax and median household income were also included in the school funding model as potential confounders.
+This allowed us to assess whether the effects of segregation and region on funding existed outside of the effects of known economic factors.
 
@@
 
-Exploratory data analysis for the two outcome variables was conducted by 1) creating scatterplots of each individual predictor against the outcome (for life expectancy) and 2) creating pairwise plots between variables (for school funding).
-During the initial stages of analysis, we found that ⅓ of our observations were missing residential segregation index data, and in addition, certain states were missing nearly all of their segregation values.
+Exploratory data analysis for the two outcome variables was conducted by 1) calculating descriptive statistics, 2) creating scatterplots of each individual predictor against the outcome (for life expectancy), and 3) creating pairwise plots between variables (for school funding).
+During the initial stages of analysis, we found that ⅓ of our observations were missing residential segregation index data, and that certain states were missing nearly all of their segregation values.
 Because of this, we decided to drop all states missing over 50% of their values for any variable.
@@
 
-For our first research question, a multiple linear regression model was fit, regressing average life expectancy on residential segregation index, percentage with limited access to healthy foods, percentage uninsured adults, and region.
+For our first research question, a multiple linear regression model was fit, regressing average life expectancy on residential segregation index, percentage with limited access to healthy foods, percentage uninsured adults, region, and median household income.
 Linear regression assumptions were assessed using diagnostic plots, especially residual vs.
 fitted and quantile-quantile plots.
-The adjusted R-squared value was used to evaluate the fit of the model, and the effect of region as an interaction with residential segregation was evaluated using nested F tests.
+The adjusted R-squared value was used to evaluate the fit of the model, and the effect of region as an interaction term with residential segregation was evaluated using nested F tests.
 We also calculated VIF to check for multicollinearity and Cook’s distance to identify influential points.
 
-For our second research question, a binary logistic regression model was fit with school funding adequacy (categorical) as the outcome and residential segregation index, median property tax, region, and median household income as predictors.
-The model was assessed using a confusion matrix, an ROC curve, and a comparison of deviance between the full model and a reduced model excluding region as a predictor.
+For our second research question, a binary logistic regression model was fit with school funding adequacy (categorical) as the outcome and residential segregation index, region, median property tax, and median household income as predictors.
+The model was assessed using a confusion matrix, an ROC curve, and a comparison of deviance between the full model and a reduced model that excluded region as a predictor.
 
@@
 
-### Data Overview
+### *Data Overview*
 
-The below tables display summary statistics on average life expectancy (1) and school funding adequacy (2).
-Both the median and mean life expectancy were 75 years.
+Table 1 displays summary statistics for the predictors in the second model, stratified by school funding adequacy.
 School funding adequacy counts show that around 66% of the counties in our analysis had under funded schools.
+Additionally, the number of counties per region indicate an imbalance between categories, with the South region being over represented (holding 60% of all counties), and the East and Northeast holding around 9% of the counties each.
+Table A1 in the Appendix provides summary statistics for all variables, and shows that both the mean and median for life expectancy are approximately 75 years.
 
-|                 | **Average Life Expectancy (years)** |
-|-----------------|-------------------------------------|
-| Min             | 65                                  |
-| 25th percentile | 73                                  |
-| Median          | 75                                  |
-| 75th percentile | 77                                  |
-| Max             | 99                                  |
-| Mean            | 75                                  |
+```{r, echo=FALSE, include=FALSE}
+completed_data |> group_by(School.Funding.Cat) |> summarise(mean(Median.Household.Income), sd(Median.Household.Income), mean(Median.Prop.Tax), sd(Median.Prop.Tax), mean(Residential.Segregation.Index), sd(Residential.Segregation.Index))
+```
 
-|                      |                           |
-|----------------------|---------------------------|
-| **0 (Under Funded)** | **1 (Adequately Funded)** |
-| 1539                 | 806                       |
+```{r, echo=FALSE, include=FALSE}
+table(completed_data$School.Funding.Cat, completed_data$Region_fac)
+```
 
-Additionally, the number of counties per region in this analysis indicate an imbalance between categories, with the South region being over represented (holding 60% of all counties), and the East and Northeast holding around 9% of the counties each.
+*Table 1:* *School Funding Adequacy Summary Statistics*
 
-|             |               |           |          |
-|-------------|---------------|-----------|----------|
-| **Midwest** | **Northeast** | **South** | **East** |
-| 552         | 203           | 1421      | 203      |
-
-As mentioned earlier, EDA for the first research question involved generating scatterplots plotting each numerical predictor plotted against life expectancy, as shown below.
-The residential segregation index vs.
-life expectancy plot was also faceted by region to explore whether each region demonstrated a different relationship in these two variables.
-Because the slopes were largely similar, this suggested that the interaction effect might not be that strong.
-In general, all three scatterplots demonstrated considerable spread and did not show a very substantial linear relationship between the numerical predictors and average life expectancy.
+| Variable                                 | **Inadequate School Funding**  | **Adequate School Funding**  |
+|------------------------------------------|--------------------------------|------------------------------|
+| Residential Segregation Index: Mean (SD) | 46 (17)                        | 56 (14)                      |
+| Region-West: Count (%)                   | 114 (7.3%)                     | 89 (10.8%)                   |
+| Region-Northeast: Count (%)              | 3 (0.2%)                       | 200 (24.3%)                  |
+| Region-Midwest: Count (%)                | 267 (17.1%)                    | 285 (34.7%)                  |
+| Region-South: Count (%)                  | 1173 (75.3%)                   | 248 (30.2%)                  |
+| Median Property Tax, USD: Mean (SD)      | 1119 (763)                     | 2282 (1648)                  |
+| Median Household Income, USD: Mean (SD)  | 56513 (13200)                  | 72218 (18255)                |
 
@@
 
-For the second research question, pairwise plots were generated to show the relationship between predictors (as seen by the scatterplots below), and also to show the relationship between each numerical predictor and the school funding outcome variable (as seen by the density plots below).
-Points and plots colored in red represent counties where school funding is inadequate (=0), and points and plots colored in blue represent counties where funding is adequate (=1).
-
 ```{r, echo=FALSE}
@@
 
-```{r, echo= FALSE, warning = FALSE, include=FALSE}
+\newpage
+
+```{r, echo=FALSE, warning = FALSE}
 # MLR
 # Outcome = life expectancy 
-# Predictors = residential segregation, limited access to healthy foods, uninsured adults, region_fac as interaction with residential segregation (or any other variable)
+# Predictors = residential segregation, limited access to healthy foods, uninsured adults, region_fac as interaction with residential segregation (or any other variable), household income
 
@@
   X..Limited.Access.to.Healthy.Foods +
-  X..Uninsured.Adults,
+  X..Uninsured.Adults + Median.Household.Income,
   data = combined)
@@
   "RSI*Region-Northeast" = "Residential.Segregation.Index:Region_facNortheast",
-  "RSI*Region-Midwest" = "Residential.Segregation.Index:Region_facMidwest")
+  "RSI*Region-Midwest" = "Residential.Segregation.Index:Region_facMidwest",
+  "Median Household Income" = "Median.Household.Income")
 
-modelsummary(mlrmodel, fmt = fmt_significant(2),
+summary(mlrmodel, fmt = fmt_significant(2),
              coef_rename = renamed, 
@@
     X..Limited.Access.to.Healthy.Foods +
-    X..Uninsured.Adults, data=completed_data)
+    X..Uninsured.Adults + Median.Household.Income, data=completed_data)
 
@@
   "Residential.Segregation.Index:Region_facNortheast"= "RSI*Region-Northeast",
-  "Residential.Segregation.Index:Region_facMidwest"="RSI*Region-Midwest")
+  "Residential.Segregation.Index:Region_facMidwest"="RSI*Region-Midwest",
+  "Median Household Income" = "Median.Household.Income")
 
-modelsummary(mlr.imp, fmt = fmt_significant(2),
+summary(mlr.imp, fmt = fmt_significant(2),
              coef_rename = renamed, 
@@
 ```{r echo= FALSE, warning = FALSE, include=FALSE}
-# Running a regression without the interaction term between residential segregation on the regions
+# Running a regression without the interaction term between residential segregation and region
 
@@
     X..Limited.Access.to.Healthy.Foods +
-    X..Uninsured.Adults + 
+    X..Uninsured.Adults + Median.Household.Income +
     Region_fac, data=completed_data)
@@
   "Region-Northeast"= "Region_facNortheast",
-  "Region-Midwest"= "Region_facMidwest")
+  "Region-Midwest"= "Region_facMidwest",
+  "Median Household Income" = "Median.Household.Income")
 
-modelsummary(mlr.imp.reduced, fmt = fmt_significant(2),
+summary(mlr.imp.reduced, fmt = fmt_significant(2),
              coef_rename = renamed, 
@@
 
+```
+
+```{r}
+##Nested F-Test to see if segregation-region interaction term makes a difference
+anova(mlr.imp, mlr.imp.reduced)
 ```
@@
 
-With variance inflation factors of roughly less than 2 for all predictors, there was little to no multicollinearity amongst the predictors.
+With variance inflation factors of roughly less than 2 for all predictors, there was little to no multicollinearity among the predictors.
 Through a nested-f test comparing a full model with a reduced model that excludes the interaction term between region and residential segregation, we concluded that a full model fit the data substantially better.
@@
 
-```{r, echo=FALSE, include=FALSE}
-#QQPlot
-qqnorm(residuals(mlr.imp))
-qqline(residuals(mlr.imp), col = "red") 
-```
+```{r, echo=FALSE}
+#Cooks Distance 
+cooksd <- cooks.distance(mlr.imp)
 
-```{r, echo=FALSE, include=FALSE}
-##Nested F-Test to see if interaction term makes a difference
-anova(mlr.imp, mlr.imp.reduced)
+# Plot Cook's Distance. Not sure if this is required or if the fourth plot in diagnostic plot shows the same thing.
+plot(cooksd, main = "Cook's Distance", ylab = "Cook's Distance")
+abline(h = 4/(nrow(combined) - length(coef(mlr.imp))), col = "red")  # Threshold line
 
@@
 
-```{r, include=FALSE, echo=FALSE}
-#Cooks Distance 
-cooksd <- cooks.distance(mlr.imp)
-
-# Plot Cook's Distance
-plot(cooksd, main = "Cook's Distance", ylab = "Cook's Distance")
-abline(h = 4/(nrow(data) - length(coef(mlr.imp))), col = "red")  # Threshold line
-
-```
-
-```{r, echo=FALSE, include = FALSE}
+```{r, echo=FALSE}
 # Identify Influential Points
@@
 
-```{r, echo=FALSE, include=FALSE}
+```{r, echo=FALSE}
 #Variance Inflaction Factor (VIF)
@@
 
-Pivoting to the second research question, we sought to assess the impact residential segregation, median property taxes, the median household income, and the region of which we are assessing has on whether or not a school is adequately funded by using a logistic regression model.
-
 #### Model Results
 
-In holding all other variables constant and using the South region as the baseline category for region, the table below displays the odds ratio results from the logistic regression model.
-
-While all estimates are statistically significant at the 95% confidence level, exponentiated regression coefficients from the logistic model demonstrate that the level of residential segregation, the median household income, and median property tax of the county have very minimal impacts on the odds of a school being adequately funded.
+The table below displays the odds ratio results from the logistic regression model for school funding adequacy, with the South as the baseline category for region. While all estimates are statistically significant at the 95% confidence level, exponentiated regression coefficients from the logistic model demonstrate that the level of residential segregation, the median household income, and median property tax of the county have very minimal impacts on the odds of a school being adequately funded.
 
@@
 Expanding the dataset to include more states and addressing missing data would also provide a more comprehensive understanding.
+
+## **Appendix**
+
+*Table A1: Summary Statistics For Numerical Variables*
+
+| Statistic          | Residential Segregation Index (RSI) | Percentage With Limited Access to Healthy Foods | Percentage of Uninsured Adults | Median Household Income, USD | Median Property Tax, USD | Average Life Expectancy |
+|--------------------|-------------------------------------|-------------------------------------------------|--------------------------------|------------------------------|--------------------------|-------------------------|
+| Min                | 1.0                                 | 0.0                                             | 2.8                            | 28972                        | 0                        | 64.9                    |
+| Median             | 50.3                                | 6.5                                             | 12.6                           | 58623                        | 1173                     | 75.1                    |
+| Mean               | 49.3                                | 8.2                                             | 14.1                           | 61940                        | 1521                     | 75.2                    |
+| Max                | 93.8                                | 62.6                                            | 47.1                           | 167605                       | 9788                     | 98.9                    |
+| Standard Deviation | 16.5                                | 6.9                                             | 6.6                            | 16877.8                      | 1274.7                   | 3.1                     |
+
+```{r}
+summary(completed_data)
+```
+
+## **References**
+
+1.
+National Academies of Sciences, Engineering, and Medicine.
+(2017).
+*Communities in action: Pathways to health equity.* Washington, DC: The National Academies Press.
+Retrieved from <https://www.ncbi.nlm.nih.gov/books/NBK425845/>
+
+2.
+Stanford Center for Opportunity Policy in Education.
+(2013.).
+How to close the opportunity gap: Key policy recommendations.
+Retrieved from <https://edpolicy.stanford.edu/sites/default/files/Opp%20Gap%20Policy%20Recommendations.pdf>
+
+3.  County Health Rankings & Roadmaps. (n.d.). *About us.* Retrieved December 8, 2024, from <https://www.countyhealthrankings.org/about-us>
+
+4.  \[INSERT IPUMS REF HERE\]
+
+5.  Halpert, C. (2014). Census regions \[GitHub repository\]. GitHub. <https://github.com/cphalpert/census-regions>
+
+6.  USDA. (2022). *Documentation*. Economic Research Service. https://www.ers.usda.gov/data-products/food-access-research-atlas/documentation/
+
+7.
+County Health Rankings & Roadmaps.
+(2024).
+*Residential segregation - Black/White.* University of Wisconsin Population Health Institute.
+Retrieved from <https://www.countyhealthrankings.org/health-data/health-factors/social-economic-factors/family-and-social-support/residential-segregation-blackwhite?utm_source=chatgpt.com&year=2024>
+
+8.
+Chetty, R., Stepner, M., Abraham, S., Lin, S., Scuderi, B., Turner, N., Bergeron, A., & Cutler, D.
+(2016).
+The Association Between Income and Life Expectancy in the United States, 2001-2014.
+*JAMA*, *315*(16), 1750–1766.
+https://doi.org/10.1001/jama.2016.4226
+
+9.
+County Health Rankings & Roadmaps.
+(2024).
+*School funding adequacy.* University of Wisconsin Population Health Institute.
+Retrieved from <https://www.countyhealthrankings.org/health-data/health-factors/social-economic-factors/education/school-funding-adequacy?utm_source=chatgpt.com&year=2024>

### 2024-12-14 paper.qmd

@@
 | Variable                                 | **Inadequate School Funding** | **Adequate School Funding** |
-|------------------------------------------|--------------------------------|------------------------------|
+|-----------------------------|----------------------|---------------------|
 | Residential Segregation Index: Mean (SD) | 46 (17)                       | 56 (14)                     |
@@
 
+```{r}
+# Remove influential points from the dataset
+combined_no_influential <- combined[-influential, ]
+
+# Refit the model without influential points
+mlr_no_influential <- lm(Life.Expectancy ~ Residential.Segregation.Index * Region_fac + X..Limited.Access.to.Healthy.Foods +X..Uninsured.Adults, data = combined_no_influential)
+
+# Compare the two models
+summary(mlr.imp)  # Original model
+summary(mlr_no_influential)  # Model without influential points
+```
+
+```{r}
+#looking for outliers 
+library(car)
+outlier_points <- outlierTest(mlr.imp)
+print(outlier_points)
+```
+
+```{r}
+outlier_indices <- as.numeric(rownames(outlier_points))  # Extract row indices
+print(outlier_indices)  # View the indices to confirm
+
+```
+
 ```{r, echo=FALSE}
@@
 plot(mlr.imp)
+```
+
+```{r}
+par(mfrow = c(2, 2), mar = c(4,4,2,2))
+plot(mlr_no_influential)
 ```
@@
 | Statistic          | Residential Segregation Index (RSI) | Percentage With Limited Access to Healthy Foods | Percentage of Uninsured Adults | Median Household Income, USD | Median Property Tax, USD | Average Life Expectancy |
-|--------------------|-------------------------------------|-------------------------------------------------|--------------------------------|------------------------------|--------------------------|-------------------------|
+|-----------|-----------|-----------|-----------|-----------|-----------|-----------|
 | Min                | 1.0                                 | 0.0                                             | 2.8                            | 28972                        | 0                        | 64.9                    |

### 2024-12-14 paper.qmd
q1 interpretation changes

@@
 | Variable                                 | **Inadequate School Funding** | **Adequate School Funding** |
-|-----------------------------|----------------------|---------------------|
+|--------------------------------|---------------------|--------------------|
 | Residential Segregation Index: Mean (SD) | 46 (17)                       | 56 (14)                     |
@@
 
-```{r, echo=FALSE}
-library(ggplot2)
-library(GGally)
-library(patchwork)
-plot1 <- ggplot(completed_data, aes(x = Residential.Segregation.Index, y = Life.Expectancy, color=Region_fac)) +
-  geom_point(color='black') + geom_smooth(method="lm",se=F) + 
-  scale_color_brewer(palette = "Dark2") +
-  labs(x = "Residential Segregation Index", y = "Life Expectancy (yrs)", color='Region')+ 
-  theme(plot.margin = margin(t = 20, r = 30, b = 20, l = 10),
-axis.title.x = element_text(size = 8),
-axis.title.y = element_text(size = 8))
-
-plot2 <- ggplot(completed_data, aes(x = X..Limited.Access.to.Healthy.Foods
-, y = Life.Expectancy)) +
-  geom_point(color = "black") + geom_smooth(method="lm",se=F)+
-  labs( x = "% Population With Limited Healthy Food Access",y = "Life Expectancy (yrs)") +
-  theme(plot.margin = margin(t = 20, r = 20, b = 20, l = 20),
-        axis.title.x = element_text(size = 8),
-        axis.title.y = element_text(size = 8))
-
-plot3 <- ggplot(completed_data, aes(x = X..Uninsured.Adults, y = Life.Expectancy)) +
-  geom_point(color = "black") +geom_smooth(method="lm",se=F)+
-  labs(x = "% Adult Population Without Health Insurance", y = "Life Expectancy (yrs)")+ 
-  theme(plot.margin = margin(t = 20, r = 20, b = 20, l = 20),
-        axis.title.x = element_text(size = 8),
-        axis.title.y = element_text(size = 8))
-
-plot1 + plot2 + plot3 + plot_layout(ncol = 2, nrow=2) + 
-  plot_annotation(title = "Relationships Between Predictors and 
-                  Life Expectancy Outcome",
-                  theme = theme(
-      plot.title = element_text(hjust=0.5)))
-```
-
-```{r, echo=FALSE}
-ggpairs(completed_data, columns = c("Residential.Segregation.Index", 
-                              "Median.Household.Income",
-                              "Median.Prop.Tax"), 
-        aes(color = factor(School.Funding.Cat))) + 
-        labs(title = "Pairwise Plots of Predictors Colored by 
-             School Funding Adequacy",
-              x = "Predictor 1",
-              y = "Predictor 2",
-              color = "School Funding Adequacy") +
-         theme(plot.title = element_text(size=15, hjust=0.5), 
-               strip.text.x = element_text(size = 5),
-               strip.text.y = element_text(size = 3))
-
-```
-
 ### Question 1
@@
 
-Based on the multiple linear regression results, percentage of uninsured adults, percentage of population with limited access to healthy foods, and region all emerged with statistically significant coefficient values.
-The interaction term between region and residential segregation, as well as the individual term of residential segregation, were not statistically significant.
-While percentage of uninsured adults and percentage with limited access to healthy foods were statistically significant, the coefficients themselves indicate very minor impacts on life expectancy (a 1 unit increase in either of these percentages leads to average life expectancy decreasing by around 0.05 years).
-The impact of region on life expectancy, however, was notable: life expectancy is on average 4.17 years higher in the West region compared to the South, 6.7 years higher in the Northeast compared to the South, and 2.41 years higher in the Midwest compared to the South.
+Based on the multiple linear regression results, residential segregation, percentage of uninsured adults, region, and median household income all emerged with statistically significant effects on life expectancy.
+However, the coefficients for most of these variables were small.
+Specifically, holding all other variables constant, for every 1 unit increase in percentage of uninsured adults, a county's average life expectancy increases by 0.032 years, which is surprising and suggests the effect of confounding variables not included in our study like ???.
+For every \$10,000 increase in median household income, life expectancy increases by 1.25 years on average.
+The impact of region on life expectancy was notable: life expectancy is on average 1.96 years higher in the West region compared to the South, 2.92 years higher in the Northeast compared to the South, and 1.92 years higher in the Midwest compared to the South, even when controlling for income, uninsured population, and racial diversity.
+This suggests that potentially the quality of healthcare and public services, which differ substantially from region to region, could be a big driver in life expectancy even outside of income (how to word / explain this better)???
+
+The percentage of population with limited access to healthy foods predictor was not statistically significant, suggesting that this environmental factor may not have an independent effect on life expectancy outside of an economic factor like income.
+In terms of residential segregation, for every 1 unit increase in a county's index value (corresponding to increased segregation), its average life expectancy increases by 0.0071 years in the South, increases by 0.0068 years in the West, decreases by 0.0055 years in the Northeast, and decreases by 0.0004 years in the Midwest, as seen in Figure 1.
+However, these differences are not statistically significant given the high p-values of all residential segregation-region interaction terms.
+We confirmed this result with a nested f-test, comparing a full model to a reduced model that excluded the interaction term between region and residential segregation.
+This did not show a statistically significant difference in RSS (*p*=0.624), indicating that the interaction between region and residential segregation does not add meaningful explanatory power to our model, and that we did not find enough evidence to support regional variations in the relationship between segregation and life expectancy (????
+is this fine????
+or am I saying something incorrect???).
+Instead, based on residential segregation's coefficient estimate in the model, life expectancy does appear to increase overall with an increase in residential segregation, however minor this increase may be.
 
@@
 
-```{r, echo=FALSE, warning = FALSE}
+```{r, echo=FALSE, warning = FALSE, include=FALSE}
 # MLR
@@
 
-summary(mlrmodel, fmt = fmt_significant(2),
+modelsummary(mlrmodel, fmt = fmt_significant(2),
              coef_rename = renamed, 
@@
 ```
+
+*Table 2: MLR Results For Average Life Expectancy Outcome*
 
@@
 
-summary(mlr.imp, fmt = fmt_significant(2),
+modelsummary(mlr.imp, fmt = fmt_significant(2),
              coef_rename = renamed, 
@@
 
-summary(mlr.imp.reduced, fmt = fmt_significant(2),
+modelsummary(mlr.imp.reduced, fmt = fmt_significant(2),
              coef_rename = renamed, 
@@
 
-```{r}
+```{r, include=FALSE, echo=FALSE}
 ##Nested F-Test to see if segregation-region interaction term makes a difference
@@
 
+*Figure 1: Segregation & Life Expectancy Across Different U.S. Regions*
+
+```{r, echo=FALSE}
+library(ggplot2)
+library(GGally)
+library(patchwork)
+
+ggplot(completed_data, aes(x = Residential.Segregation.Index, y = Life.Expectancy, color=Region_fac)) +
+  geom_point(color='black') + geom_smooth(method="lm",se=F) + 
+  scale_color_brewer(palette = "Dark2") +
+  labs(x = "Residential Segregation Index", y = "Average Life Expectancy (Years)", color='U.S. Region', title='Relationship Between Residential Segregation and Life Expectancy')+ 
+  theme(plot.margin = margin(t = 20, r = 30, b = 20, l = 10),
+axis.title.x = element_text(size = 8),
+axis.title.y = element_text(size = 8),
+title=element_text(size=9))
+```
+
 #### Model Assessment
 
-The adjusted R-squared value for the model was around 0.2, indicating that only a small amount of the variability in life expectancy was explained by the predictors in the model.
-This points to the fact that are a plethora of other socioeconomic and environmental factors that have an effect on life expectancy, and expanding our model to include more of these would have likely yielded a better fit.
-
-In conducting diagnostics on the multilinear regression model, we first checked to see if the model meets our assumptions of linearity, independence, normality, and homoscedasticity.
+The adjusted R-squared value for the model was 0.6065, indicating that a majority of the variability in life expectancy was explained by the socioeconomic and racial predictors included the model.
+In conducting diagnostics on the multilinear regression model, we first checked to see if the model met assumptions of linearity, independence, normality, and homoscedasticity.
 The residuals vs.
 fitted plot showed relatively equal variance and no discernible pattern, meaning that the linearity and homoscedasticity assumptions were reasonably met.
-In addition, because each observation in the dataset represented a unique county in the U.S., independence can be reasonably assumed.
+In addition, because each observation in the dataset represented a unique county in the U.S., independence could be reasonably assumed.
 Assessing the Q-Q Plot for normality, the residual distribution was approximately normal, despite a slight divergence on both tails of the plot.
-Finally, through observing the Residuals vs.
-Leverage plot (not pictured), we observed only a few points falling near or beyond the Cook’s distance threshold, suggesting that there may be a few influential points that can be seen as significantly impacting the results of the regression.
-These points could affect the validity of the model.
+Cook's distance was also assessed using the residuals vs.
+leverage plot; no points were found to be influential.
+There was a high leverage point (Mineral County, Nevada), and refitting the model without this point yielded no substantial changes in the coefficient estimates for statistically significant predictors; however, the p-value for residential segregation increased from 0.031 to 0.054.
+Finally, with a VIF \<2 for all predictors, there was little to no multicollinearity among the predictors.
 
-With variance inflation factors of roughly less than 2 for all predictors, there was little to no multicollinearity among the predictors.
-Through a nested-f test comparing a full model with a reduced model that excludes the interaction term between region and residential segregation, we concluded that a full model fit the data substantially better.
-The reduction in RSS in the full model compared to the reduced model was highly significant, indicating that the interaction between region and residential segregation added meaningful explanatory power to the model.
+```{r, echo=FALSE, include=FALSE}
+completed_data[1180,]
+completed_data_lev <- completed_data[-1180,]
+```
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, include=FALSE}
 #Cooks Distance 
@@
 
-```{r, echo=FALSE}
-# Identify Influential Points
-influential <- which(cooksd > 4/(nrow(combined) - length(coef(mlr.imp))))
-print(influential)
-```
-
-```{r}
-# Remove influential points from the dataset
-combined_no_influential <- combined[-influential, ]
-
-# Refit the model without influential points
-mlr_no_influential <- lm(Life.Expectancy ~ Residential.Segregation.Index * Region_fac + X..Limited.Access.to.Healthy.Foods +X..Uninsured.Adults, data = combined_no_influential)
-
-# Compare the two models
-summary(mlr.imp)  # Original model
-summary(mlr_no_influential)  # Model without influential points
-```
-
-```{r}
+```{r, echo=FALSE, include=FALSE}
 #looking for outliers 
@@
 print(outlier_points)
+outlier_indices <- as.numeric(rownames(outlier_points))  # Extract row indices
+print(outlier_indices)  # View the indices to confirm
 ```
 
-```{r}
-outlier_indices <- as.numeric(rownames(outlier_points))  # Extract row indices
-print(outlier_indices)  # View the indices to confirm
-
-```
-
-```{r, echo=FALSE}
+```{r, echo=FALSE, include=FALSE}
 #Variance Inflaction Factor (VIF)
 
-vif(mlr.imp)
+#vif(mlr.imp)
 vif(mlr.imp.reduced)
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, include=FALSE}
 #Generate 4 Diagnostic Plots at once
@@
 
-```{r}
-par(mfrow = c(2, 2), mar = c(4,4,2,2))
-plot(mlr_no_influential)
+```{r, echo= FALSE, warning = FALSE, include=FALSE}
+# Running a regression with obs 1180 removed
+
+mlr.lev <- lm(Life.Expectancy ~ Residential.Segregation.Index * Region_fac +
+    X..Limited.Access.to.Healthy.Foods +
+    X..Uninsured.Adults + Median.Household.Income, data=completed_data_lev)
+
+renamed <- c(
+    "Residential.Segregation.Index" = "Residential Segregation Index (RSI)",
+  "X..Limited.Access.to.Healthy.Foods" = "Percentage With Access to Healthy Foods",
+  "X..Uninsured.Adults"= "Percentage of Uninsured Adults",
+  "Region_facWest"="Region-West", 
+  "Region_facNortheast"="Region-Northeast",
+  "Region_facMidwest"="Region-Midwest" ,
+  "Residential.Segregation.Index:Region_facWest" = "RSI*Region-West",
+  "Residential.Segregation.Index:Region_facNortheast"= "RSI*Region-Northeast",
+  "Residential.Segregation.Index:Region_facMidwest"="RSI*Region-Midwest",
+  "Median Household Income" = "Median.Household.Income")
+
+modelsummary(mlr.lev, fmt = fmt_significant(2),
+             coef_rename = renamed, 
+             shape = term ~ model + statistic,
+             statistic = c("std.error","conf.int","p.value"),
+             exponentiate = FALSE,
+             gof_map=NA,
+             )
+```
+
+```{r, echo=FALSE, include=FALSE}
+plot(mlr.lev)
 ```
@@
 
-```{r}
+```{r, include=FALSE, echo=FALSE}
 summary(completed_data)

### 2024-12-15 paper.qmd
reformatted q2 text

--- 2024-12-14 paper.qmd
@@
 library(car)
+library(GGally)
+library(ggplot2)
+library(patchwork)
+library(caret)
+library(pROC)
+library(pander)
+library(kableExtra)
 ```
@@
 
-For this purpose, we focus our analysis on two key research questions.  First, we investigate how socioeconomic factors—residential segregation, access to healthy foods, and health insurance coverage—impact average life expectancy.
-This analysis includes a comparison across different regions to identify variations in the relationship between residential segregation and life expectancy.
+For this purpose, we focus our analysis on two key research questions.
+First, we investigate how socioeconomic factors—residential segregation, access to healthy foods, and health insurance coverage—impact average life expectancy.
+**We seek to answer the question, "What is the impact of residential segregation, access to healthy foods, and health insurance coverage on a county's average life expectancy?".** This analysis includes a comparison across different regions to identify variations in the relationship between residential segregation and life expectancy.
 Second, we evaluate how residential segregation and region influence school funding adequacy.
+**We seek the answer the question, "Does residential segregation and region impact the likelihood of a school being adequately funded?"**.
 This question allows us to assess how the racial makeup and geographic location—given that each region of the U.S.
@@
 | Variable | **Inadequate School Funding** | **Adequate School Funding** |
-|-----------------------------|----------------------|---------------------|
+|----|----|----|
 | Residential Segregation Index: Mean (SD) | 46 (17) | 56 (14) |
@@
 ```{r, echo=FALSE}
-library(ggplot2)
-library(GGally)
-library(patchwork)
 plot1 <- ggplot(completed_data, aes(x = Residential.Segregation.Index, y = Life.Expectancy, color=Region_fac)) +
@@
 
-```{r, echo=FALSE}
-ggpairs(completed_data, columns = c("Residential.Segregation.Index", 
-                              "Median.Household.Income",
-                              "Median.Prop.Tax"), 
-        aes(color = factor(School.Funding.Cat))) + 
-        labs(title = "Pairwise Plots of Predictors Colored by 
-             School Funding Adequacy",
-              x = "Predictor 1",
-              y = "Predictor 2",
-              color = "School Funding Adequacy") +
-         theme(plot.title = element_text(size=15, hjust=0.5), 
+```{r}
+library(GGally)
+library(ggplot2)
+
+ggpairs(
+  completed_data,
+  columns = c("Residential.Segregation.Index", "Median.Household.Income", "Median.Prop.Tax"),
+  aes(color = factor(School.Funding.Cat,
+                     labels = c("Inadequate", "Adequate")))
+) +
+  labs(
+    title = "Pairwise Plots of Predictors Colored by School Funding Adequacy",
+    color = "School Funding Adequacy"
+  ) +
+  theme(
+    plot.title = element_text(size = 15, hjust = 0.5),
     strip.text.x = element_text(size = 5),
-               strip.text.y = element_text(size = 3))
+    strip.text.y = element_text(size = 3)
+  )
 
@@
 #looking for outliers 
-library(car)
 outlier_points <- outlierTest(mlr.imp)
@@
 
-The table below displays the odds ratio results from the logistic regression model for school funding adequacy, with the South as the baseline category for region. While all estimates are statistically significant at the 95% confidence level, exponentiated regression coefficients from the logistic model demonstrate that the level of residential segregation, the median household income, and median property tax of the county have very minimal impacts on the odds of a school being adequately funded.
+The table below displays the odds ratio results from the logistic regression model for school funding adequacy, with the South as the baseline category for the categorical region variable. While all estimates are statistically significant at the 0.05 significance level, exponentiated regression coefficients from the logistic model demonstrate that the level of residential segregation, the median household income, and median property tax of the county have very minimal impacts on the odds of a school being adequately funded.
 
@@
 Wealthier states with a high fiscal capacity, (typically those in the Northeast), have more funding available to spend on education than states with more limited resources (typically those in the South and the West).
+
+In comparing our full model (with predictor variables for residential segregation, median household income, median property tax, and region) to a model with including all of these variables except for region, we found that there is statistical significance in the deviance between the residuals of the full model and reduced.
+As such, we can conclude that including region as a predictor refines and creates a better fit for our model.
 
@@
 Evaluating our model, we found that the logistic regression model demonstrated a strong overall performance with an accuracy rate of 80.33%.
-The model had a high sensitivity of 91%, indicating that it accurately predicted 91% of actual positive cases of inadequate school funding when evaluated on our dataset.
+The model had a high sensitivity of 91%, indicating that it accurately predicted 91% of actual positive cases of adequate school funding when evaluated on our dataset.
 However, it had a 60% specificity rate, highlighting it correctly identified cases of adequate school funding only 60% of the time.
@@
 
-In comparing our full model (with predictor variables for residential segregation, median household income, median property tax, and region) to a model with including all of these variables except for region, we found that there is statistical significance in the deviance between the residuals of the full model and reduced.
-As such, we can conclude that including region as a predictor refines and creates a better fit for our model.
-
 ```{r, echo=FALSE, include=FALSE}
 # Creating the confusion matrix
-
-library(caret)
-library(pROC)
 
@@
 # Create a confusion matrix
-library("caret")
 conf_matrix <- confusionMatrix(factor(test_pred), factor(completed_data$School.Funding.Cat))
 #print(conf_matrix)
-
-library(pander)
 
@@
 | Statistic | Residential Segregation Index (RSI) | Percentage With Limited Access to Healthy Foods | Percentage of Uninsured Adults | Median Household Income, USD | Median Property Tax, USD | Average Life Expectancy |
-|-----------|-----------|-----------|-----------|-----------|-----------|-----------|
+|----|----|----|----|----|----|----|
 | Min | 1.0 | 0.0 | 2.8 | 28972 | 0 | 64.9 |
@@
 ```{r}
-summary(completed_data)
+library(kableExtra)
+datasummary_skim(completed_data)
 ```
@@
 
-4.  \[INSERT IPUMS REF HERE\]
+4.
+Steven Ruggles, Sarah Flood, Matthew Sobek, Daniel Backman, Annie Chen, Grace Cooper, Stephanie Richards, Renae Rodgers, and Megan Schouweiler.
+IPUMS USA: Version 15.0 \[dataset\].
+Minneapolis, MN: IPUMS, 2024.
+Retrieved from <https://doi.org/10.18128/D010.V15.0>
 

### 2024-12-15 paper.qmd

--- 2024-12-14 paper.qmd
@@
 | Variable | **Inadequate School Funding** | **Adequate School Funding** |
+|----|----|----|
+| Residential Segregation Index: Mean (SD) | 46 (17) | 56 (14) |
+| Region-West: Count (%) | 114 (7.3%) | 89 (10.8%) |
+| Region-Northeast: Count (%) | 3 (0.2%) | 200 (24.3%) |
+| Region-Midwest: Count (%) | 267 (17.1%) | 285 (34.7%) |
+| Region-South: Count (%) | 1173 (75.3%) | 248 (30.2%) |
+| Median Property Tax, USD: Mean (SD) | 1119 (763) | 2282 (1648) |
+| Median Household Income, USD: Mean (SD) | 56513 (13200) | 72218 (18255) |
+
+=======
+| Variable                                 | **Inadequate School Funding** | **Adequate School Funding** |
 |--------------------------------|---------------------|--------------------|
@@
 
+```{r}
+# Plot for Q2 EDA
+ggpairs(
+  completed_data,
+  columns = c("Residential.Segregation.Index", "Median.Household.Income", "Median.Prop.Tax"),
+  aes(color = factor(School.Funding.Cat,
+                     labels = c("Inadequate", "Adequate")))
+) +
+  labs(
+    title = "Pairwise Plots of Predictors Colored by School Funding Adequacy",
+    color = "School Funding Adequacy"
+  ) +
+  theme(
+    plot.title = element_text(size = 15, hjust = 0.5),
+    strip.text.x = element_text(size = 5),
+    strip.text.y = element_text(size = 3)
+  )
+
+```
 ### Question 1

### 2024-12-15 paper.qmd
more changes

--- 2024-12-14 paper.qmd
@@
 format: pdf
+header-includes:
+      - \usepackage{float}
+      - \floatplacement{table}{H}
+fontsize: 10pt
 editor: visual
@@
 Table 1 displays summary statistics for the predictors in the second model, stratified by school funding adequacy.
-School funding adequacy counts show that around 66% of the counties in our analysis had under funded schools.
+School funding adequacy counts show that around 66% of the counties in our analysis had underfunded schools.
 Additionally, the number of counties per region indicate an imbalance between categories, with the South region being over represented (holding 60% of all counties), and the East and Northeast holding around 9% of the counties each.
@@
 
-\newpage
-
 ```{r, echo=FALSE, warning = FALSE, include=FALSE}
@@
 
+\newpage
+
+*Figure 1: Segregation & Life Expectancy Across Different U.S. Regions*
+
+```{r, echo=FALSE, fig.pos='H'}
+library(ggplot2)
+library(GGally)
+library(patchwork)
+
+ggplot(completed_data, aes(x = Residential.Segregation.Index, y = Life.Expectancy, color=Region_fac)) +
+  geom_point(color='black') + geom_smooth(method="lm",se=F) + 
+  scale_color_brewer(palette = "Dark2") +
+  labs(x = "Residential Segregation Index", y = "Average Life Expectancy (Years)", color='U.S. Region', title='Relationship Between Residential Segregation and Life Expectancy')+ 
+  theme(plot.margin = margin(t = 20, r = 30, b = 20, l = 10),
+axis.title.x = element_text(size = 8),
+axis.title.y = element_text(size = 8),
+title=element_text(size=9))
+```
+
 *Table 2: MLR Results For Average Life Expectancy Outcome*
 
-```{r, echo= FALSE, warning = FALSE}
+```{r, echo= FALSE, warning = FALSE, fig.pos='H'}
 # Running a regression with the imputated data (completed_data) to understand the impact of missing data
@@
 anova(mlr.imp, mlr.imp.reduced)
-```
-
-*Figure 1: Segregation & Life Expectancy Across Different U.S. Regions*
-
-```{r, echo=FALSE}
-library(ggplot2)
-library(GGally)
-library(patchwork)
-
-ggplot(completed_data, aes(x = Residential.Segregation.Index, y = Life.Expectancy, color=Region_fac)) +
-  geom_point(color='black') + geom_smooth(method="lm",se=F) + 
-  scale_color_brewer(palette = "Dark2") +
-  labs(x = "Residential Segregation Index", y = "Average Life Expectancy (Years)", color='U.S. Region', title='Relationship Between Residential Segregation and Life Expectancy')+ 
-  theme(plot.margin = margin(t = 20, r = 30, b = 20, l = 10),
-axis.title.x = element_text(size = 8),
-axis.title.y = element_text(size = 8),
-title=element_text(size=9))
 ```
@@
 ```
+
+*Table 3: Logistic Regression Results For School Funding Adequacy Outcome*
 

### 2024-12-15 paper.qmd
most changes

@@
 This study examines the relationships between socioeconomic, health, and education indicators in U.S.
-counties using data from the American Community Survey, the USDA Food Environment Atlas, small area income and poverty estimates, and other sources.
-We will explore regional influences and predictors of life expectancy, and also identify socioeconomic factors associated with school funding adequacy.
-By analyzing differences in life expectancy and education funding across socioeconomic and regional contexts, this study aims to gain insight into public well-being and economic inequality across the U.S.
-Through exploratory data analysis and regression modeling with interaction terms, we identify significant regional disparities in both life expectancy and school funding adequacy.
-The findings highlight that regions, particularly the West and Northeast, tend to have better health and education outcomes, emphasizing the critical role of location and economic conditions in shaping societal well-being.
+counties using data from the American Community Survey, the USDA Food Environment Atlas, Small Area Income and Poverty Estimates, and other sources.We analyze how factors such as residential segregation, health insurance coverage, access to healthy foods, and region influence life expectancy, and how residential segregation and region affect school funding adequacy.
+Our findings reveal that life expectancy is influenced by residential segregation and uninsured rates, though these effects are minor compared to the more substantial influence of household income and regional differences.
+Life expectancy averages are highest in the Northeast, followed by the Midwest, West, and South.
+Region also plays the most significant role out of all included variables in determining the likelihood of school funding adequacy, with schools in the Northeast and Midwest being far more likely to receive adequate funding than those in the South.
+Other factors, such as residential segregation, property taxes, and household income, show statistically significant but smaller effects.
+These results highlight the complex interplay between socioeconomic variables and geographic location in shaping public well-being.
+By identifying key disparities in health and education, this study underscores the importance of equitable resource allocation and targeted policies to address regional and socioeconomic inequalities.
 
@@
 Life expectancy is an important health indicator that reflects not only the overall health of a society, but also its stability, economic resilience, and levels of inequality.
-According to "The Root Causes of Health Inequity - Communities in Action" report from the National Institutes of Health, we can infer that differences in life expectancy between regions and socioeconomic groups often reveal systemic inequalities, with marginalized communities experiencing shorter life spans due to limited access to health care, healthy living environments, and economic opportunities.^1^ Similarly, adequately funded schools ensure that students have equal access to learning opportunities, serving as a crucial investment in social mobility.
+While access to quality healthcare and nutritious foods are drivers of well-being, where you live and the nature of the community you live in likely also have an influence on health outcomes such as life expectancy.
+According to "The Root Causes of Health Inequity - Communities in Action" report from the National Institutes of Health, we can infer that differences in life expectancy between regions and socioeconomic groups often reveal systemic inequalities, with marginalized communities experiencing shorter life spans due to limited access to health care, healthy living environments, and economic opportunities.^1^ When it comes to educational indicators, adequately funded schools ensure that students have equal access to learning opportunities, serving as a crucial investment in social mobility.
 The Stanford Center for Opportunity Policy in Education highlights that "the unequal allocation and inadequate levels of resources in schools and communities \[are\] at the heart of many gaps in student opportunity."^2^ This emphasizes the importance of equitable and adequate school funding in closing opportunity gaps and empowering students from disadvantaged backgrounds to achieve academic and economic success.
-Consequently, it can be inferred that underfunded schools, often situated in economically disadvantaged or segregated areas, exacerbate educational disparities and limit students' ability to break the cycle of poverty.
-These gaps are rooted in systemic inequalities, such as residential segregation, which concentrates poverty, limits access to quality education and healthcare, and perpetuates income and health disparities across generations.
-By addressing these inequalities, policymakers can foster fairer, more prosperous communities and ensure that everyone benefits from improved health and education systems.
+By better understanding the correlates of school funding adequacy, we can take a major step towards implementing policy that targets the areas and communities most affected by this issue.
 
 For this purpose, we focus our analysis on two key research questions.
-First, we investigate how socioeconomic factors—residential segregation, access to healthy foods, and health insurance coverage—impact average life expectancy.
+First, we investigate how socioeconomic factors impact average life expectancy.
 **We seek to answer the question, "What is the impact of residential segregation, access to healthy foods, and health insurance coverage on a county's average life expectancy?".** This analysis includes a comparison across different regions to identify variations in the relationship between residential segregation and life expectancy.
-Second, we evaluate how residential segregation and region influence school funding adequacy.
-**We seek the answer the question, "Does residential segregation and region impact the likelihood of a school being adequately funded?"**.
+Second, **we seek to answer the question, "Do residential segregation and region impact the likelihood of a school being adequately funded?"**.
 This question allows us to assess how the racial makeup and geographic location—given that each region of the U.S.
@@
 
-Variables of interest were selected *a priori* based on their perceived relevance to the two outcomes we wanted to better understand: average life expectancy and school funding adequacy.
+Variables of interest were selected a priori based on their perceived relevance to the two outcomes we wanted to better understand: average life expectancy and school funding adequacy.
 While nutrition and access to quality healthcare are inherently linked to general wellbeing, it is worthwhile to assess how influential the environmental and financial aspects of these factors are on our tangible outcome.
@@
 We also were interested in seeing how residential segregation would impact life expectancy among these predictors, given existing studies on the negative health effects of segregation, and whether this relationship would exhibit any variations across different regions.
-This is why we chose to include residential segregation index values.
+This is why we chose to include residential segregation index values as a predictor.
 The residential segregation index, used by the County Health Rankings & Roadmaps program, measures how evenly two groups, such as Black and white residents, are distributed across geographic areas, with values ranging from 0 (complete integration) to 100 (complete segregation).^7^
 
-*Income is known to have a strong relationship with.*^8^
+Past studies have established a correlation between income and life expectancy.^8^ Because we wanted to control for its effects and examine how insurance, healthy food access, segregation, and region influence life expectancy even within a given income bracket, we selected income as our final predictor for the first model.
 
@@
 In addition, given that state policy has a large influence on school administration and the determination of school needs, and that states in the same region tend to have similar political leanings and cultural beliefs, we wanted to analyze how much of a relationship exists between region and actual school funding levels.
-
 Because a major source of public school funding is tax revenue, median property tax and median household income were also included in the school funding model as potential confounders.
@@
 
-After this, remaining missing values in all variables were imputed using the mice package and predictive mean matching to generate one complete dataset.
+After this, the remaining missing values in all variables were imputed using the mice package and predictive mean matching to generate one complete dataset.
 The original school funding adequacy numerical variable was first imputed, and then converted into binary values for subsequent analysis.
@@
 The model was assessed using a confusion matrix, an ROC curve, and a comparison of deviance between the full model and a reduced model that excluded region as a predictor.
+VIF and Cook's distance were again used to evaluate multicollinearity and influential points.
 
@@
 Table 1 displays summary statistics for the predictors in the second model, stratified by school funding adequacy.
+Table A1 in the Appendix provides summary statistics for all variables, and shows that both the mean and median for life expectancy are approximately 75 years.
 School funding adequacy counts show that around 66% of the counties in our analysis had underfunded schools.
 Additionally, the number of counties per region indicate an imbalance between categories, with the South region being over represented (holding 60% of all counties), and the East and Northeast holding around 9% of the counties each.
-Table A1 in the Appendix provides summary statistics for all variables, and shows that both the mean and median for life expectancy are approximately 75 years.
 
@@
 
+\newpage
+
 *Table 1:* *School Funding Adequacy Summary Statistics*
@@
 | Variable                                 | **Inadequate School Funding** | **Adequate School Funding** |
-|----|----|----|
+|------------------------------|----------------------|--------------------|
 | Residential Segregation Index: Mean (SD) | 46 (17)                       | 56 (14)                     |
@@
 
-=======
-| Variable                                 | **Inadequate School Funding** | **Adequate School Funding** |
-|--------------------------------|---------------------|--------------------|
-| Residential Segregation Index: Mean (SD) | 46 (17)                       | 56 (14)                     |
-| Region-West: Count (%)                   | 114 (7.3%)                    | 89 (10.8%)                  |
-| Region-Northeast: Count (%)              | 3 (0.2%)                      | 200 (24.3%)                 |
-| Region-Midwest: Count (%)                | 267 (17.1%)                   | 285 (34.7%)                 |
-| Region-South: Count (%)                  | 1173 (75.3%)                  | 248 (30.2%)                 |
-| Median Property Tax, USD: Mean (SD)      | 1119 (763)                    | 2282 (1648)                 |
-| Median Household Income, USD: Mean (SD)  | 56513 (13200)                 | 72218 (18255)               |
-
-```{r}
+```{r, echo=FALSE, include=FALSE}
 # Plot for Q2 EDA
@@
 ```
+
 ### Question 1
@@
 
-Based on the multiple linear regression results, residential segregation, percentage of uninsured adults, region, and median household income all emerged with statistically significant effects on life expectancy.
-However, the coefficients for most of these variables were small.
-Specifically, holding all other variables constant, for every 1 unit increase in percentage of uninsured adults, a county's average life expectancy increases by 0.032 years, which is surprising and suggests the effect of confounding variables not included in our study like ???.
-For every \$10,000 increase in median household income, life expectancy increases by 1.25 years on average.
-The impact of region on life expectancy was notable: life expectancy is on average 1.96 years higher in the West region compared to the South, 2.92 years higher in the Northeast compared to the South, and 1.92 years higher in the Midwest compared to the South, even when controlling for income, uninsured population, and racial diversity.
-This suggests that potentially the quality of healthcare and public services, which differ substantially from region to region, could be a big driver in life expectancy even outside of income (how to word / explain this better)???
+Based on the multiple linear regression results, residential segregation, percentage of uninsured adults, region, and median household income all showed statistically significant effects on life expectancy.
+Specifically, for every \$10,000 increase in median household income, life expectancy increases by 1.25 years on average.
+In addition, the impact of region on life expectancy was substantial: life expectancy is on average 1.96 years higher in the West region compared to the South, 2.92 years higher in the Northeast compared to the South, and 1.92 years higher in the Midwest compared to the South, even when controlling for income, uninsured population, and racial diversity.
+This suggests, income as a factor withstanding, the differing quality of and access to healthcare and public services found in each region can have a significant impact on life expectancy.
 
-The percentage of population with limited access to healthy foods predictor was not statistically significant, suggesting that this environmental factor may not have an independent effect on life expectancy outside of an economic factor like income.
+Surprisingly, holding all other variables constant, for every 1 unit increase in percentage of uninsured adults in a county, a county's average life expectancy increases by 0.032 years.
+While statistically significant, many studies covering the relationship between an individual’s health and having insurance coverage have proven its positive correlation, suggesting our results may benefit from a closer look.
+In assessing our dataset, as seen in Appendix 1A, we find that our dataset contains more observations for counties with low uninsured individuals, or equivalently, counties where most people are covered.
+As the timeframe of our dataset is 2018-2022, it is likely that the aforementioned observations require greater scrutiny of the significance of the impact of this variable on average life expectancy of a county.
+
+The percentage of population with limited access to healthy foods predictor was not statistically significant in our model, suggesting that the effect of this environmental factor is minor or intertwined with economic variables, such as income, rather than being a direct or standalone influence.
 In terms of residential segregation, for every 1 unit increase in a county's index value (corresponding to increased segregation), its average life expectancy increases by 0.0071 years in the South, increases by 0.0068 years in the West, decreases by 0.0055 years in the Northeast, and decreases by 0.0004 years in the Midwest, as seen in Figure 1.
-However, these differences are not statistically significant given the high p-values of all residential segregation-region interaction terms.
-We confirmed this result with a nested f-test, comparing a full model to a reduced model that excluded the interaction term between region and residential segregation.
-This did not show a statistically significant difference in RSS (*p*=0.624), indicating that the interaction between region and residential segregation does not add meaningful explanatory power to our model, and that we did not find enough evidence to support regional variations in the relationship between segregation and life expectancy (????
-is this fine????
-or am I saying something incorrect???).
+However, these regional differences are not statistically significant at the 0.05 significance level given the high p-values of all residential segregation-region interaction terms.
+We confirmed this result with a nested F-test, comparing a full model to a reduced model that excluded the interaction term between region and residential segregation.
+This did not show a statistically significant difference in RSS (p=0.624), indicating that the interaction between region and residential segregation does not add meaningful explanatory power to our model, Consequently, we did not find sufficient evidence to support regional variations in the relationship between segregation and life expectancy.
 Instead, based on residential segregation's coefficient estimate in the model, life expectancy does appear to increase overall with an increase in residential segregation, however minor this increase may be.
+One possible theory for this observation is that counties with more racial uniformity might experience less racial conflict and more social cohesion, especially for minority communities.
 
-```{r, echo=FALSE, warning = FALSE, include=FALSE}
-# MLR
-# Outcome = life expectancy 
-# Predictors = residential segregation, limited access to healthy foods, uninsured adults, region_fac as interaction with residential segregation (or any other variable), household income
-
-# To get a general sense of what the data tells us, we use the orignial dataset with missing values.
-
-mlrmodel <- lm(
-  Life.Expectancy ~ Residential.Segregation.Index * Region_fac +
-  X..Limited.Access.to.Healthy.Foods +
-  X..Uninsured.Adults + Median.Household.Income,
-  data = combined)
-
-# Renaming the output names 
-renamed <- c(
-   "Residential Segregation Index (RSI)" = "Residential.Segregation.Index",
-  "Percentage With Limited Access to Healthy Foods" = "X..Limited.Access.to.Healthy.Foods",
-  "Percentage of Uninsured Adults"= "X..Uninsured.Adults",
-  "Region-West" = "Region_facWest", 
-  "Region-Northeast"= "Region_facNortheast",
-  "Region-Midwest"= "Region_facMidwest",
-  "RSI*Region-West" = "Residential.Segregation.Index:Region_facWest",
-  "RSI*Region-Northeast" = "Residential.Segregation.Index:Region_facNortheast",
-  "RSI*Region-Midwest" = "Residential.Segregation.Index:Region_facMidwest",
-  "Median Household Income" = "Median.Household.Income")
-
-modelsummary(mlrmodel, fmt = fmt_significant(2),
-             coef_rename = renamed, 
-             shape = term ~ model + statistic,
-             statistic = c("std.error","conf.int","p.value"),
-             exponentiate = FALSE,
-             gof_map=NA,
-             )
-
-```
-
-\newpage
-
-*Figure 1: Segregation & Life Expectancy Across Different U.S. Regions*
+*Figure 1: Segregation vs. Life Expectancy Across Different U.S. Regions*
 
@@
 
-The adjusted R-squared value for the model was 0.6065, indicating that a majority of the variability in life expectancy was explained by the socioeconomic and racial predictors included the model.
+The adjusted R-squared value for the model was 0.6065, indicating that a majority of the variability in life expectancy was explained by the socioeconomic and racial predictors included in the model.
 In conducting diagnostics on the multilinear regression model, we first checked to see if the model met assumptions of linearity, independence, normality, and homoscedasticity.
@@
 completed_data_lev <- completed_data[-1180,]
-```
-
-```{r, echo=FALSE, include=FALSE}
-#Cooks Distance 
-cooksd <- cooks.distance(mlr.imp)
-
-# Plot Cook's Distance. Not sure if this is required or if the fourth plot in diagnostic plot shows the same thing.
-plot(cooksd, main = "Cook's Distance", ylab = "Cook's Distance")
-abline(h = 4/(nrow(combined) - length(coef(mlr.imp))), col = "red")  # Threshold line
-
 ```
@@
 
-The table below displays the odds ratio results from the logistic regression model for school funding adequacy, with the South as the baseline category for the categorical region variable. While all estimates are statistically significant at the 0.05 significance level, exponentiated regression coefficients from the logistic model demonstrate that the level of residential segregation, the median household income, and median property tax of the county have very minimal impacts on the odds of a school being adequately funded.
+Table 3 below displays the odds ratio results from the logistic regression model for school funding adequacy, with the South as the baseline category for the categorical region variable. While all estimates are statistically significant at the 0.05 significance level, exponentiated regression coefficient estimates from the logistic model demonstrate that the level of residential segregation, the median household income, and median property tax of the county have very minimal impacts on the odds of a school being adequately funded.
 
@@
 
-In comparing our full model (with predictor variables for residential segregation, median household income, median property tax, and region) to a model with including all of these variables except for region, we found that there is statistical significance in the deviance between the residuals of the full model and reduced.
+In comparing our full model (with predictor variables for residential segregation, median household income, median property tax, and region) to a model including all of these variables except for region, we found that there is statistical significance in the deviance between the residuals of the full model and reduced.
 As such, we can conclude that including region as a predictor refines and creates a better fit for our model.
@@
 
+*Figure 2: Relationship Between Residential Segregation, Region, & School Funding Adequacy*
+
+```{r}
+ggplot(completed_data, aes(x=Region_fac, y = Residential.Segregation.Index, color=factor(School.Funding.Cat))) +
+  geom_boxplot() +
+  labs(x = "U.S. Region", y='Residential Segregation Index', color='School Funding Adequacy', title='Residential Segregation Stratified by Region & School Funding') +
+  scale_color_hue(labels = c("Inadequate", "Adequate"))+ 
+  theme(plot.margin = margin(t = 20, r = 30, b = 20, l = 10),
+axis.title.x = element_text(size = 8),
+axis.title.y = element_text(size = 8),
+title=element_text(size=9))
+```
+
 *Table 3: Logistic Regression Results For School Funding Adequacy Outcome*
@@
 
+To assess logistic regression assumptions, we first plotted each numerical predictor against the log odds of school funding adequacy.
+All plots showed a mostly linear relationship, indicating that the linearity assumption was largely met; interestingly, different linear trends were observed within each plot based on region, pointing to potential interaction effects between region and the three numerical predictors.
+GVIF for all predictors ranged between 1 and 3, suggesting that there could be moderate multicollinearity between predictors, but not enough to warrant a reconsideration of included variables.
+Cook's distance was also less than 0.05 for all points, meaning that there were no highly influential observations in the model.
 Evaluating our model, we found that the logistic regression model demonstrated a strong overall performance with an accuracy rate of 80.33%.
 The model had a high sensitivity of 91%, indicating that it accurately predicted 91% of actual positive cases of adequate school funding when evaluated on our dataset.
-However, it had a 60% specificity rate, highlighting it correctly identified cases of adequate school funding only 60% of the time.
-Moreover, looking at the ROC curve below, the AUC is 0.858, indicating a moderately strong discriminate ability.
+However, it had a 60% specificity rate, highlighting it correctly identified cases of inadequate school funding only 60% of the time.
+Overall, looking at the ROC curve below, the AUC is 0.858, indicating a moderately strong discriminate ability.
+
+```{r, include=FALSE, echo=FALSE}
+plot(completed_data$Residential.Segregation.Index, predict(imp_mods), col = completed_data$Region_fac)
+plot(completed_data$Median.Household.Income, predict(imp_mods), col = completed_data$Region_fac)
+plot(completed_data$Median.Prop.Tax, predict(imp_mods), col = completed_data$Region_fac)
+```
+
+```{r, include=FALSE, echo=FALSE}
+#Cooks Distance 
+cooksd <- cooks.distance(imp_mods)
+
+# Plot Cook's Distance
+plot(cooksd, main = "Cook's Distance", ylab = "Cook's Distance")
+abline(h = 4/(nrow(completed_data) - length(coef(imp_mods))), col = "red")  # Threshold line
+4/(nrow(data) - length(coef(imp_mods)))
+```
+
+```{r, echo=FALSE, include=FALSE}
+vif(imp_mods)
+plot(imp_mods)
+```
 
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, include=FALSE}
 #Creating the AUC curves
@@
 
-By analyzing these two research questions, we observed the significant impact of region on both life expectancy (a health indicator) and school funding (an education indicator), both of which are key components of societal well-being.
-These indices tend to be higher in the West and Northeast, which are more developed regions.
-The findings provide empirical evidence supporting the significant disparities in life expectancy across regions with differing economic conditions.
-They also provide insights into social policy investment decisions, highlighting the importance of fair resource allocation to people's well-being.
+In our analysis, we observed the significant impact of the regional location of a county on both its average life expectancy (a health indicator) and school funding adequacy (an educational indicator)--two key components of societal and an individual’s well-being.
+Life expectancy is highest in the Northeast, followed by the Midwest, West, and South, while school funding adequacy is significantly greater in the Northeast and Midwest compared to the South.
+Our findings highlight substantial differences in both outcomes between regions, which are often caused by different economic capabilities and different allocations of resources among countries.
+This study provides empirical evidence that underscores the importance for equitable social policies and targeted investments to address regional disparities and promote well-being across all communities.
 
-Our analysis faced limitations due to missing data, causing our dataset to be restricted to 33 states.
-Additionally, because The County Health Rankings program consolidates the latest from a variety of governmental sources, most of our variables were pulled from different years between 2018 and 2022; we assumed that trends remained largely consistent between these years, treating time as insignificant, but this may not actually be the case.
+Our analysis faced limitations due to missing data, which restricted the dataset to 33 states and reduced the generalizability of our findings.
+Additionally, some variables were drawn from different years between 2018 and 2022, requiring us to assume that trends remained consistent over a 1–5 year period and that time was an insignificant factor; however, this may not actually be the case.
 
-Future studies could incorporate more social variables to enhance the analysis and investigate the causal relationships between region and outcomes.
-Expanding the dataset to include more states and addressing missing data would also provide a more comprehensive understanding.
+Future studies could incorporate additional social variables to enhance the analysis and explore causal relationships between region and outcomes.
+For example, integrating a dataset on employment rates and school performance could offer deeper insights into how socioeconomic factors shape life expectancy and educational funding, complementing existing variables in our models.
+Another direction for future work is to delve deeper into understanding the surprising findings regarding the effects of uninsured adult percentage and residential segregation on life expectancy.
+Expanding the dataset to include a temporal scope, and finding more complete sources of data on variables of interest so that no states have to be excluded from the analysis, would also contribute to a more comprehensive understanding.
 
@@
 
-*Table A1: Summary Statistics For Numerical Variables*
+*Table A1: Summary Statistics For All Variables*
 
-| Statistic | Residential Segregation Index (RSI) | Percentage With Limited Access to Healthy Foods | Percentage of Uninsured Adults | Median Household Income, USD | Median Property Tax, USD | Average Life Expectancy |
-|----|----|----|----|----|----|----|
-| Min | 1.0 | 0.0 | 2.8 | 28972 | 0 | 64.9 |
-| Median | 50.3 | 6.5 | 12.6 | 58623 | 1173 | 75.1 |
-| Mean | 49.3 | 8.2 | 14.1 | 61940 | 1521 | 75.2 |
-| Max | 93.8 | 62.6 | 47.1 | 167605 | 9788 | 98.9 |
-| Standard Deviation | 16.5 | 6.9 | 6.6 | 16877.8 | 1274.7 | 3.1 |
+![](images/Screenshot 2024-12-15 at 2.38.35 AM.png)
+
+```{r}
+renamed_data <- completed_data %>%
+  rename(
+    "Residential Segregation Index" = Residential.Segregation.Index,
+    "Median Household Income" = Median.Household.Income,
+    "Median Property Tax" = Median.Prop.Tax,
+    "School Funding Adequacy" = School.Funding.Adequacy, 
+    "Average Life Expectancy" = Life.Expectancy,
+    "% Limited Access to Healthy Foods" = X..Limited.Access.to.Healthy.Foods,
+    "% Uninsured Adults" = X..Uninsured.Adults,
+    "Region" = Region_fac,
+    "School Funding Binary" =School.Funding.Cat)
+
+datasummary_skim(
+  renamed_data,
+  title = "Customized Summary Table")
+
+```
 

### 2024-12-15 paper.qmd
fixes

@@
 
-## **Abstract**
+## Abstract
 
@@
 
-![](images/Screenshot 2024-12-15 at 2.38.35 AM.png)
-
 ```{r}
@@
   rename(
-    "Residential Segregation Index" = Residential.Segregation.Index,
+    "Residential Segregation" = Residential.Segregation.Index,
     "Median Household Income" = Median.Household.Income,
@@
     "School Funding Adequacy" = School.Funding.Adequacy, 
-    "Average Life Expectancy" = Life.Expectancy,
-    "% Limited Access to Healthy Foods" = X..Limited.Access.to.Healthy.Foods,
+    "Life Expectancy" = Life.Expectancy,
+    "% Limited Access, Healthy Foods" = X..Limited.Access.to.Healthy.Foods,
     "% Uninsured Adults" = X..Uninsured.Adults,
@@
 
-datasummary_skim(
+skim_table <- datasummary_skim(
   renamed_data,
-  title = "Customized Summary Table")
+  title = "Summary Statistics")
 

### 2024-12-15 paper.tex
other files

--- 2024-12-15 paper.qmd
@@
----
-title: "Understanding the Impact of Socioeconomic and Regional Factors on Health and Educational Outcomes"
-author: "Rishika Randev, Jenny Wu, Uzoma Uwazurike Jr., Shiyue Zhou"
-instructor: "Andrea Lane"
-institution: "Duke University, IDS 702 Data Modeling and Presentation"
-format: pdf
-header-includes:
-      - \usepackage{float}
-      - \floatplacement{table}{H}
-fontsize: 10pt
-editor: visual
-execute:
-  echo: false
-  warning: false
----
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
+%
+\documentclass[
+  10pt,
+  letterpaper,
+  DIV=11,
+  numbers=noendperiod]{scrartcl}
 
-```{r, echo= FALSE, warning = FALSE, message=FALSE}
-library("dplyr") 
-library(mice) 
-library(sjlabelled) 
-library(tidyverse) 
-library("modelsummary") 
-library(car)
-library(GGally)
-library(ggplot2)
-library(patchwork)
-library(caret)
-library(pROC)
-library(pander)
-library(kableExtra)
-```
+\usepackage{amsmath,amssymb}
+\usepackage{iftex}
+\ifPDFTeX
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+\usepackage{lmodern}
+\ifPDFTeX\else  
+    % xetex/luatex font selection
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
+  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+% Make \paragraph and \subparagraph free-standing
+\ifx\paragraph\undefined\else
+  \let\oldparagraph\paragraph
+  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+  \let\oldsubparagraph\subparagraph
+  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
 
-```{r,echo= FALSE, warning = FALSE, message=FALSE}
-combined <- read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/combined.csv')
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
+\usepackage{graphicx}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
 
-```
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
+\usepackage{tabularray}
+\usepackage[normalem]{ulem}
+\usepackage{graphicx}
+\UseTblrLibrary{booktabs}
+\UseTblrLibrary{rotating}
+\UseTblrLibrary{siunitx}
+\NewTableCommand{\tinytableDefineColor}[3]{\definecolor{#1}{#2}{#3}}
+\newcommand{\tinytableTabularrayUnderline}[1]{\underline{#1}}
+\newcommand{\tinytableTabularrayStrikeout}[1]{\sout{#1}}
+\KOMAoption{captions}{tableheading}
+\usepackage{float}
+\floatplacement{table}{H}
+\makeatletter
+\@ifpackageloaded{caption}{}{\usepackage{caption}}
+\AtBeginDocument{%
+\ifdefined\contentsname
+  \renewcommand*\contentsname{Table of contents}
+\else
+  \newcommand\contentsname{Table of contents}
+\fi
+\ifdefined\listfigurename
+  \renewcommand*\listfigurename{List of Figures}
+\else
+  \newcommand\listfigurename{List of Figures}
+\fi
+\ifdefined\listtablename
+  \renewcommand*\listtablename{List of Tables}
+\else
+  \newcommand\listtablename{List of Tables}
+\fi
+\ifdefined\figurename
+  \renewcommand*\figurename{Figure}
+\else
+  \newcommand\figurename{Figure}
+\fi
+\ifdefined\tablename
+  \renewcommand*\tablename{Table}
+\else
+  \newcommand\tablename{Table}
+\fi
+}
+\@ifpackageloaded{float}{}{\usepackage{float}}
+\floatstyle{ruled}
+\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
+\floatname{codelisting}{Listing}
+\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
+\makeatother
+\makeatletter
+\makeatother
+\makeatletter
+\@ifpackageloaded{caption}{}{\usepackage{caption}}
+\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
+\makeatother
+\ifLuaTeX
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
+\usepackage{bookmark}
 
-```{r, echo= FALSE, warning = FALSE}
-# Making the state_fac variable a factor variable for future use
-combined <- combined |> mutate(State_fac = factor(combined$State), Region_fac = factor(combined$Region, levels=c('South', 'West', 'Northeast', 'Midwest')))
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same} % disable monospaced font for URLs
+\hypersetup{
+  pdftitle={Understanding the Impact of Socioeconomic and Regional Factors on Health and Educational Outcomes},
+  pdfauthor={Rishika Randev, Jenny Wu, Uzoma Uwazurike Jr., Shiyue Zhou},
+  colorlinks=true,
+  linkcolor={blue},
+  filecolor={Maroon},
+  citecolor={Blue},
+  urlcolor={Blue},
+  pdfcreator={LaTeX via pandoc}}
 
-```
+\title{Understanding the Impact of Socioeconomic and Regional Factors on
+Health and Educational Outcomes}
+\author{Rishika Randev, Jenny Wu, Uzoma Uwazurike Jr., Shiyue Zhou}
+\date{}
 
-```{r, echo= FALSE, warning = FALSE}
-## Imputation Code 
-# In this block, we are looking to use predictive mean matching to combing regression methods and hot deck imputing to predicted value closest to the predicted value of the observation with missing data; we also needed to convert school funding adequacy imputed values to school funding categorical variable before using in models
+\begin{document}
+\maketitle
 
-sub <- combined |> select(c("Life.Expectancy","Residential.Segregation.Index","X..Limited.Access.to.Healthy.Foods","X..Uninsured.Adults","Median.Household.Income", "School.Funding.Adequacy", "Median.Prop.Tax", "Region_fac"))
-sub <- unlabel(sub) #unlabel the data (labels cause problem for the mice function)
+\subsection{Abstract}\label{abstract}
 
-# Add School.Funding.Cat to the orginal dataset.
-Since we do not want to imputate on the categorical outcome variable,  this creation of a variable exists to create the proper dataframe within the imputated data
-sub$School.Funding.Cat <- ifelse(sub$School.Funding.Adequacy > 0, 1, 0)
+This study examines the relationships between socioeconomic, health, and
+education indicators in U.S. counties using data from the American
+Community Survey, the USDA Food Environment Atlas, Small Area Income and
+Poverty Estimates, and other sources.We analyze how factors such as
+residential segregation, health insurance coverage, access to healthy
+foods, and region influence life expectancy, and how residential
+segregation and region affect school funding adequacy. Our findings
+reveal that life expectancy is influenced by residential segregation and
+uninsured rates, though these effects are minor compared to the more
+substantial influence of household income and regional differences. Life
+expectancy averages are highest in the Northeast, followed by the
+Midwest, West, and South. Region also plays the most significant role
+out of all included variables in determining the likelihood of school
+funding adequacy, with schools in the Northeast and Midwest being far
+more likely to receive adequate funding than those in the South. Other
+factors, such as residential segregation, property taxes, and household
+income, show statistically significant but smaller effects. These
+results highlight the complex interplay between socioeconomic variables
+and geographic location in shaping public well-being. By identifying key
+disparities in health and education, this study underscores the
+importance of equitable resource allocation and targeted policies to
+address regional and socioeconomic inequalities.
 
-# Perform the imputation
-sub.imp <- mice(sub, m = 1, method = "pmm", print = FALSE, seed=3)
+\subsection{\texorpdfstring{\textbf{Introduction}}{Introduction}}\label{introduction}
 
-```
-
-```{r, echo= FALSE, warning = FALSE}
-# Since we should not imputate on the outcome, but need to imputate the School.Funding.Adequacy variable, we are feeding the School.Funding.Adequacy variable into a ifelse conditional that will replaced the imputed outcome variable.
-
-# Define the categorization function
-categorize_school_funding <- function(x) {
-  ifelse(x > 0, 1, 0)  # Binary classification
-}
-
-# Loop over the imputations
-for (i in 1:1) {
-  # Extract the imputed values for the i-th dataset
-  imputed_values <- sub.imp$imp$School.Funding.Adequacy[, i]
-
-  # Categorize the imputed values
-  categorized_values <- categorize_school_funding(imputed_values)
-  complete_cat <- rep(NA, nrow(sub))  # Initialize with NAs
-
-  # Get the indices where imputation occurred
-  imp_indices <- which(!is.na(categorized_values))  # Only for imputed values
-
-  # Assign the categorized values to the rows that were imputed
-  complete_cat[imp_indices] <- categorized_values[imp_indices]
-
-  # Save into the `imp` list of the `sub.imp` object
-  sub.imp$imp$School.Funding.Cat[[i]] <- complete_cat[imp_indices]
-}
-
-```
-
-```{r, echo= FALSE, warning = FALSE, include = FALSE }
-# To have a full dataset, we are taking the first imputed dataset R provided and using that as our full dataset sample. 
-completed_data <- complete(sub.imp, 1)
-
-# Check the structure
-str(completed_data)
-
-# Ensure that there are no missing data 
-colSums(is.na(completed_data))
-```
-
-## Abstract
-
-This study examines the relationships between socioeconomic, health, and education indicators in U.S.
-counties using data from the American Community Survey, the USDA Food Environment Atlas, Small Area Income and Poverty Estimates, and other sources.We analyze how factors such as residential segregation, health insurance coverage, access to healthy foods, and region influence life expectancy, and how residential segregation and region affect school funding adequacy.
-Our findings reveal that life expectancy is influenced by residential segregation and uninsured rates, though these effects are minor compared to the more substantial influence of household income and regional differences.
-Life expectancy averages are highest in the Northeast, followed by the Midwest, West, and South.
-Region also plays the most significant role out of all included variables in determining the likelihood of school funding adequacy, with schools in the Northeast and Midwest being far more likely to receive adequate funding than those in the South.
-Other factors, such as residential segregation, property taxes, and household income, show statistically significant but smaller effects.
-These results highlight the complex interplay between socioeconomic variables and geographic location in shaping public well-being.
-By identifying key disparities in health and education, this study underscores the importance of equitable resource allocation and targeted policies to address regional and socioeconomic inequalities.
-
-## **Introduction**
-
-Life expectancy and school funding are important indicators of public well-being because they capture fundamental aspects of health, education, and social equity that are essential for the development of individuals and communities.
-Life expectancy is an important health indicator that reflects not only the overall health of a society, but also its stability, economic resilience, and levels of inequality.
-While access to quality healthcare and nutritious foods are drivers of well-being, where you live and the nature of the community you live in likely also have an influence on health outcomes such as life expectancy.
-According to "The Root Causes of Health Inequity - Communities in Action" report from the National Institutes of Health, we can infer that differences in life expectancy between regions and socioeconomic groups often reveal systemic inequalities, with marginalized communities experiencing shorter life spans due to limited access to health care, healthy living environments, and economic opportunities.^1^ When it comes to educational indicators, adequately funded schools ensure that students have equal access to learning opportunities, serving as a crucial investment in social mobility.
-The Stanford Center for Opportunity Policy in Education highlights that "the unequal allocation and inadequate levels of resources in schools and communities \[are\] at the heart of many gaps in student opportunity."^2^ This emphasizes the importance of equitable and adequate school funding in closing opportunity gaps and empowering students from disadvantaged backgrounds to achieve academic and economic success.
-By better understanding the correlates of school funding adequacy, we can take a major step towards implementing policy that targets the areas and communities most affected by this issue.
+Life expectancy and school funding are important indicators of public
+well-being because they capture fundamental aspects of health,
+education, and social equity that are essential for the development of
+individuals and communities. Life expectancy is an important health
+indicator that reflects not only the overall health of a society, but
+also its stability, economic resilience, and levels of inequality. While
+access to quality healthcare and nutritious foods are drivers of
+well-being, where you live and the nature of the community you live in
+likely also have an influence on health outcomes such as life
+expectancy. According to ``The Root Causes of Health Inequity -
+Communities in Action'' report from the National Institutes of Health,
+we can infer that differences in life expectancy between regions and
+socioeconomic groups often reveal systemic inequalities, with
+marginalized communities experiencing shorter life spans due to limited
+access to health care, healthy living environments, and economic
+opportunities.\textsuperscript{1}~When it comes to educational
+indicators, adequately funded schools ensure that students have equal
+access to learning opportunities, serving as a crucial investment in
+social mobility. The Stanford Center for Opportunity Policy in Education
+highlights that ``the unequal allocation and inadequate levels of
+resources in schools and communities {[}are{]} at the heart of many gaps
+in student opportunity.''\textsuperscript{2} This emphasizes the
+importance of equitable and adequate school funding in closing
+opportunity gaps and empowering students from disadvantaged backgrounds
+to achieve academic and economic success. By better understanding the
+correlates of school funding adequacy, we can take a major step towards
+implementing policy that targets the areas and communities most affected
+by this issue.
 
 For this purpose, we focus our analysis on two key research questions.
-First, we investigate how socioeconomic factors impact average life expectancy.
-**We seek to answer the question, "What is the impact of residential segregation, access to healthy foods, and health insurance coverage on a county's average life expectancy?".** This analysis includes a comparison across different regions to identify variations in the relationship between residential segregation and life expectancy.
-Second, **we seek to answer the question, "Do residential segregation and region impact the likelihood of a school being adequately funded?"**.
-This question allows us to assess how the racial makeup and geographic location—given that each region of the U.S.
-is characterized by differing political and cultural norms—affect a county's educational investment.
-The observational unit for this analysis is the U.S.
-county.
+First, we investigate how socioeconomic factors impact average life
+expectancy. \textbf{We seek to answer the question, ``What is the impact
+of residential segregation, access to healthy foods, and health
+insurance coverage on a county's average life expectancy?''.} This
+analysis includes a comparison across different regions to identify
+variations in the relationship between residential segregation and life
+expectancy. Second, \textbf{we seek to answer the question, ``Do
+residential segregation and region impact the likelihood of a school
+being adequately funded?''}. This question allows us to assess how the
+racial makeup and geographic location---given that each region of the
+U.S. is characterized by differing political and cultural norms---affect
+a county's educational investment. The observational unit for this
+analysis is the U.S. county.
 
-To obtain data on the aforementioned variables, we used three primary sources.
+To obtain data on the aforementioned variables, we used three primary
+sources.~
 
-The first source was the County Health Rankings & Roadmaps program,^3^ associated with the University of Wisconsin Population Health Institute, which provided us with 2024 datasets for all U.S.
-states.
-This program consolidates the latest county-level measurements of various population health, economic, demographic, and social factors from the American Community Survey, the USDA Food Environment Atlas, Small Area Income and Poverty Estimates, and other reliable government sources into publicly available datasets every year.
+The first source was the County Health Rankings \& Roadmaps
+program,\textsuperscript{3} associated with the University of Wisconsin
+Population Health Institute, which provided us with 2024 datasets for
+all U.S. states. This program consolidates the latest county-level
+measurements of various population health, economic, demographic, and
+social factors from the American Community Survey, the USDA Food
+Environment Atlas, Small Area Income and Poverty Estimates, and other
+reliable government sources into publicly available datasets every year.
 Our specific variables of interest were originally collected as part of:
 
--   National Center for Health Statistics - Natality & Mortality Files, 2019-2021 (average life expectancy)
+\begin{itemize}
+\item
+  National Center for Health Statistics - Natality \& Mortality Files,
+  2019-2021 (average life expectancy)
+\item
+  American Community Survey 5-year estimates, 2018-2022 (residential
+  segregation index \& median household income)
+\item
+  USDA Food Environment Atlas, 2019 (percentage of population with
+  limited access to healthy foods)
+\item
+  Small Area Health Insurance Estimates, 2021 (percentage of adults
+  uninsured)
+\item
+  School Finance Indicators Database, 2021 (school funding adequacy)
+\end{itemize}
 
--   American Community Survey 5-year estimates, 2018-2022 (residential segregation index & median household income)
+The second source was IPUMS, which provided us with median poverty tax
+data for every county from 2018-2022.\textsuperscript{4}
 
--   USDA Food Environment Atlas, 2019 (percentage of population with limited access to healthy foods)
+The third source was the U.S. Census Bureau's Regions and Divisions of
+the United States, which maps every U.S. to one of four geographic
+regions, and also a division within the regions. We specifically
+included regions as a variable in this analysis.~This data was loaded
+into R using a
+\href{https://github.com/cphalpert/census-regions}{publicly available
+GitHub repository} titled \emph{Census Regions}.\textsuperscript{5}
 
--   Small Area Health Insurance Estimates, 2021 (percentage of adults uninsured)
+\subsection{\texorpdfstring{\textbf{Methods}}{Methods}}\label{methods}
 
--   School Finance Indicators Database, 2021 (school funding adequacy)
+\subsubsection{Variable Selection}\label{variable-selection}
 
-The second source was IPUMS, which provided us with median poverty tax data for every county from 2018-2022.^4^
+Variables of interest were selected a priori based on their perceived
+relevance to the two outcomes we wanted to better understand: average
+life expectancy and school funding adequacy. While nutrition and access
+to quality healthcare are inherently linked to general wellbeing, it is
+worthwhile to assess how influential the environmental and financial
+aspects of these factors are on our tangible outcome. Therefore, we
+selected two key predictors for our life expectancy model: the
+percentage of the population with limited access to healthy foods and
+the percentage of uninsured adults in a county. The former is defined by
+the US Department of Agriculture as ``the percentage of the {[}county{]}
+population that is low-income and does not live close to a grocery
+store'', where ``close'' is defined differently for rural and nonrural
+areas.\textsuperscript{6} The latter variable represents the percentage
+of adults in a county younger than 65 who lack health insurance. We also
+were interested in seeing how residential segregation would impact life
+expectancy among these predictors, given existing studies on the
+negative health effects of segregation, and whether this relationship
+would exhibit any variations across different regions. This is why we
+chose to include residential segregation index values as a predictor.
+The residential segregation index, used by the County Health Rankings \&
+Roadmaps program, measures how evenly two groups, such as Black and
+white residents, are distributed across geographic areas, with values
+ranging from 0 (complete integration) to 100 (complete
+segregation).\textsuperscript{7}~
 
-The third source was the U.S.
-Census Bureau’s Regions and Divisions of the United States, which maps every U.S.
-to one of four geographic regions, and also a division within the regions.
-We specifically included regions as a variable in this analysis. This data was loaded into R using a [publicly available GitHub repository](https://github.com/cphalpert/census-regions) titled *Census Regions*.^5^
+Past studies have established a correlation between income and life
+expectancy.\textsuperscript{8} Because we wanted to control for its
+effects and examine how insurance, healthy food access, segregation, and
+region influence life expectancy even within a given income bracket, we
+selected income as our final predictor for the first model.
 
-## **Methods**
+School funding adequacy is defined by the County Health Rankings program
+as ``the average dollar gap between actual per-pupil spending and the
+amount needed for students to achieve national average test scores,
+considering districts' varying equity-based needs.''\textsuperscript{9}
+To explore how racial diversity interface with variations in school
+funding, we selected residential segregation to be a part of this
+regression. In addition, given that state policy has a large influence
+on school administration and the determination of school needs, and that
+states in the same region tend to have similar political leanings and
+cultural beliefs, we wanted to analyze how much of a relationship exists
+between region and actual school funding levels. Because a major source
+of public school funding is tax revenue, median property tax and median
+household income were also included in the school funding model as
+potential confounders. This allowed us to assess whether the effects of
+segregation and region on funding existed outside of the effects of
+known economic factors.
 
-### Variable Selection
+\subsubsection{Exploratory Data Analysis \& Data
+Preprocessing}\label{exploratory-data-analysis-data-preprocessing}
 
-Variables of interest were selected a priori based on their perceived relevance to the two outcomes we wanted to better understand: average life expectancy and school funding adequacy.
-While nutrition and access to quality healthcare are inherently linked to general wellbeing, it is worthwhile to assess how influential the environmental and financial aspects of these factors are on our tangible outcome.
-Therefore, we selected two key predictors for our life expectancy model: the percentage of the population with limited access to healthy foods and the percentage of uninsured adults in a county.
-The former is defined by the US Department of Agriculture as "the percentage of the \[county\] population that is low-income and does not live close to a grocery store", where "close" is defined differently for rural and nonrural areas.^6^ The latter variable represents the percentage of adults in a county younger than 65 who lack health insurance.
-We also were interested in seeing how residential segregation would impact life expectancy among these predictors, given existing studies on the negative health effects of segregation, and whether this relationship would exhibit any variations across different regions.
-This is why we chose to include residential segregation index values as a predictor.
-The residential segregation index, used by the County Health Rankings & Roadmaps program, measures how evenly two groups, such as Black and white residents, are distributed across geographic areas, with values ranging from 0 (complete integration) to 100 (complete segregation).^7^
+2024 datasets for all U.S. states were combined and merged with median
+property tax and region mapping data, resulting in a final dataset with
+3144 county observations and 9 different variables (state, average life
+expectancy, residential segregation index, \% limited access to healthy
+foods, \% uninsured adults, school funding adequacy, median household
+income, median property tax, and region). Because school funding
+adequacy was given as a numerical value, and we wanted to focus on
+better understanding the dichotomy between school funding adequacy and
+inadequacy, this variable was converted into a binary categorical
+variable (where the value was set to 1 if the numerical value was
+positive, indicating the counties' schools were adequately funded, and
+set to 0 if the numerical value was negative, indicating underfunding).
 
-Past studies have established a correlation between income and life expectancy.^8^ Because we wanted to control for its effects and examine how insurance, healthy food access, segregation, and region influence life expectancy even within a given income bracket, we selected income as our final predictor for the first model.
+Exploratory data analysis for the two outcome variables was conducted by
+1) calculating descriptive statistics, 2) creating scatterplots of each
+individual predictor against the outcome (for life expectancy), and 3)
+creating pairwise plots between variables (for school funding). During
+the initial stages of analysis, we found that ⅓ of our observations were
+missing residential segregation index data, and that certain states were
+missing nearly all of their segregation values. Because of this, we
+decided to drop all states missing over 50\% of their values for any
+variable. This led to our analysis being limited to 33 states, and
+excluding the following: Alaska, Colorado, Idaho, Iowa, Kansas,
+Minnesota, Montana, Nebraska, North Dakota, South Dakota, Utah, Wyoming,
+and Vermont. We were left with 2379 total observations, and the missing
+percentage for residential segregation dropped to 22\%.
 
-School funding adequacy is defined by the County Health Rankings program as "the average dollar gap between actual per-pupil spending and the amount needed for students to achieve national average test scores, considering districts' varying equity-based needs."^9^ To explore how racial diversity interface with variations in school funding, we selected residential segregation to be a part of this regression.
-In addition, given that state policy has a large influence on school administration and the determination of school needs, and that states in the same region tend to have similar political leanings and cultural beliefs, we wanted to analyze how much of a relationship exists between region and actual school funding levels.
-Because a major source of public school funding is tax revenue, median property tax and median household income were also included in the school funding model as potential confounders.
-This allowed us to assess whether the effects of segregation and region on funding existed outside of the effects of known economic factors.
+After this, the remaining missing values in all variables were imputed
+using the mice package and predictive mean matching to generate one
+complete dataset. The original school funding adequacy numerical
+variable was first imputed, and then converted into binary values for
+subsequent analysis.
 
-### Exploratory Data Analysis & Data Preprocessing
+\subsubsection{Model Fitting \&
+Assessment}\label{model-fitting-assessment}
 
-2024 datasets for all U.S.
-states were combined and merged with median property tax and region mapping data, resulting in a final dataset with 3144 county observations and 9 different variables (state, average life expectancy, residential segregation index, % limited access to healthy foods, % uninsured adults, school funding adequacy, median household income, median property tax, and region).
-Because school funding adequacy was given as a numerical value, and we wanted to focus on better understanding the dichotomy between school funding adequacy and inadequacy, this variable was converted into a binary categorical variable (where the value was set to 1 if the numerical value was positive, indicating the counties’ schools were adequately funded, and set to 0 if the numerical value was negative, indicating underfunding).
+For our first research question, a multiple linear regression model was
+fit, regressing average life expectancy on residential segregation
+index, percentage with limited access to healthy foods, percentage
+uninsured adults, region, and median household income. Linear regression
+assumptions were assessed using diagnostic plots, especially residual
+vs.~fitted and quantile-quantile plots. The adjusted R-squared value was
+used to evaluate the fit of the model, and the effect of region as an
+interaction term with residential segregation was evaluated using nested
+F tests. We also calculated VIF to check for multicollinearity and
+Cook's distance to identify influential points.~
 
-Exploratory data analysis for the two outcome variables was conducted by 1) calculating descriptive statistics, 2) creating scatterplots of each individual predictor against the outcome (for life expectancy), and 3) creating pairwise plots between variables (for school funding).
-During the initial stages of analysis, we found that ⅓ of our observations were missing residential segregation index data, and that certain states were missing nearly all of their segregation values.
-Because of this, we decided to drop all states missing over 50% of their values for any variable.
-This led to our analysis being limited to 33 states, and excluding the following: Alaska, Colorado, Idaho, Iowa, Kansas, Minnesota, Montana, Nebraska, North Dakota, South Dakota, Utah, Wyoming, and Vermont.
-We were left with 2379 total observations, and the missing percentage for residential segregation dropped to 22%.
+For our second research question, a binary logistic regression model was
+fit with school funding adequacy (categorical) as the outcome and
+residential segregation index, region, median property tax, and median
+household income as predictors. The model was assessed using a confusion
+matrix, an ROC curve, and a comparison of deviance between the full
+model and a reduced model that excluded region as a predictor. VIF and
+Cook's distance were again used to evaluate multicollinearity and
+influential points.
 
-After this, the remaining missing values in all variables were imputed using the mice package and predictive mean matching to generate one complete dataset.
-The original school funding adequacy numerical variable was first imputed, and then converted into binary values for subsequent analysis.
+\subsection{Results}\label{results}
 
-### Model Fitting & Assessment
+\subsubsection{\texorpdfstring{\emph{Data
+Overview}}{Data Overview}}\label{data-overview}
 
-For our first research question, a multiple linear regression model was fit, regressing average life expectancy on residential segregation index, percentage with limited access to healthy foods, percentage uninsured adults, region, and median household income.
-Linear regression assumptions were assessed using diagnostic plots, especially residual vs.
-fitted and quantile-quantile plots.
-The adjusted R-squared value was used to evaluate the fit of the model, and the effect of region as an interaction term with residential segregation was evaluated using nested F tests.
-We also calculated VIF to check for multicollinearity and Cook’s distance to identify influential points.
-
-For our second research question, a binary logistic regression model was fit with school funding adequacy (categorical) as the outcome and residential segregation index, region, median property tax, and median household income as predictors.
-The model was assessed using a confusion matrix, an ROC curve, and a comparison of deviance between the full model and a reduced model that excluded region as a predictor.
-VIF and Cook's distance were again used to evaluate multicollinearity and influential points.
-
-## Results
-
-### *Data Overview*
-
-Table 1 displays summary statistics for the predictors in the second model, stratified by school funding adequacy.
-Table A1 in the Appendix provides summary statistics for all variables, and shows that both the mean and median for life expectancy are approximately 75 years.
-School funding adequacy counts show that around 66% of the counties in our analysis had underfunded schools.
-Additionally, the number of counties per region indicate an imbalance between categories, with the South region being over represented (holding 60% of all counties), and the East and Northeast holding around 9% of the counties each.
-
-```{r, echo=FALSE, include=FALSE}
-completed_data |> group_by(School.Funding.Cat) |> summarise(mean(Median.Household.Income), sd(Median.Household.Income), mean(Median.Prop.Tax), sd(Median.Prop.Tax), mean(Residential.Segregation.Index), sd(Residential.Segregation.Index))
-```
-
-```{r, echo=FALSE, include=FALSE}
-table(completed_data$School.Funding.Cat, completed_data$Region_fac)
-```
+Table 1 displays summary statistics for the predictors in the second
+model, stratified by school funding adequacy. Table A1 in the Appendix
+provides summary statistics for all variables, and shows that both the
+mean and median for life expectancy are approximately 75 years. School
+funding adequacy counts show that around 66\% of the counties in our
+analysis had underfunded schools. Additionally, the number of counties
+per region indicate an imbalance between categories, with the South
+region being over represented (holding 60\% of all counties), and the
+East and Northeast holding around 9\% of the counties each.
 
@@
 
-*Table 1:* *School Funding Adequacy Summary Statistics*
+\emph{Table 1:} \emph{School Funding Adequacy Summary Statistics}
 
-| Variable                                 | **Inadequate School Funding** | **Adequate School Funding** |
-|------------------------------|----------------------|--------------------|
-| Residential Segregation Index: Mean (SD) | 46 (17)                       | 56 (14)                     |
-| Region-West: Count (%)                   | 114 (7.3%)                    | 89 (10.8%)                  |
-| Region-Northeast: Count (%)              | 3 (0.2%)                      | 200 (24.3%)                 |
-| Region-Midwest: Count (%)                | 267 (17.1%)                   | 285 (34.7%)                 |
-| Region-South: Count (%)                  | 1173 (75.3%)                  | 248 (30.2%)                 |
-| Median Property Tax, USD: Mean (SD)      | 1119 (763)                    | 2282 (1648)                 |
-| Median Household Income, USD: Mean (SD)  | 56513 (13200)                 | 72218 (18255)               |
+\begin{longtable}[]{@{}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4167}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3056}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2778}}@{}}
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\raggedright
+Variable
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+\textbf{Inadequate School Funding}
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+\textbf{Adequate School Funding}
+\end{minipage} \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+Residential Segregation Index: Mean (SD) & 46 (17) & 56 (14) \\
+Region-West: Count (\%) & 114 (7.3\%) & 89 (10.8\%) \\
+Region-Northeast: Count (\%) & 3 (0.2\%) & 200 (24.3\%) \\
+Region-Midwest: Count (\%) & 267 (17.1\%) & 285 (34.7\%) \\
+Region-South: Count (\%) & 1173 (75.3\%) & 248 (30.2\%) \\
+Median Property Tax, USD: Mean (SD) & 1119 (763) & 2282 (1648) \\
+Median Household Income, USD: Mean (SD) & 56513 (13200) & 72218
+(18255) \\
+\end{longtable}
 
-```{r, echo=FALSE, include=FALSE}
-# Plot for Q2 EDA
-ggpairs(
-  completed_data,
-  columns = c("Residential.Segregation.Index", "Median.Household.Income", "Median.Prop.Tax"),
-  aes(color = factor(School.Funding.Cat,
-                     labels = c("Inadequate", "Adequate")))
-) +
-  labs(
-    title = "Pairwise Plots of Predictors Colored by School Funding Adequacy",
-    color = "School Funding Adequacy"
-  ) +
-  theme(
-    plot.title = element_text(size = 15, hjust = 0.5),
-    strip.text.x = element_text(size = 5),
-    strip.text.y = element_text(size = 3)
-  )
+\subsubsection{Question 1}\label{question-1}
 
-```
+\paragraph{Model Results}\label{model-results}
 
-### Question 1
+Based on the multiple linear regression results, residential
+segregation, percentage of uninsured adults, region, and median
+household income all showed statistically significant effects on life
+expectancy. Specifically, for every \$10,000 increase in median
+household income, life expectancy increases by 1.25 years on average. In
+addition, the impact of region on life expectancy was substantial: life
+expectancy is on average 1.96 years higher in the West region compared
+to the South, 2.92 years higher in the Northeast compared to the South,
+and 1.92 years higher in the Midwest compared to the South, even when
+controlling for income, uninsured population, and racial diversity. This
+suggests, income as a factor withstanding, the differing quality of and
+access to healthcare and public services found in each region can have a
+significant impact on life expectancy.
 
-#### Model Results
+Surprisingly, holding all other variables constant, for every 1 unit
+increase in percentage of uninsured adults in a county, a county's
+average life expectancy increases by 0.032 years. While statistically
+significant, many studies covering the relationship between an
+individual's health and having insurance coverage have proven its
+positive correlation, suggesting our results may benefit from a closer
+look. In assessing our dataset, as seen in Appendix 1A, we find that our
+dataset contains more observations for counties with low uninsured
+individuals, or equivalently, counties where most people are covered. As
+the timeframe of our dataset is 2018-2022, it is likely that the
+aforementioned observations require greater scrutiny of the significance
+of the impact of this variable on average life expectancy of a county.~
 
-Based on the multiple linear regression results, residential segregation, percentage of uninsured adults, region, and median household income all showed statistically significant effects on life expectancy.
-Specifically, for every \$10,000 increase in median household income, life expectancy increases by 1.25 years on average.
-In addition, the impact of region on life expectancy was substantial: life expectancy is on average 1.96 years higher in the West region compared to the South, 2.92 years higher in the Northeast compared to the South, and 1.92 years higher in the Midwest compared to the South, even when controlling for income, uninsured population, and racial diversity.
-This suggests, income as a factor withstanding, the differing quality of and access to healthcare and public services found in each region can have a significant impact on life expectancy.
+The percentage of population with limited access to healthy foods
+predictor was not statistically significant in our model, suggesting
+that the effect of this environmental factor is minor or intertwined
+with economic variables, such as income, rather than being a direct or
+standalone influence. In terms of residential segregation, for every 1
+unit increase in a county's index value (corresponding to increased
+segregation), its average life expectancy increases by 0.0071 years in
+the South, increases by 0.0068 years in the West, decreases by 0.0055
+years in the Northeast, and decreases by 0.0004 years in the Midwest, as
+seen in Figure 1. However, these regional differences are not
+statistically significant at the 0.05 significance level given the high
+p-values of all residential segregation-region interaction terms. We
+confirmed this result with a nested F-test, comparing a full model to a
+reduced model that excluded the interaction term between region and
+residential segregation. This did not show a statistically significant
+difference in RSS (p=0.624), indicating that the interaction between
+region and residential segregation does not add meaningful explanatory
+power to our model, Consequently, we did not find sufficient evidence to
+support regional variations in the relationship between segregation and
+life expectancy. Instead, based on residential segregation's coefficient
+estimate in the model, life expectancy does appear to increase overall
+with an increase in residential segregation, however minor this increase
+may be. One possible theory for this observation is that counties with
+more racial uniformity might experience less racial conflict and more
+social cohesion, especially for minority communities.
 
-Surprisingly, holding all other variables constant, for every 1 unit increase in percentage of uninsured adults in a county, a county's average life expectancy increases by 0.032 years.
-While statistically significant, many studies covering the relationship between an individual’s health and having insurance coverage have proven its positive correlation, suggesting our results may benefit from a closer look.
-In assessing our dataset, as seen in Appendix 1A, we find that our dataset contains more observations for counties with low uninsured individuals, or equivalently, counties where most people are covered.
-As the timeframe of our dataset is 2018-2022, it is likely that the aforementioned observations require greater scrutiny of the significance of the impact of this variable on average life expectancy of a county.
+\emph{Figure 1: Segregation vs.~Life Expectancy Across Different U.S.
+Regions}
 
-The percentage of population with limited access to healthy foods predictor was not statistically significant in our model, suggesting that the effect of this environmental factor is minor or intertwined with economic variables, such as income, rather than being a direct or standalone influence.
-In terms of residential segregation, for every 1 unit increase in a county's index value (corresponding to increased segregation), its average life expectancy increases by 0.0071 years in the South, increases by 0.0068 years in the West, decreases by 0.0055 years in the Northeast, and decreases by 0.0004 years in the Midwest, as seen in Figure 1.
-However, these regional differences are not statistically significant at the 0.05 significance level given the high p-values of all residential segregation-region interaction terms.
-We confirmed this result with a nested F-test, comparing a full model to a reduced model that excluded the interaction term between region and residential segregation.
-This did not show a statistically significant difference in RSS (p=0.624), indicating that the interaction between region and residential segregation does not add meaningful explanatory power to our model, Consequently, we did not find sufficient evidence to support regional variations in the relationship between segregation and life expectancy.
-Instead, based on residential segregation's coefficient estimate in the model, life expectancy does appear to increase overall with an increase in residential segregation, however minor this increase may be.
-One possible theory for this observation is that counties with more racial uniformity might experience less racial conflict and more social cohesion, especially for minority communities.
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-10-1.pdf}
 
-*Figure 1: Segregation vs. Life Expectancy Across Different U.S. Regions*
+\emph{Table 2: MLR Results For Average Life Expectancy Outcome}
 
-```{r, echo=FALSE, fig.pos='H'}
-library(ggplot2)
-library(GGally)
-library(patchwork)
+\begin{table}
+\centering
+\begin{tblr}[         %% tabularray outer open
+]                     %% tabularray outer close
+{                     %% tabularray inner open
+colspec={Q[]Q[]Q[]Q[]Q[]Q[]},
+cell{1}{2}={c=5,}{halign=c,},
+column{1}={halign=l,},
+column{2}={halign=c,},
+column{3}={halign=c,},
+column{4}={halign=c,},
+column{5}={halign=c,},
+column{6}={halign=c,},
+row{1}={halign=c,},
+}                     %% tabularray inner close
+\toprule
+& (1) &  &  &  &  \\ \cmidrule[lr]{2-6}
+& Est. & S.E. & 2.5 \% & 97.5 \% & p \\ \midrule %% TinyTableHeader
+(Intercept)                             & \num{65.88}    & \num{0.27}    & \num{65.34}    & \num{66.41}   & \num{<0.01} \\
+Residential Segregation Index (RSI)     & \num{0.00715}  & \num{0.00332} & \num{0.00065}  & \num{0.01366} & \num{0.03}  \\
+Region-West                             & \num{1.96}     & \num{0.69}    & \num{0.60}     & \num{3.32}    & \num{<0.01} \\
+Region-Northeast                        & \num{2.92}     & \num{0.84}    & \num{1.27}     & \num{4.57}    & \num{<0.01} \\
+Region-Midwest                          & \num{1.91}     & \num{0.44}    & \num{1.06}     & \num{2.77}    & \num{<0.01} \\
+Percentage With Access to Healthy Foods & \num{0.0040}   & \num{0.0062}  & \num{-0.0082}  & \num{0.0162}  & \num{0.52}  \\
+Percentage of Uninsured Adults          & \num{0.0328}   & \num{0.0075}  & \num{0.0181}   & \num{0.0476}  & \num{<0.01} \\
+Median.Household.Income                 & \num{1.3e-04}  & \num{2.7e-06} & \num{1.2e-04}  & \num{1.3e-04} & \num{<0.01} \\
+RSI*Region-West                         & \num{-0.00032} & \num{0.01164} & \num{-0.02313} & \num{0.02250} & \num{0.98}  \\
+RSI*Region-Northeast                    & \num{-0.013}   & \num{0.014}   & \num{-0.040}   & \num{0.015}   & \num{0.36}  \\
+RSI*Region-Midwest                      & \num{-0.0075}  & \num{0.0075}  & \num{-0.0222}  & \num{0.0072}  & \num{0.32}  \\
+\bottomrule
+\end{tblr}
+\end{table}
 
-ggplot(completed_data, aes(x = Residential.Segregation.Index, y = Life.Expectancy, color=Region_fac)) +
-  geom_point(color='black') + geom_smooth(method="lm",se=F) + 
-  scale_color_brewer(palette = "Dark2") +
-  labs(x = "Residential Segregation Index", y = "Average Life Expectancy (Years)", color='U.S. Region', title='Relationship Between Residential Segregation and Life Expectancy')+ 
-  theme(plot.margin = margin(t = 20, r = 30, b = 20, l = 10),
-axis.title.x = element_text(size = 8),
-axis.title.y = element_text(size = 8),
-title=element_text(size=9))
-```
+\paragraph{Model Assessment}\label{model-assessment}
 
-*Table 2: MLR Results For Average Life Expectancy Outcome*
+The adjusted R-squared value for the model was 0.6065, indicating that a
+majority of the variability in life expectancy was explained by the
+socioeconomic and racial predictors included in the model. In conducting
+diagnostics on the multilinear regression model, we first checked to see
+if the model met assumptions of linearity, independence, normality, and
+homoscedasticity. The residuals vs.~fitted plot showed relatively equal
+variance and no discernible pattern, meaning that the linearity and
+homoscedasticity assumptions were reasonably met. In addition, because
+each observation in the dataset represented a unique county in the U.S.,
+independence could be reasonably assumed. Assessing the Q-Q Plot for
+normality, the residual distribution was approximately normal, despite a
+slight divergence on both tails of the plot. Cook's distance was also
+assessed using the residuals vs.~leverage plot; no points were found to
+be influential. There was a high leverage point (Mineral County,
+Nevada), and refitting the model without this point yielded no
+substantial changes in the coefficient estimates for statistically
+significant predictors; however, the p-value for residential segregation
+increased from 0.031 to 0.054. Finally, with a VIF \textless2 for all
+predictors, there was little to no multicollinearity among the
+predictors.
 
-```{r, echo= FALSE, warning = FALSE, fig.pos='H'}
-# Running a regression with the imputated data (completed_data) to understand the impact of missing data
+\subsection{Question 2}\label{question-2}
 
-mlr.imp <- lm(Life.Expectancy ~ Residential.Segregation.Index * Region_fac +
-    X..Limited.Access.to.Healthy.Foods +
-    X..Uninsured.Adults + Median.Household.Income, data=completed_data)
+\paragraph{Model Results}\label{model-results-1}
 
-renamed <- c(
-    "Residential.Segregation.Index" = "Residential Segregation Index (RSI)",
-  "X..Limited.Access.to.Healthy.Foods" = "Percentage With Access to Healthy Foods",
-  "X..Uninsured.Adults"= "Percentage of Uninsured Adults",
-  "Region_facWest"="Region-West", 
-  "Region_facNortheast"="Region-Northeast",
-  "Region_facMidwest"="Region-Midwest" ,
-  "Residential.Segregation.Index:Region_facWest" = "RSI*Region-West",
-  "Residential.Segregation.Index:Region_facNortheast"= "RSI*Region-Northeast",
-  "Residential.Segregation.Index:Region_facMidwest"="RSI*Region-Midwest",
-  "Median Household Income" = "Median.Household.Income")
+Table 3 below displays the odds ratio results from the logistic
+regression model for school funding adequacy, with the South as the
+baseline category for the categorical region variable.~While all
+estimates are statistically significant at the 0.05 significance level,
+exponentiated regression coefficient estimates from the logistic model
+demonstrate that the level of residential segregation, the median
+household income, and median property tax of the county have very
+minimal impacts on the odds of a school being adequately funded. This
+finding is interesting as looking at Figure 2, we see that across
+regions, the average racial segregation index is consistent across
+school funding adequacy categories.~
 
-modelsummary(mlr.imp, fmt = fmt_significant(2),
-             coef_rename = renamed, 
-             shape = term ~ model + statistic,
-             statistic = c("std.error","conf.int","p.value"),
-             exponentiate = FALSE,
-             gof_map=NA,
-             )
-```
+Region, however, has a substantial impact on the odds of school funding
+adequacy.~ Being located in the Northeast or the Midwest has the
+greatest impact on how well the school is funded, with the odds of
+adequate school funding being over 200 times higher in a Northeast
+county compared to a Southern county, and the odds of adequate school
+funding being over 3 times higher in a Northeast county compared to a
+Southern county, holding all other variables constant. Such disparities
+across the country are caused by a number of factors, including: (1)
+capacity - how well off a state is based on its economy and resources,
+and (2) effort - the states willingness to provide funding for
+education. Wealthier states with a high fiscal capacity, (typically
+those in the Northeast), have more funding available to spend on
+education than states with more limited resources (typically those in
+the South and the West).~
 
-```{r echo= FALSE, warning = FALSE, include=FALSE}
-# Running a regression without the interaction term between residential segregation and region
+In comparing our full model (with predictor variables for residential
+segregation, median household income, median property tax, and region)
+to a model including all of these variables except for region, we found
+that there is statistical significance in the deviance between the
+residuals of the full model and reduced. As such, we can conclude that
+including region as a predictor refines and creates a better fit for our
+model.
 
-mlr.imp.reduced <- lm(Life.Expectancy ~ Residential.Segregation.Index +
-    X..Limited.Access.to.Healthy.Foods +
-    X..Uninsured.Adults + Median.Household.Income +
-    Region_fac, data=completed_data)
+\emph{Figure 2: Relationship Between Residential Segregation, Region, \&
+School Funding Adequacy}
 
-renamed <- c(
-   "Residential Segregation Index (RSI)" = "Residential.Segregation.Index",
-  "Percentage With Limited Access to Healthy Foods" = "X..Limited.Access.to.Healthy.Foods",
-  "Percentage of Uninsured Adults"= "X..Uninsured.Adults",
-  "Region-West" = "Region_facWest", 
-  "Region-Northeast"= "Region_facNortheast",
-  "Region-Midwest"= "Region_facMidwest",
-  "Median Household Income" = "Median.Household.Income")
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-22-1.pdf}
 
-modelsummary(mlr.imp.reduced, fmt = fmt_significant(2),
-             coef_rename = renamed, 
-             shape = term ~ model + statistic,
-             statistic = c("std.error","conf.int","p.value"),
-             exponentiate = FALSE,
-             gof_map=NA,
-             )
+\emph{Table 3: Logistic Regression Results For School Funding Adequacy
+Outcome}
 
-```
+\begin{table}
+\centering
+\begin{tblr}[         %% tabularray outer open
+]                     %% tabularray outer close
+{                     %% tabularray inner open
+colspec={Q[]Q[]Q[]Q[]Q[]Q[]},
+cell{1}{2}={c=5,}{halign=c,},
+column{1}={halign=l,},
+column{2}={halign=c,},
+column{3}={halign=c,},
+column{4}={halign=c,},
+column{5}={halign=c,},
+column{6}={halign=c,},
+row{1}={halign=c,},
+}                     %% tabularray inner close
+\toprule
+& (1) &  &  &  &  \\ \cmidrule[lr]{2-6}
+& Est. & S.E. & 2.5 \% & 97.5 \% & p \\ \midrule %% TinyTableHeader
+(Intercept)                         & \num{0.00135} & \num{0.00050} & \num{0.00064} & \num{0.00273} & \num{<0.01} \\
+Residential Segregation Index (RSI) & \num{1.0196}  & \num{0.0041}  & \num{1.0117}  & \num{1.0277}  & \num{<0.01} \\
+Median Household Income, USD        & \num{1.0e+00} & \num{5.5e-06} & \num{1.0e+00} & \num{1.0e+00} & \num{<0.01} \\
+Median Property Tax, USD            & \num{1.0e+00} & \num{8.5e-05} & \num{1.0e+00} & \num{1.0e+00} & \num{0.01}  \\
+Region-West                         & \num{1.83}    & \num{0.35}    & \num{1.26}    & \num{2.66}    & \num{<0.01} \\
+Region-Northeast                    & \num{232}     & \num{141}     & \num{82}      & \num{974}     & \num{<0.01} \\
+Region-Midwest                      & \num{3.84}    & \num{0.52}    & \num{2.95}    & \num{5.02}    & \num{<0.01} \\
+\bottomrule
+\end{tblr}
+\end{table}
 
-```{r, include=FALSE, echo=FALSE}
-##Nested F-Test to see if segregation-region interaction term makes a difference
-anova(mlr.imp, mlr.imp.reduced)
-```
+\paragraph{Model Assessment}\label{model-assessment-1}
 
-#### Model Assessment
+To assess logistic regression assumptions, we first plotted each
+numerical predictor against the log odds of school funding adequacy. All
+plots showed a mostly linear relationship, indicating that the linearity
+assumption was largely met; interestingly, different linear trends were
+observed within each plot based on region, pointing to potential
+interaction effects between region and the three numerical predictors.
+GVIF for all predictors ranged between 1 and 3, suggesting that there
+could be moderate multicollinearity between predictors, but not enough
+to warrant a reconsideration of included variables. Cook's distance was
+also less than 0.05 for all points, meaning that there were no highly
+influential observations in the model. Evaluating our model, we found
+that the logistic regression model demonstrated a strong overall
+performance with an accuracy rate of 80.33\%. The model had a high
+sensitivity of 91\%, indicating that it accurately predicted 91\% of
+actual positive cases of adequate school funding when evaluated on our
+dataset. However, it had a 60\% specificity rate, highlighting it
+correctly identified cases of inadequate school funding only 60\% of the
+time. Overall, looking at the ROC curve below, the AUC is 0.858,
+indicating a moderately strong discriminate ability.~
 
-The adjusted R-squared value for the model was 0.6065, indicating that a majority of the variability in life expectancy was explained by the socioeconomic and racial predictors included in the model.
-In conducting diagnostics on the multilinear regression model, we first checked to see if the model met assumptions of linearity, independence, normality, and homoscedasticity.
-The residuals vs.
-fitted plot showed relatively equal variance and no discernible pattern, meaning that the linearity and homoscedasticity assumptions were reasonably met.
-In addition, because each observation in the dataset represented a unique county in the U.S., independence could be reasonably assumed.
-Assessing the Q-Q Plot for normality, the residual distribution was approximately normal, despite a slight divergence on both tails of the plot.
-Cook's distance was also assessed using the residuals vs.
-leverage plot; no points were found to be influential.
-There was a high leverage point (Mineral County, Nevada), and refitting the model without this point yielded no substantial changes in the coefficient estimates for statistically significant predictors; however, the p-value for residential segregation increased from 0.031 to 0.054.
-Finally, with a VIF \<2 for all predictors, there was little to no multicollinearity among the predictors.
+\subsection{\texorpdfstring{\textbf{Conclusion}}{Conclusion}}\label{conclusion}
 
-```{r, echo=FALSE, include=FALSE}
-completed_data[1180,]
-completed_data_lev <- completed_data[-1180,]
-```
+In our analysis, we observed the significant impact of the regional
+location of a county on both its average life expectancy (a health
+indicator) and school funding adequacy (an educational indicator)--two
+key components of societal and an individual's well-being. Life
+expectancy is highest in the Northeast, followed by the Midwest, West,
+and South, while school funding adequacy is significantly greater in the
+Northeast and Midwest compared to the South. Our findings highlight
+substantial differences in both outcomes between regions, which are
+often caused by different economic capabilities and different
+allocations of resources among countries. This study provides empirical
+evidence that underscores the importance for equitable social policies
+and targeted investments to address regional disparities and promote
+well-being across all communities.
 
-```{r, echo=FALSE, include=FALSE}
-#looking for outliers 
-outlier_points <- outlierTest(mlr.imp)
-print(outlier_points)
-outlier_indices <- as.numeric(rownames(outlier_points))  # Extract row indices
-print(outlier_indices)  # View the indices to confirm
-```
+Our analysis faced limitations due to missing data, which restricted the
+dataset to 33 states and reduced the generalizability of our findings.
+Additionally, some variables were drawn from different years between
+2018 and 2022, requiring us to assume that trends remained consistent
+over a 1--5 year period and that time was an insignificant factor;
+however, this may not actually be the case.
 
-```{r, echo=FALSE, include=FALSE}
-#Variance Inflaction Factor (VIF)
+Future studies could incorporate additional social variables to enhance
+the analysis and explore causal relationships between region and
+outcomes. For example, integrating a dataset on employment rates and
+school performance could offer deeper insights into how socioeconomic
+factors shape life expectancy and educational funding, complementing
+existing variables in our models. Another direction for future work is
+to delve deeper into understanding the surprising findings regarding the
+effects of uninsured adult percentage and residential segregation on
+life expectancy. Expanding the dataset to include a temporal scope, and
+finding more complete sources of data on variables of interest so that
+no states have to be excluded from the analysis, would also contribute
+to a more comprehensive understanding.
 
-#vif(mlr.imp)
-vif(mlr.imp.reduced)
-```
+\subsection{\texorpdfstring{\textbf{Appendix}}{Appendix}}\label{appendix}
 
-```{r, echo=FALSE, include=FALSE}
-#Generate 4 Diagnostic Plots at once
-par(mfrow = c(2, 2), mar = c(4,4,2,2))
-plot(mlr.imp)
-```
+\emph{Table A1: Summary Statistics For All Variables}
 
-```{r, echo= FALSE, warning = FALSE, include=FALSE}
-# Running a regression with obs 1180 removed
+\subsection{\texorpdfstring{\textbf{References}}{References}}\label{references}
 
-mlr.lev <- lm(Life.Expectancy ~ Residential.Segregation.Index * Region_fac +
-    X..Limited.Access.to.Healthy.Foods +
-    X..Uninsured.Adults + Median.Household.Income, data=completed_data_lev)
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  National Academies of Sciences, Engineering, and Medicine. (2017).
+  \emph{Communities in action: Pathways to health equity.} Washington,
+  DC: The National Academies Press. Retrieved from
+  \url{https://www.ncbi.nlm.nih.gov/books/NBK425845/}
+\item
+  Stanford Center for Opportunity Policy in Education. (2013.). How to
+  close the opportunity gap: Key policy recommendations. Retrieved from
+  \url{https://edpolicy.stanford.edu/sites/default/files/Opp\%20Gap\%20Policy\%20Recommendations.pdf}
+\item
+  County Health Rankings \& Roadmaps. (n.d.). \emph{About us.} Retrieved
+  December 8, 2024, from
+  \url{https://www.countyhealthrankings.org/about-us}
+\item
+  Steven Ruggles, Sarah Flood, Matthew Sobek, Daniel Backman, Annie
+  Chen, Grace Cooper, Stephanie Richards, Renae Rodgers, and Megan
+  Schouweiler. IPUMS USA: Version 15.0 {[}dataset{]}. Minneapolis, MN:
+  IPUMS, 2024. Retrieved from \url{https://doi.org/10.18128/D010.V15.0}
+\item
+  Halpert, C. (2014). Census regions {[}GitHub repository{]}. GitHub.
+  \url{https://github.com/cphalpert/census-regions}
+\item
+  USDA. (2022). \emph{Documentation}. Economic Research Service.
+  https://www.ers.usda.gov/data-products/food-access-research-atlas/documentation/
+\item
+  County Health Rankings \& Roadmaps. (2024). \emph{Residential
+  segregation - Black/White.} University of Wisconsin Population Health
+  Institute. Retrieved from
+  \url{https://www.countyhealthrankings.org/health-data/health-factors/social-economic-factors/family-and-social-support/residential-segregation-blackwhite?utm_source=chatgpt.com&year=2024}
+\item
+  Chetty, R., Stepner, M., Abraham, S., Lin, S., Scuderi, B., Turner,
+  N., Bergeron, A., \& Cutler, D. (2016). The Association Between Income
+  and Life Expectancy in the United States, 2001-2014. \emph{JAMA},
+  \emph{315}(16), 1750--1766. https://doi.org/10.1001/jama.2016.4226
+\item
+  County Health Rankings \& Roadmaps. (2024). \emph{School funding
+  adequacy.} University of Wisconsin Population Health Institute.
+  Retrieved from
+  \url{https://www.countyhealthrankings.org/health-data/health-factors/social-economic-factors/education/school-funding-adequacy?utm_source=chatgpt.com&year=2024}
+\end{enumerate}
 
-renamed <- c(
-    "Residential.Segregation.Index" = "Residential Segregation Index (RSI)",
-  "X..Limited.Access.to.Healthy.Foods" = "Percentage With Access to Healthy Foods",
-  "X..Uninsured.Adults"= "Percentage of Uninsured Adults",
-  "Region_facWest"="Region-West", 
-  "Region_facNortheast"="Region-Northeast",
-  "Region_facMidwest"="Region-Midwest" ,
-  "Residential.Segregation.Index:Region_facWest" = "RSI*Region-West",
-  "Residential.Segregation.Index:Region_facNortheast"= "RSI*Region-Northeast",
-  "Residential.Segregation.Index:Region_facMidwest"="RSI*Region-Midwest",
-  "Median Household Income" = "Median.Household.Income")
-
-modelsummary(mlr.lev, fmt = fmt_significant(2),
-             coef_rename = renamed, 
-             shape = term ~ model + statistic,
-             statistic = c("std.error","conf.int","p.value"),
-             exponentiate = FALSE,
-             gof_map=NA,
-             )
-```
-
-```{r, echo=FALSE, include=FALSE}
-plot(mlr.lev)
-```
-
-## Question 2
-
-#### Model Results
-
-Table 3 below displays the odds ratio results from the logistic regression model for school funding adequacy, with the South as the baseline category for the categorical region variable. While all estimates are statistically significant at the 0.05 significance level, exponentiated regression coefficient estimates from the logistic model demonstrate that the level of residential segregation, the median household income, and median property tax of the county have very minimal impacts on the odds of a school being adequately funded.
-
-Region, however, has a substantial impact on the odds of school funding adequacy.  Being located in the Northeast or the Midwest has the greatest impact on how well the school is funded, with the odds of adequate school funding being over 200 times higher in a Northeast county compared to a Southern county, and the odds of adequate school funding being over 3 times higher in a Northeast county compared to a Southern county, holding all other variables constant.
-Such disparities across the country are caused by a number of factors, including: (1) capacity - how well off a state is based on its economy and resources, and (2) effort - the states willingness to provide funding for education.
-Wealthier states with a high fiscal capacity, (typically those in the Northeast), have more funding available to spend on education than states with more limited resources (typically those in the South and the West).
-
-In comparing our full model (with predictor variables for residential segregation, median household income, median property tax, and region) to a model including all of these variables except for region, we found that there is statistical significance in the deviance between the residuals of the full model and reduced.
-As such, we can conclude that including region as a predictor refines and creates a better fit for our model.
-
-```{r, echo=FALSE, include=FALSE}
-## Checking to see if the imputated data has a large effect on the outcome
-
-# Logistic Regression without Imputated Data
-# Outcome = School.Funding.Cat
-# Predictors = residential segregation, median.household.income, median.prop.tax
-
-lmod <- glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax + Region_fac,
-              data=combined,
-              family="binomial")
-
-#log-odds scale
-modelsummary(lmod,
-             fmt = fmt_significant(2),
-             shape = term ~ model + statistic,
-             statistic = c("std.error","conf.int","p.value"),
-             exponentiate = FALSE,
-             gof_map=NA)
-
-modelsummary(lmod,
-             fmt = fmt_significant(2),
-             shape = term ~ model + statistic,
-             statistic = c("std.error","conf.int","p.value"),
-             exponentiate = TRUE,
-             gof_map=NA)
-
-```
-
-```{r, include=FALSE, echo=FALSE}
-imp_mods <- glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax + Region_fac, data=completed_data, family="binomial")
-
-#log-odds scale
-modelsummary(imp_mods,
-             fmt = fmt_significant(2),
-             shape = term ~ model + statistic,
-             statistic = c("std.error","conf.int","p.value"),
-             exponentiate = FALSE,
-             gof_map=NA)
-```
-
-*Figure 2: Relationship Between Residential Segregation, Region, & School Funding Adequacy*
-
-```{r}
-ggplot(completed_data, aes(x=Region_fac, y = Residential.Segregation.Index, color=factor(School.Funding.Cat))) +
-  geom_boxplot() +
-  labs(x = "U.S. Region", y='Residential Segregation Index', color='School Funding Adequacy', title='Residential Segregation Stratified by Region & School Funding') +
-  scale_color_hue(labels = c("Inadequate", "Adequate"))+ 
-  theme(plot.margin = margin(t = 20, r = 30, b = 20, l = 10),
-axis.title.x = element_text(size = 8),
-axis.title.y = element_text(size = 8),
-title=element_text(size=9))
-```
-
-*Table 3: Logistic Regression Results For School Funding Adequacy Outcome*
-
-```{r, echo=FALSE}
-
-renamed <- c(
-    "Residential.Segregation.Index" = "Residential Segregation Index (RSI)",
-    "Median.Household.Income"="Median Household Income, USD",
-    "Median.Prop.Tax" = "Median Property Tax, USD",
-  "Region_facWest"="Region-West", 
-  "Region_facNortheast"="Region-Northeast",
-  "Region_facMidwest"="Region-Midwest"
-)
-modelsummary(imp_mods,
-             fmt = fmt_significant(2),
-             shape = term ~ model + statistic,
-             coef_rename = renamed,
-             statistic = c("std.error","conf.int","p.value"),
-             exponentiate = TRUE,
-             gof_map=NA)
-```
-
-#### Model Assessment
-
-To assess logistic regression assumptions, we first plotted each numerical predictor against the log odds of school funding adequacy.
-All plots showed a mostly linear relationship, indicating that the linearity assumption was largely met; interestingly, different linear trends were observed within each plot based on region, pointing to potential interaction effects between region and the three numerical predictors.
-GVIF for all predictors ranged between 1 and 3, suggesting that there could be moderate multicollinearity between predictors, but not enough to warrant a reconsideration of included variables.
-Cook's distance was also less than 0.05 for all points, meaning that there were no highly influential observations in the model.
-Evaluating our model, we found that the logistic regression model demonstrated a strong overall performance with an accuracy rate of 80.33%.
-The model had a high sensitivity of 91%, indicating that it accurately predicted 91% of actual positive cases of adequate school funding when evaluated on our dataset.
-However, it had a 60% specificity rate, highlighting it correctly identified cases of inadequate school funding only 60% of the time.
-Overall, looking at the ROC curve below, the AUC is 0.858, indicating a moderately strong discriminate ability.
-
-```{r, include=FALSE, echo=FALSE}
-plot(completed_data$Residential.Segregation.Index, predict(imp_mods), col = completed_data$Region_fac)
-plot(completed_data$Median.Household.Income, predict(imp_mods), col = completed_data$Region_fac)
-plot(completed_data$Median.Prop.Tax, predict(imp_mods), col = completed_data$Region_fac)
-```
-
-```{r, include=FALSE, echo=FALSE}
-#Cooks Distance 
-cooksd <- cooks.distance(imp_mods)
-
-# Plot Cook's Distance
-plot(cooksd, main = "Cook's Distance", ylab = "Cook's Distance")
-abline(h = 4/(nrow(completed_data) - length(coef(imp_mods))), col = "red")  # Threshold line
-4/(nrow(data) - length(coef(imp_mods)))
-```
-
-```{r, echo=FALSE, include=FALSE}
-vif(imp_mods)
-plot(imp_mods)
-```
-
-```{r, echo=FALSE, include=FALSE}
-# Creating the confusion matrix
-
-# set.seed(42)
-# train_idx <- sample(1:nrow(long_data), size = 0.5 * nrow(long_data))
-# train_data <- long_data[train_idx, ]
-# test_data <- long_data[-train_idx, ]
-
-#t_glm_model <- glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax, data = completed_data, family = binomial)
-
-# Make predictions on the test set
-test_probs <- predict(imp_mods, newdata = completed_data, type = "response")
-test_pred <- ifelse(test_probs > 0.5, 1, 0)  # Classify based on threshold 0.5
-```
-
-```{r, echo=FALSE, include=FALSE}
-# Create a confusion matrix
-conf_matrix <- confusionMatrix(factor(test_pred), factor(completed_data$School.Funding.Cat))
-#print(conf_matrix)
-
-# Format and display confusion matrix as a Markdown table
-pander(conf_matrix, caption = "Confusion Matrix")
-
-```
-
-```{r, echo=FALSE, include=FALSE}
-#Creating the AUC curves
-roc_result <- roc(
-  completed_data$School.Funding.Cat,
-  predict(imp_mods, type = "response"),
-  legacy.axes = TRUE,
-  plot = TRUE,
-  print.thres = 0.5,
-  print.auc = FALSE  # Disable automatic AUC
-)
-
-# Add AUC manually
-text(
-  x = 1.2,  # Adjust X-coordinate
-  y = 1,  # Adjust Y-coordinate
-  labels = sprintf("AUC = %.3f", auc(roc_result)),  # Add AUC value
-  cex = 1,  # Label size
-  col = "red"  # Label color
-)
-```
-
-```{r, include=FALSE, echo=FALSE}
-imp_mods_red <- glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax, data=completed_data, family="binomial")
-
-anova(imp_mods_red, imp_mods, test="Chisq")
-```
-
-## **Conclusion**
-
-In our analysis, we observed the significant impact of the regional location of a county on both its average life expectancy (a health indicator) and school funding adequacy (an educational indicator)--two key components of societal and an individual’s well-being.
-Life expectancy is highest in the Northeast, followed by the Midwest, West, and South, while school funding adequacy is significantly greater in the Northeast and Midwest compared to the South.
-Our findings highlight substantial differences in both outcomes between regions, which are often caused by different economic capabilities and different allocations of resources among countries.
-This study provides empirical evidence that underscores the importance for equitable social policies and targeted investments to address regional disparities and promote well-being across all communities.
-
-Our analysis faced limitations due to missing data, which restricted the dataset to 33 states and reduced the generalizability of our findings.
-Additionally, some variables were drawn from different years between 2018 and 2022, requiring us to assume that trends remained consistent over a 1–5 year period and that time was an insignificant factor; however, this may not actually be the case.
-
-Future studies could incorporate additional social variables to enhance the analysis and explore causal relationships between region and outcomes.
-For example, integrating a dataset on employment rates and school performance could offer deeper insights into how socioeconomic factors shape life expectancy and educational funding, complementing existing variables in our models.
-Another direction for future work is to delve deeper into understanding the surprising findings regarding the effects of uninsured adult percentage and residential segregation on life expectancy.
-Expanding the dataset to include a temporal scope, and finding more complete sources of data on variables of interest so that no states have to be excluded from the analysis, would also contribute to a more comprehensive understanding.
-
-## **Appendix**
-
-*Table A1: Summary Statistics For All Variables*
-
-```{r}
-renamed_data <- completed_data %>%
-  rename(
-    "Residential Segregation" = Residential.Segregation.Index,
-    "Median Household Income" = Median.Household.Income,
-    "Median Property Tax" = Median.Prop.Tax,
-    "School Funding Adequacy" = School.Funding.Adequacy, 
-    "Life Expectancy" = Life.Expectancy,
-    "% Limited Access, Healthy Foods" = X..Limited.Access.to.Healthy.Foods,
-    "% Uninsured Adults" = X..Uninsured.Adults,
-    "Region" = Region_fac,
-    "School Funding Binary" =School.Funding.Cat)
-
-skim_table <- datasummary_skim(
-  renamed_data,
-  title = "Summary Statistics")
-
-```
-
-```{r, include=FALSE, echo=FALSE}
-summary(completed_data)
-```
-
-## **References**
-
-1.
-National Academies of Sciences, Engineering, and Medicine.
-(2017).
-*Communities in action: Pathways to health equity.* Washington, DC: The National Academies Press.
-Retrieved from <https://www.ncbi.nlm.nih.gov/books/NBK425845/>
-
-2.
-Stanford Center for Opportunity Policy in Education.
-(2013.).
-How to close the opportunity gap: Key policy recommendations.
-Retrieved from <https://edpolicy.stanford.edu/sites/default/files/Opp%20Gap%20Policy%20Recommendations.pdf>
-
-3.  County Health Rankings & Roadmaps. (n.d.). *About us.* Retrieved December 8, 2024, from <https://www.countyhealthrankings.org/about-us>
-
-4.
-Steven Ruggles, Sarah Flood, Matthew Sobek, Daniel Backman, Annie Chen, Grace Cooper, Stephanie Richards, Renae Rodgers, and Megan Schouweiler.
-IPUMS USA: Version 15.0 \[dataset\].
-Minneapolis, MN: IPUMS, 2024.
-Retrieved from <https://doi.org/10.18128/D010.V15.0>
-
-5.  Halpert, C. (2014). Census regions \[GitHub repository\]. GitHub. <https://github.com/cphalpert/census-regions>
-
-6.  USDA. (2022). *Documentation*. Economic Research Service. https://www.ers.usda.gov/data-products/food-access-research-atlas/documentation/
-
-7.
-County Health Rankings & Roadmaps.
-(2024).
-*Residential segregation - Black/White.* University of Wisconsin Population Health Institute.
-Retrieved from <https://www.countyhealthrankings.org/health-data/health-factors/social-economic-factors/family-and-social-support/residential-segregation-blackwhite?utm_source=chatgpt.com&year=2024>
-
-8.
-Chetty, R., Stepner, M., Abraham, S., Lin, S., Scuderi, B., Turner, N., Bergeron, A., & Cutler, D.
-(2016).
-The Association Between Income and Life Expectancy in the United States, 2001-2014.
-*JAMA*, *315*(16), 1750–1766.
-https://doi.org/10.1001/jama.2016.4226
-
-9.
-County Health Rankings & Roadmaps.
-(2024).
-*School funding adequacy.* University of Wisconsin Population Health Institute.
-Retrieved from <https://www.countyhealthrankings.org/health-data/health-factors/social-economic-factors/education/school-funding-adequacy?utm_source=chatgpt.com&year=2024>
+\end{document}

### 2024-12-15 paper.qmd
fixed rendered photo
fixing format
other files
final

@@
 library(pander)
-library(kableExtra)
 ```
@@
 
-Table 3 below displays the odds ratio results from the logistic regression model for school funding adequacy, with the South as the baseline category for the categorical region variable. While all estimates are statistically significant at the 0.05 significance level, exponentiated regression coefficient estimates from the logistic model demonstrate that the level of residential segregation, the median household income, and median property tax of the county have very minimal impacts on the odds of a school being adequately funded.
+Table 3 below displays the odds ratio results from the logistic regression model for school funding adequacy, with the South as the baseline category for the categorical region variable.
+While all estimates are statistically significant at the 0.05 significance level, exponentiated regression coefficients from the logistic model demonstrate median household income and median property tax of the county have almost no impact on the odds of a school being adequately funded.
+Our results demonstrate that a one unit increase to the residential segregation index increases the odds of school funding adequacy by 0.02 times, allowing us to conclude that residential segregation has a minimal effect on the odds of the school being adequately funded.
+Figure 2 demonstrates an interesting trend–each region shows a different relationship between the residential segregation distribution of adequately and inadequately funded schools, pointing to a potential interaction between region and segregation.
 
@@
 
-```{r}
+```{r, include=FALSE, echo=FALSE}
 renamed_data <- completed_data %>%
@@
 
-```{r, include=FALSE, echo=FALSE}
-summary(completed_data)
-```
+## ![](images/clipboard-2912764107.png)
 

### 2024-12-15 paper.qmd
final with edits

@@
 
-This study examines the relationships between socioeconomic, health, and education indicators in U.S.
-counties using data from the American Community Survey, the USDA Food Environment Atlas, Small Area Income and Poverty Estimates, and other sources.We analyze how factors such as residential segregation, health insurance coverage, access to healthy foods, and region influence life expectancy, and how residential segregation and region affect school funding adequacy.
+This study examines the relationships between socioeconomic, health, and educational indicators in U.S.
+counties using data from the County Health Rankings & Roadmaps program and IPUMS.
+We analyze how factors such as residential segregation, health insurance coverage, access to healthy foods, and region impact life expectancy, and how residential segregation and region affect school funding adequacy.
 Our findings reveal that life expectancy is influenced by residential segregation and uninsured rates, though these effects are minor compared to the more substantial influence of household income and regional differences.
@@
 Life expectancy and school funding are important indicators of public well-being because they capture fundamental aspects of health, education, and social equity that are essential for the development of individuals and communities.
-Life expectancy is an important health indicator that reflects not only the overall health of a society, but also its stability, economic resilience, and levels of inequality.
+Life expectancy is an important indicator that reflects not only the overall health of a society, but also its stability, economic resilience, and levels of inequality.
 While access to quality healthcare and nutritious foods are drivers of well-being, where you live and the nature of the community you live in likely also have an influence on health outcomes such as life expectancy.
@@
 
-Variables of interest were selected a priori based on their perceived relevance to the two outcomes we wanted to better understand: average life expectancy and school funding adequacy.
+Variables of interest were selected *a priori* based on their perceived relevance to the two outcomes we wanted to better understand: average life expectancy and school funding adequacy.
 While nutrition and access to quality healthcare are inherently linked to general wellbeing, it is worthwhile to assess how influential the environmental and financial aspects of these factors are on our tangible outcome.
@@
 
-### *Data Overview*
+### Data Overview
 
@@
 Based on the multiple linear regression results, residential segregation, percentage of uninsured adults, region, and median household income all showed statistically significant effects on life expectancy.
-Specifically, for every \$10,000 increase in median household income, life expectancy increases by 1.25 years on average.
+Specifically, for every \$10,000 increase in median household income, life expectancy increases by 1.25 years on average, holding other variables constant.
 In addition, the impact of region on life expectancy was substantial: life expectancy is on average 1.96 years higher in the West region compared to the South, 2.92 years higher in the Northeast compared to the South, and 1.92 years higher in the Midwest compared to the South, even when controlling for income, uninsured population, and racial diversity.
@@
 Surprisingly, holding all other variables constant, for every 1 unit increase in percentage of uninsured adults in a county, a county's average life expectancy increases by 0.032 years.
-While statistically significant, many studies covering the relationship between an individual’s health and having insurance coverage have proven its positive correlation, suggesting our results may benefit from a closer look.
-In assessing our dataset, as seen in Appendix 1A, we find that our dataset contains more observations for counties with low uninsured individuals, or equivalently, counties where most people are covered.
+While this was statistically significant, many studies covering the relationship between an individual’s health and having insurance coverage have proven its positive correlation, suggesting our results may benefit from a closer look.
+In assessing our dataset, as seen in Appendix A1, we find that our dataset contains more observations for counties with low uninsured individuals, or equivalently, counties where most people are covered.
 As the timeframe of our dataset is 2018-2022, it is likely that the aforementioned observations require greater scrutiny of the significance of the impact of this variable on average life expectancy of a county.
 
-The percentage of population with limited access to healthy foods predictor was not statistically significant in our model, suggesting that the effect of this environmental factor is minor or intertwined with economic variables, such as income, rather than being a direct or standalone influence.
+The percentage of population with limited access to healthy foods predictor was not statistically significant in our model, suggesting that the effect of this environmental factor is minor or intertwined with economic variables, such as income, rather than being a standalone influence.
 In terms of residential segregation, for every 1 unit increase in a county's index value (corresponding to increased segregation), its average life expectancy increases by 0.0071 years in the South, increases by 0.0068 years in the West, decreases by 0.0055 years in the Northeast, and decreases by 0.0004 years in the Midwest, as seen in Figure 1.
@@
 We confirmed this result with a nested F-test, comparing a full model to a reduced model that excluded the interaction term between region and residential segregation.
-This did not show a statistically significant difference in RSS (p=0.624), indicating that the interaction between region and residential segregation does not add meaningful explanatory power to our model, Consequently, we did not find sufficient evidence to support regional variations in the relationship between segregation and life expectancy.
+This did not show a statistically significant difference in RSS (*p*=0.624), indicating that the interaction between region and residential segregation does not add meaningful explanatory power to our model, Consequently, we did not find sufficient evidence to support regional variations in the relationship between segregation and life expectancy.
 Instead, based on residential segregation's coefficient estimate in the model, life expectancy does appear to increase overall with an increase in residential segregation, however minor this increase may be.
-One possible theory for this observation is that counties with more racial uniformity might experience less racial conflict and more social cohesion, especially for minority communities.
+One possible theory for this observation is that counties with more racial uniformity might experience less racial conflict and more social cohesion, especially for minority communities, but this should be further investigated in future studies.
 
@@
 leverage plot; no points were found to be influential.
-There was a high leverage point (Mineral County, Nevada), and refitting the model without this point yielded no substantial changes in the coefficient estimates for statistically significant predictors; however, the p-value for residential segregation increased from 0.031 to 0.054.
+There was a high leverage point (Mineral County, Nevada), and refitting the model without this point yielded no substantial changes in the coefficient estimates for statistically significant predictors; however, the *p*-value for residential segregation increased from 0.031 to 0.054.
 Finally, with a VIF \<2 for all predictors, there was little to no multicollinearity among the predictors.
@@
 Table 3 below displays the odds ratio results from the logistic regression model for school funding adequacy, with the South as the baseline category for the categorical region variable.
-While all estimates are statistically significant at the 0.05 significance level, exponentiated regression coefficients from the logistic model demonstrate median household income and median property tax of the county have almost no impact on the odds of a school being adequately funded.
+While all estimates are statistically significant at the 0.05 significance level, exponentiated regression coefficients from the logistic model demonstrate that a county's median household income and median property tax have almost no impact on the odds of its schools being adequately funded.
 Our results demonstrate that a one unit increase to the residential segregation index increases the odds of school funding adequacy by 0.02 times, allowing us to conclude that residential segregation has a minimal effect on the odds of the school being adequately funded.
-Figure 2 demonstrates an interesting trend–each region shows a different relationship between the residential segregation distribution of adequately and inadequately funded schools, pointing to a potential interaction between region and segregation.
+Figure 2 demonstrates an interesting trend—each region shows a different relationship between the residential segregation distribution of adequately and inadequately funded schools, pointing to a potential interaction between region and segregation.
 
-Region, however, has a substantial impact on the odds of school funding adequacy.  Being located in the Northeast or the Midwest has the greatest impact on how well the school is funded, with the odds of adequate school funding being over 200 times higher in a Northeast county compared to a Southern county, and the odds of adequate school funding being over 3 times higher in a Northeast county compared to a Southern county, holding all other variables constant.
-Such disparities across the country are caused by a number of factors, including: (1) capacity - how well off a state is based on its economy and resources, and (2) effort - the states willingness to provide funding for education.
+Region has a substantial impact on the odds of school funding adequacy. Being located in the Northeast or the Midwest has the greatest impact on how well a school is funded, with the odds of adequate school funding being over 200 times higher in a Northeast county compared to a Southern county, and the odds of adequate school funding being over 3 times higher in a Midwest county compared to a Southern county, holding all other variables constant.
+Such disparities across the country are caused by a number of factors, including: (1) capacity - how well off a state is based on its economy and resources, and (2) effort - the state's willingness to provide funding for education.
 Wealthier states with a high fiscal capacity, (typically those in the Northeast), have more funding available to spend on education than states with more limited resources (typically those in the South and the West).
@@
 To assess logistic regression assumptions, we first plotted each numerical predictor against the log odds of school funding adequacy.
-All plots showed a mostly linear relationship, indicating that the linearity assumption was largely met; interestingly, different linear trends were observed within each plot based on region, pointing to potential interaction effects between region and the three numerical predictors.
+All plots showed a mostly linear relationship, indicating that the linearity assumption was largely met; interestingly, different linear trends were observed within each plot based on region, further supporting potential interaction effects between region and the three numerical predictors.
 GVIF for all predictors ranged between 1 and 3, suggesting that there could be moderate multicollinearity between predictors, but not enough to warrant a reconsideration of included variables.
@@
 
-In our analysis, we observed the significant impact of the regional location of a county on both its average life expectancy (a health indicator) and school funding adequacy (an educational indicator)--two key components of societal and an individual’s well-being.
+In our analysis, we observed the significant impact of the regional location of a county on both its average life expectancy (a health indicator) and school funding adequacy (an educational indicator)--two key components of societal and individual well-being.
 Life expectancy is highest in the Northeast, followed by the Midwest, West, and South, while school funding adequacy is significantly greater in the Northeast and Midwest compared to the South.
@@
 
-Future studies could incorporate additional social variables to enhance the analysis and explore causal relationships between region and outcomes.
+Future studies could look at additional interaction effects with region and incorporate more social variables to enhance the analysis.
 For example, integrating a dataset on employment rates and school performance could offer deeper insights into how socioeconomic factors shape life expectancy and educational funding, complementing existing variables in our models.

### 2024-12-15 paper.qmd
minor code changes to Q2

@@
 modelsummary(lmod,
-             fmt = fmt_significant(2),
+             fmt = fmt_significant(7),
              shape = term ~ model + statistic,
@@
 # Creating the confusion matrix
-
-# set.seed(42)
-# train_idx <- sample(1:nrow(long_data), size = 0.5 * nrow(long_data))
-# train_data <- long_data[train_idx, ]
-# test_data <- long_data[-train_idx, ]
-
-#t_glm_model <- glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax, data = completed_data, family = binomial)
 

### 2024-12-15 final_code.qmd
final code file

--- 2024-12-15 paper.qmd
@@
 
+## Cleaning
+
+```{r, echo=FALSE}
+# Load necessary library for rename
+library(dplyr)
+library(readxl)
+
+# Vector of all 50 states
+states <- c("Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", 
+            "Connecticut", "Delaware", "Florida", "Georgia", "Hawaii", "Idaho", 
+            "Illinois", "Indiana", "Iowa", "Kansas", "Kentucky", "Louisiana", 
+            "Maine", "Maryland", "Massachusetts", "Michigan", "Minnesota", 
+            "Mississippi", "Missouri", "Montana", "Nebraska", "Nevada", 
+            "New Hampshire", "New Jersey", "New Mexico", "New York", 
+            "North Carolina", "North Dakota", "Ohio", "Oklahoma", "Oregon", 
+            "Pennsylvania", "Rhode Island", "South Carolina", "South Dakota", 
+            "Tennessee", "Texas", "Utah", "Vermont", "Virginia", "Washington", 
+            "West Virginia", "Wisconsin", "Wyoming")
+
+# Vector of desired columns to keep
+desired_cols <- c('FIPS', 'State',  
+                  "Life Expectancy", "Segregation Index...200",
+                  "% Limited Access to Healthy Foods", 
+                  "% Uninsured Adults", 
+                  "Spending per Pupil", "School Funding Adequacy", 
+                  "Median Household Income" )
+
+# Base URL
+base_path <- "F24_STATS_FINAL/county_data/"
+
+# Initialize an empty list to store processed data frames
+all_data <- list()
+
+# Loop through each state
+for (state in states) {
+  # Format the state name for URL (replace spaces with %20)
+  state_url <- paste0(base_path, gsub(" ", "\ ", state), ".xlsx")
+
+  # Read the CSV
+  full_data <- read_excel(state_url, sheet="Additional Measure Data", skip=1)
+
+  # Subset columns
+  subset_data <- full_data[desired_cols]
+
+  # Rename columns
+  renamed_data <- subset_data %>%
+    rename(
+      "Residential.Segregation.Index" = "Segregation Index...200",
+      "School.Funding.Adequacy" = "School Funding Adequacy"
+    )
+
+  # Drop the first row
+  processed_data <- renamed_data[-1, ]
+
+  # Save processed data to the list with a dynamic name
+  all_data[[tolower(state)]] <- processed_data
+}
+
+# Combine all processed data frames into one
+combined <- do.call(rbind, all_data)
+
+#Create categorical variable for school funding adequacy
+combined_pre_merge <- combined |>
+  mutate(
+    School.Funding.Cat = factor(case_when(
+      School.Funding.Adequacy < 0 ~ 'Inadequate', 
+      School.Funding.Adequacy > 0  ~ 'Adequate',
+      TRUE ~ NA
+    ), levels=c('Inadequate', 'Adequate'), labels=c(0, 1)))
+
+#Merge with ACS (property tax) data
+acs <- read.csv("https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/county_data/acs_prop_2.csv")
+combined_pre_merge$FIPS <- as.numeric(combined_pre_merge$FIPS)
+
+merged_df <- full_join(combined_pre_merge, acs, join_by("FIPS" == "fips")) %>%
+  mutate(
+    indicator = case_when(
+      is.na(State) & !is.na(state) ~ "Right Only",  # Row from acs only
+      !is.na(State) & is.na(state) ~ "Left Only",   # Row from combined_pre_merge only
+      !is.na(State) & !is.na(state) ~ "Both"        # Row present in both dfs
+    )
+  )
+
+print(merged_df[merged_df$indicator == "Right Only", ])
+print(merged_df[merged_df$indicator == "Left Only", ])
+#Difference of 8 rows, indicating that there are 8 counties with health data but no property tax data. 
+#Alaska and South Dakota are going to be dropped anyways due to missing residential segregation index data.
+#Looking into the VA county with missing property tax data (51515), it turns out that this county does not have any
+#health data either in the original spreadsheet, but is still recorded as a county. Upon further analysis
+#it was found that this FIPS code corresponded to Bedford City, which was merged into 51519 in 2010. Therefore it was safe to drop.
+
+merged_df <- merged_df[merged_df$FIPS != 51515, ]
+merged_df <- merged_df |> 
+  mutate(
+    Median.Prop.Tax = as.numeric(med_.re_taxes_paid))
+
+intermed_merged = subset(merged_df, select = -c(med_.re_taxes_paid, med_.re_taxes_paid_real, county.1, Location, indicator, county, state) )
+
+#Now merge with regions data
+regions <- read.csv("https://github.com/cphalpert/census-regions/raw/refs/heads/master/us%20census%20bureau%20regions%20and%20divisions.csv")
+final_merged <- left_join(intermed_merged, regions, by = c("State" = "State"))
+final_merged = subset(final_merged, select = -c(State.Code))
+
+#Due to substantial missing data in certain states, remove them from the final dataframe
+states_to_exclude <- c('Alaska', 'Colorado', 'Idaho', 'Iowa', 'Kansas', 'Minnesota', 'Montana', 'Nebraska', 'North Dakota', 'South Dakota', 'Utah', 'Wyoming', 'Vermont')
+combined <- final_merged %>%
+  filter(!State %in% states_to_exclude)
+
+# The final data frame `combined` contains all processed state data.
+#write.csv(final, '~/Documents/R\ Course\ Code/F24_STATS_FINAL/combined.csv')
+```
+
+## Imputation
+
 ```{r,echo= FALSE, warning = FALSE, message=FALSE}
-combined <- read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/combined.csv')
+#combined <- read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/combined.csv')
 
@@
 
-## Abstract
-
-This study examines the relationships between socioeconomic, health, and education indicators in U.S.
-counties using data from the American Community Survey, the USDA Food Environment Atlas, Small Area Income and Poverty Estimates, and other sources.We analyze how factors such as residential segregation, health insurance coverage, access to healthy foods, and region influence life expectancy, and how residential segregation and region affect school funding adequacy.
-Our findings reveal that life expectancy is influenced by residential segregation and uninsured rates, though these effects are minor compared to the more substantial influence of household income and regional differences.
-Life expectancy averages are highest in the Northeast, followed by the Midwest, West, and South.
-Region also plays the most significant role out of all included variables in determining the likelihood of school funding adequacy, with schools in the Northeast and Midwest being far more likely to receive adequate funding than those in the South.
-Other factors, such as residential segregation, property taxes, and household income, show statistically significant but smaller effects.
-These results highlight the complex interplay between socioeconomic variables and geographic location in shaping public well-being.
-By identifying key disparities in health and education, this study underscores the importance of equitable resource allocation and targeted policies to address regional and socioeconomic inequalities.
-
-## **Introduction**
-
-Life expectancy and school funding are important indicators of public well-being because they capture fundamental aspects of health, education, and social equity that are essential for the development of individuals and communities.
-Life expectancy is an important health indicator that reflects not only the overall health of a society, but also its stability, economic resilience, and levels of inequality.
-While access to quality healthcare and nutritious foods are drivers of well-being, where you live and the nature of the community you live in likely also have an influence on health outcomes such as life expectancy.
-According to "The Root Causes of Health Inequity - Communities in Action" report from the National Institutes of Health, we can infer that differences in life expectancy between regions and socioeconomic groups often reveal systemic inequalities, with marginalized communities experiencing shorter life spans due to limited access to health care, healthy living environments, and economic opportunities.^1^ When it comes to educational indicators, adequately funded schools ensure that students have equal access to learning opportunities, serving as a crucial investment in social mobility.
-The Stanford Center for Opportunity Policy in Education highlights that "the unequal allocation and inadequate levels of resources in schools and communities \[are\] at the heart of many gaps in student opportunity."^2^ This emphasizes the importance of equitable and adequate school funding in closing opportunity gaps and empowering students from disadvantaged backgrounds to achieve academic and economic success.
-By better understanding the correlates of school funding adequacy, we can take a major step towards implementing policy that targets the areas and communities most affected by this issue.
-
-For this purpose, we focus our analysis on two key research questions.
-First, we investigate how socioeconomic factors impact average life expectancy.
-**We seek to answer the question, "What is the impact of residential segregation, access to healthy foods, and health insurance coverage on a county's average life expectancy?".** This analysis includes a comparison across different regions to identify variations in the relationship between residential segregation and life expectancy.
-Second, **we seek to answer the question, "Do residential segregation and region impact the likelihood of a school being adequately funded?"**.
-This question allows us to assess how the racial makeup and geographic location—given that each region of the U.S.
-is characterized by differing political and cultural norms—affect a county's educational investment.
-The observational unit for this analysis is the U.S.
-county.
-
-To obtain data on the aforementioned variables, we used three primary sources.
-
-The first source was the County Health Rankings & Roadmaps program,^3^ associated with the University of Wisconsin Population Health Institute, which provided us with 2024 datasets for all U.S.
-states.
-This program consolidates the latest county-level measurements of various population health, economic, demographic, and social factors from the American Community Survey, the USDA Food Environment Atlas, Small Area Income and Poverty Estimates, and other reliable government sources into publicly available datasets every year.
-Our specific variables of interest were originally collected as part of:
-
--   National Center for Health Statistics - Natality & Mortality Files, 2019-2021 (average life expectancy)
-
--   American Community Survey 5-year estimates, 2018-2022 (residential segregation index & median household income)
-
--   USDA Food Environment Atlas, 2019 (percentage of population with limited access to healthy foods)
-
--   Small Area Health Insurance Estimates, 2021 (percentage of adults uninsured)
-
--   School Finance Indicators Database, 2021 (school funding adequacy)
-
-The second source was IPUMS, which provided us with median poverty tax data for every county from 2018-2022.^4^
-
-The third source was the U.S.
-Census Bureau’s Regions and Divisions of the United States, which maps every U.S.
-to one of four geographic regions, and also a division within the regions.
-We specifically included regions as a variable in this analysis. This data was loaded into R using a [publicly available GitHub repository](https://github.com/cphalpert/census-regions) titled *Census Regions*.^5^
-
-## **Methods**
-
-### Variable Selection
-
-Variables of interest were selected a priori based on their perceived relevance to the two outcomes we wanted to better understand: average life expectancy and school funding adequacy.
-While nutrition and access to quality healthcare are inherently linked to general wellbeing, it is worthwhile to assess how influential the environmental and financial aspects of these factors are on our tangible outcome.
-Therefore, we selected two key predictors for our life expectancy model: the percentage of the population with limited access to healthy foods and the percentage of uninsured adults in a county.
-The former is defined by the US Department of Agriculture as "the percentage of the \[county\] population that is low-income and does not live close to a grocery store", where "close" is defined differently for rural and nonrural areas.^6^ The latter variable represents the percentage of adults in a county younger than 65 who lack health insurance.
-We also were interested in seeing how residential segregation would impact life expectancy among these predictors, given existing studies on the negative health effects of segregation, and whether this relationship would exhibit any variations across different regions.
-This is why we chose to include residential segregation index values as a predictor.
-The residential segregation index, used by the County Health Rankings & Roadmaps program, measures how evenly two groups, such as Black and white residents, are distributed across geographic areas, with values ranging from 0 (complete integration) to 100 (complete segregation).^7^
-
-Past studies have established a correlation between income and life expectancy.^8^ Because we wanted to control for its effects and examine how insurance, healthy food access, segregation, and region influence life expectancy even within a given income bracket, we selected income as our final predictor for the first model.
-
-School funding adequacy is defined by the County Health Rankings program as "the average dollar gap between actual per-pupil spending and the amount needed for students to achieve national average test scores, considering districts' varying equity-based needs."^9^ To explore how racial diversity interface with variations in school funding, we selected residential segregation to be a part of this regression.
-In addition, given that state policy has a large influence on school administration and the determination of school needs, and that states in the same region tend to have similar political leanings and cultural beliefs, we wanted to analyze how much of a relationship exists between region and actual school funding levels.
-Because a major source of public school funding is tax revenue, median property tax and median household income were also included in the school funding model as potential confounders.
-This allowed us to assess whether the effects of segregation and region on funding existed outside of the effects of known economic factors.
-
-### Exploratory Data Analysis & Data Preprocessing
-
-2024 datasets for all U.S.
-states were combined and merged with median property tax and region mapping data, resulting in a final dataset with 3144 county observations and 9 different variables (state, average life expectancy, residential segregation index, % limited access to healthy foods, % uninsured adults, school funding adequacy, median household income, median property tax, and region).
-Because school funding adequacy was given as a numerical value, and we wanted to focus on better understanding the dichotomy between school funding adequacy and inadequacy, this variable was converted into a binary categorical variable (where the value was set to 1 if the numerical value was positive, indicating the counties’ schools were adequately funded, and set to 0 if the numerical value was negative, indicating underfunding).
-
-Exploratory data analysis for the two outcome variables was conducted by 1) calculating descriptive statistics, 2) creating scatterplots of each individual predictor against the outcome (for life expectancy), and 3) creating pairwise plots between variables (for school funding).
-During the initial stages of analysis, we found that ⅓ of our observations were missing residential segregation index data, and that certain states were missing nearly all of their segregation values.
-Because of this, we decided to drop all states missing over 50% of their values for any variable.
-This led to our analysis being limited to 33 states, and excluding the following: Alaska, Colorado, Idaho, Iowa, Kansas, Minnesota, Montana, Nebraska, North Dakota, South Dakota, Utah, Wyoming, and Vermont.
-We were left with 2379 total observations, and the missing percentage for residential segregation dropped to 22%.
-
-After this, the remaining missing values in all variables were imputed using the mice package and predictive mean matching to generate one complete dataset.
-The original school funding adequacy numerical variable was first imputed, and then converted into binary values for subsequent analysis.
-
-### Model Fitting & Assessment
-
-For our first research question, a multiple linear regression model was fit, regressing average life expectancy on residential segregation index, percentage with limited access to healthy foods, percentage uninsured adults, region, and median household income.
-Linear regression assumptions were assessed using diagnostic plots, especially residual vs.
-fitted and quantile-quantile plots.
-The adjusted R-squared value was used to evaluate the fit of the model, and the effect of region as an interaction term with residential segregation was evaluated using nested F tests.
-We also calculated VIF to check for multicollinearity and Cook’s distance to identify influential points.
-
-For our second research question, a binary logistic regression model was fit with school funding adequacy (categorical) as the outcome and residential segregation index, region, median property tax, and median household income as predictors.
-The model was assessed using a confusion matrix, an ROC curve, and a comparison of deviance between the full model and a reduced model that excluded region as a predictor.
-VIF and Cook's distance were again used to evaluate multicollinearity and influential points.
-
-## Results
-
-### *Data Overview*
-
-Table 1 displays summary statistics for the predictors in the second model, stratified by school funding adequacy.
-Table A1 in the Appendix provides summary statistics for all variables, and shows that both the mean and median for life expectancy are approximately 75 years.
-School funding adequacy counts show that around 66% of the counties in our analysis had underfunded schools.
-Additionally, the number of counties per region indicate an imbalance between categories, with the South region being over represented (holding 60% of all counties), and the East and Northeast holding around 9% of the counties each.
+## EDA
 
@@
 ```
-
-\newpage
 
@@
 
-```{r, echo=FALSE, include=FALSE}
+```{r, echo=FALSE}
 # Plot for Q2 EDA
@@
 
-### Question 1
+```{r, echo=FALSE}
+#Plot for Q1 EDA
+plot1 <- ggplot(completed_data, aes(x = Residential.Segregation.Index, y = Life.Expectancy, color=Region_fac)) +
+  geom_point(color='black') + geom_smooth(method="lm",se=F) + 
+  scale_color_brewer(palette = "Dark2") +
+  labs(x = "Residential Segregation Index", y = "Average Life Expectancy (Years)", color='U.S. Region')+ 
+  theme(plot.margin = margin(t = 20, r = 30, b = 20, l = 10),
+axis.title.x = element_text(size = 8),
+axis.title.y = element_text(size = 6),
+title=element_text(size=9))
+
+plot2 <- ggplot(completed_data, aes(x = X..Limited.Access.to.Healthy.Foods
+, y = Life.Expectancy)) +
+  geom_point(color='black') + geom_smooth(method="lm",se=F) + 
+  scale_color_brewer(palette = "Dark2") +
+  labs( x = "% Population With Limited Healthy Food Access", y="Average Life Expectancy (Years)") +
+  theme(plot.margin = margin(t = 20, r = 20, b = 20, l = 20),
+        axis.title.x = element_text(size = 8),
+        axis.title.y = element_text(size = 6))
+
+plot3 <- ggplot(completed_data, aes(x = X..Uninsured.Adults, y = Life.Expectancy)) +
+  geom_point(color='black') + geom_smooth(method="lm",se=F) + 
+  scale_color_brewer(palette = "Dark2") +
+  labs(x = "% Adult Population Without Health Insurance", y = "Average Life Expectancy (Years)")+ 
+  theme(plot.margin = margin(t = 20, r = 20, b = 20, l = 20),
+        axis.title.x = element_text(size = 8),
+        axis.title.y = element_text(size = 6))
+
+plot1 + plot2 + plot3 + plot_layout(ncol = 2, nrow=2) + 
+  plot_annotation(title = "Relationships Between Predictors and 
+                  Life Expectancy Outcome",
+                  theme = theme(
+      plot.title = element_text(hjust=0.5)))
+```
+
+### Question 1 Analysis
 
 #### Model Results
-
-Based on the multiple linear regression results, residential segregation, percentage of uninsured adults, region, and median household income all showed statistically significant effects on life expectancy.
-Specifically, for every \$10,000 increase in median household income, life expectancy increases by 1.25 years on average.
-In addition, the impact of region on life expectancy was substantial: life expectancy is on average 1.96 years higher in the West region compared to the South, 2.92 years higher in the Northeast compared to the South, and 1.92 years higher in the Midwest compared to the South, even when controlling for income, uninsured population, and racial diversity.
-This suggests, income as a factor withstanding, the differing quality of and access to healthcare and public services found in each region can have a significant impact on life expectancy.
-
-Surprisingly, holding all other variables constant, for every 1 unit increase in percentage of uninsured adults in a county, a county's average life expectancy increases by 0.032 years.
-While statistically significant, many studies covering the relationship between an individual’s health and having insurance coverage have proven its positive correlation, suggesting our results may benefit from a closer look.
-In assessing our dataset, as seen in Appendix 1A, we find that our dataset contains more observations for counties with low uninsured individuals, or equivalently, counties where most people are covered.
-As the timeframe of our dataset is 2018-2022, it is likely that the aforementioned observations require greater scrutiny of the significance of the impact of this variable on average life expectancy of a county.
-
-The percentage of population with limited access to healthy foods predictor was not statistically significant in our model, suggesting that the effect of this environmental factor is minor or intertwined with economic variables, such as income, rather than being a direct or standalone influence.
-In terms of residential segregation, for every 1 unit increase in a county's index value (corresponding to increased segregation), its average life expectancy increases by 0.0071 years in the South, increases by 0.0068 years in the West, decreases by 0.0055 years in the Northeast, and decreases by 0.0004 years in the Midwest, as seen in Figure 1.
-However, these regional differences are not statistically significant at the 0.05 significance level given the high p-values of all residential segregation-region interaction terms.
-We confirmed this result with a nested F-test, comparing a full model to a reduced model that excluded the interaction term between region and residential segregation.
-This did not show a statistically significant difference in RSS (p=0.624), indicating that the interaction between region and residential segregation does not add meaningful explanatory power to our model, Consequently, we did not find sufficient evidence to support regional variations in the relationship between segregation and life expectancy.
-Instead, based on residential segregation's coefficient estimate in the model, life expectancy does appear to increase overall with an increase in residential segregation, however minor this increase may be.
-One possible theory for this observation is that counties with more racial uniformity might experience less racial conflict and more social cohesion, especially for minority communities.
-
-*Figure 1: Segregation vs. Life Expectancy Across Different U.S. Regions*
 
@@
 
-```{r echo= FALSE, warning = FALSE, include=FALSE}
+```{r echo= FALSE, warning = FALSE}
 # Running a regression without the interaction term between residential segregation and region
@@
 
-```{r, include=FALSE, echo=FALSE}
+```{r, echo=FALSE}
 ##Nested F-Test to see if segregation-region interaction term makes a difference
@@
 
-The adjusted R-squared value for the model was 0.6065, indicating that a majority of the variability in life expectancy was explained by the socioeconomic and racial predictors included in the model.
-In conducting diagnostics on the multilinear regression model, we first checked to see if the model met assumptions of linearity, independence, normality, and homoscedasticity.
-The residuals vs.
-fitted plot showed relatively equal variance and no discernible pattern, meaning that the linearity and homoscedasticity assumptions were reasonably met.
-In addition, because each observation in the dataset represented a unique county in the U.S., independence could be reasonably assumed.
-Assessing the Q-Q Plot for normality, the residual distribution was approximately normal, despite a slight divergence on both tails of the plot.
-Cook's distance was also assessed using the residuals vs.
-leverage plot; no points were found to be influential.
-There was a high leverage point (Mineral County, Nevada), and refitting the model without this point yielded no substantial changes in the coefficient estimates for statistically significant predictors; however, the p-value for residential segregation increased from 0.031 to 0.054.
-Finally, with a VIF \<2 for all predictors, there was little to no multicollinearity among the predictors.
-
-```{r, echo=FALSE, include=FALSE}
+```{r, echo=FALSE}
 completed_data[1180,]
@@
 
-```{r, echo=FALSE, include=FALSE}
+```{r, echo=FALSE}
 #looking for outliers 
@@
 
-```{r, echo=FALSE, include=FALSE}
+```{r, echo=FALSEE}
 #Variance Inflaction Factor (VIF)
@@
 
-```{r, echo=FALSE, include=FALSE}
+```{r, echo=FALSE}
 #Generate 4 Diagnostic Plots at once
@@
 
-```{r, echo= FALSE, warning = FALSE, include=FALSE}
+```{r, echo= FALSE, warning = FALSE}
 # Running a regression with obs 1180 removed
@@
 
-```{r, echo=FALSE, include=FALSE}
+```{r, echo=FALSE}
 plot(mlr.lev)
@@
 
-Table 3 below displays the odds ratio results from the logistic regression model for school funding adequacy, with the South as the baseline category for the categorical region variable.
-While all estimates are statistically significant at the 0.05 significance level, exponentiated regression coefficients from the logistic model demonstrate median household income and median property tax of the county have almost no impact on the odds of a school being adequately funded.
-Our results demonstrate that a one unit increase to the residential segregation index increases the odds of school funding adequacy by 0.02 times, allowing us to conclude that residential segregation has a minimal effect on the odds of the school being adequately funded.
-Figure 2 demonstrates an interesting trend–each region shows a different relationship between the residential segregation distribution of adequately and inadequately funded schools, pointing to a potential interaction between region and segregation.
-
-Region, however, has a substantial impact on the odds of school funding adequacy.  Being located in the Northeast or the Midwest has the greatest impact on how well the school is funded, with the odds of adequate school funding being over 200 times higher in a Northeast county compared to a Southern county, and the odds of adequate school funding being over 3 times higher in a Northeast county compared to a Southern county, holding all other variables constant.
-Such disparities across the country are caused by a number of factors, including: (1) capacity - how well off a state is based on its economy and resources, and (2) effort - the states willingness to provide funding for education.
-Wealthier states with a high fiscal capacity, (typically those in the Northeast), have more funding available to spend on education than states with more limited resources (typically those in the South and the West).
-
-In comparing our full model (with predictor variables for residential segregation, median household income, median property tax, and region) to a model including all of these variables except for region, we found that there is statistical significance in the deviance between the residuals of the full model and reduced.
-As such, we can conclude that including region as a predictor refines and creates a better fit for our model.
-
-```{r, echo=FALSE, include=FALSE}
+```{r, echo=FALSE}
 ## Checking to see if the imputated data has a large effect on the outcome
@@
 
-```{r, include=FALSE, echo=FALSE}
+```{r, echo=FALSE}
 imp_mods <- glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax + Region_fac, data=completed_data, family="binomial")
@@
 
-To assess logistic regression assumptions, we first plotted each numerical predictor against the log odds of school funding adequacy.
-All plots showed a mostly linear relationship, indicating that the linearity assumption was largely met; interestingly, different linear trends were observed within each plot based on region, pointing to potential interaction effects between region and the three numerical predictors.
-GVIF for all predictors ranged between 1 and 3, suggesting that there could be moderate multicollinearity between predictors, but not enough to warrant a reconsideration of included variables.
-Cook's distance was also less than 0.05 for all points, meaning that there were no highly influential observations in the model.
-Evaluating our model, we found that the logistic regression model demonstrated a strong overall performance with an accuracy rate of 80.33%.
-The model had a high sensitivity of 91%, indicating that it accurately predicted 91% of actual positive cases of adequate school funding when evaluated on our dataset.
-However, it had a 60% specificity rate, highlighting it correctly identified cases of inadequate school funding only 60% of the time.
-Overall, looking at the ROC curve below, the AUC is 0.858, indicating a moderately strong discriminate ability.
-
-```{r, include=FALSE, echo=FALSE}
+```{r, echo=FALSE}
 plot(completed_data$Residential.Segregation.Index, predict(imp_mods), col = completed_data$Region_fac)
@@
 
-```{r, include=FALSE, echo=FALSE}
+```{r, echo=FALSE}
 #Cooks Distance 
@@
 
-```{r, echo=FALSE, include=FALSE}
+```{r, echo=FALSE}
 vif(imp_mods)
@@
 
-```{r, echo=FALSE, include=FALSE}
+```{r, echo=FALSE}
 # Creating the confusion matrix
@@
 
-```{r, echo=FALSE, include=FALSE}
+```{r, echo=FALSE}
 # Create a confusion matrix
@@
 
-```{r, echo=FALSE, include=FALSE}
+```{r, echo=FALSE}
 #Creating the AUC curves
@@
 
-```{r, include=FALSE, echo=FALSE}
+```{r, echo=FALSE}
 imp_mods_red <- glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax, data=completed_data, family="binomial")
@@
 ```
-
-## **Conclusion**
-
-In our analysis, we observed the significant impact of the regional location of a county on both its average life expectancy (a health indicator) and school funding adequacy (an educational indicator)--two key components of societal and an individual’s well-being.
-Life expectancy is highest in the Northeast, followed by the Midwest, West, and South, while school funding adequacy is significantly greater in the Northeast and Midwest compared to the South.
-Our findings highlight substantial differences in both outcomes between regions, which are often caused by different economic capabilities and different allocations of resources among countries.
-This study provides empirical evidence that underscores the importance for equitable social policies and targeted investments to address regional disparities and promote well-being across all communities.
-
-Our analysis faced limitations due to missing data, which restricted the dataset to 33 states and reduced the generalizability of our findings.
-Additionally, some variables were drawn from different years between 2018 and 2022, requiring us to assume that trends remained consistent over a 1–5 year period and that time was an insignificant factor; however, this may not actually be the case.
-
-Future studies could incorporate additional social variables to enhance the analysis and explore causal relationships between region and outcomes.
-For example, integrating a dataset on employment rates and school performance could offer deeper insights into how socioeconomic factors shape life expectancy and educational funding, complementing existing variables in our models.
-Another direction for future work is to delve deeper into understanding the surprising findings regarding the effects of uninsured adult percentage and residential segregation on life expectancy.
-Expanding the dataset to include a temporal scope, and finding more complete sources of data on variables of interest so that no states have to be excluded from the analysis, would also contribute to a more comprehensive understanding.
 
@@
 
-## ![](images/clipboard-2912764107.png)
+##
 
-## **References**
-
-1.
-National Academies of Sciences, Engineering, and Medicine.
-(2017).
-*Communities in action: Pathways to health equity.* Washington, DC: The National Academies Press.
-Retrieved from <https://www.ncbi.nlm.nih.gov/books/NBK425845/>
-
-2.
-Stanford Center for Opportunity Policy in Education.
-(2013.).
-How to close the opportunity gap: Key policy recommendations.
-Retrieved from <https://edpolicy.stanford.edu/sites/default/files/Opp%20Gap%20Policy%20Recommendations.pdf>
-
-3.  County Health Rankings & Roadmaps. (n.d.). *About us.* Retrieved December 8, 2024, from <https://www.countyhealthrankings.org/about-us>
-
-4.
-Steven Ruggles, Sarah Flood, Matthew Sobek, Daniel Backman, Annie Chen, Grace Cooper, Stephanie Richards, Renae Rodgers, and Megan Schouweiler.
-IPUMS USA: Version 15.0 \[dataset\].
-Minneapolis, MN: IPUMS, 2024.
-Retrieved from <https://doi.org/10.18128/D010.V15.0>
-
-5.  Halpert, C. (2014). Census regions \[GitHub repository\]. GitHub. <https://github.com/cphalpert/census-regions>
-
-6.  USDA. (2022). *Documentation*. Economic Research Service. https://www.ers.usda.gov/data-products/food-access-research-atlas/documentation/
-
-7.
-County Health Rankings & Roadmaps.
-(2024).
-*Residential segregation - Black/White.* University of Wisconsin Population Health Institute.
-Retrieved from <https://www.countyhealthrankings.org/health-data/health-factors/social-economic-factors/family-and-social-support/residential-segregation-blackwhite?utm_source=chatgpt.com&year=2024>
-
-8.
-Chetty, R., Stepner, M., Abraham, S., Lin, S., Scuderi, B., Turner, N., Bergeron, A., & Cutler, D.
-(2016).
-The Association Between Income and Life Expectancy in the United States, 2001-2014.
-*JAMA*, *315*(16), 1750–1766.
-https://doi.org/10.1001/jama.2016.4226
-
-9.
-County Health Rankings & Roadmaps.
-(2024).
-*School funding adequacy.* University of Wisconsin Population Health Institute.
-Retrieved from <https://www.countyhealthrankings.org/health-data/health-factors/social-economic-factors/education/school-funding-adequacy?utm_source=chatgpt.com&year=2024>
+##

### 2024-12-15 code.qmd
local to fix merge conflict

@@
 
+```{r}
+library("dplyr") 
+library(mice) 
+library(sjlabelled) 
+library(tidyverse) 
+library("modelsummary") 
+library(car)
+library(GGally)
+library(ggplot2)
+library(patchwork)
+library(caret)
+library(pROC)
+library(pander)
+```
+
+Imputation
+
+```{r}
+combined <- read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/combined.csv')
+```
+
+```{r, echo= FALSE, warning = FALSE}
+# Making the state_fac variable a factor variable for future use
+combined <- combined |> mutate(State_fac = factor(combined$State), Region_fac = factor(combined$Region, levels=c('South', 'West', 'Northeast', 'Midwest')))
+```
+
+```{r, echo= FALSE, warning = FALSE}
+## Imputation Code 
+# In this block, we are looking to use predictive mean matching to combing regression methods and hot deck imputing to predicted value closest to the predicted value of the observation with missing data; we also needed to convert school funding adequacy imputed values to school funding categorical variable before using in models
+
+sub <- combined |> select(c("Life.Expectancy","Residential.Segregation.Index","X..Limited.Access.to.Healthy.Foods","X..Uninsured.Adults","Median.Household.Income", "School.Funding.Adequacy", "Median.Prop.Tax", "Region_fac"))
+sub <- unlabel(sub) #unlabel the data (labels cause problem for the mice function)
+
+# Add School.Funding.Cat to the orginal dataset.
+Since we do not want to imputate on the categorical outcome variable,  this creation of a variable exists to create the proper dataframe within the imputated data
+sub$School.Funding.Cat <- ifelse(sub$School.Funding.Adequacy > 0, 1, 0)
+
+# Perform the imputation
+sub.imp <- mice(sub, m = 1, method = "pmm", print = FALSE, seed=3)
+```
+
 ### Exploratory Data Analysis & Data Preprocessing
 
-(Rishika will add here)
+```{r}
+plot1 <- ggplot(final_df, aes(x = Residential.Segregation.Index, y = Life.Expectancy)) +
+  geom_point(color='purple') +
+  labs(x = "Residential Segregation", y = "Life Expectancy")+ 
+  theme(plot.margin = margin(t = 20, r = 30, b = 20, l = 10),
+axis.title.x = element_text(size = 8),
+axis.title.y = element_text(size = 8))
+
+plot2 <- ggplot(final_df, aes(x = Percent.Limited.Access.Healthy.Foods
+, y = Life.Expectancy)) +
+  geom_point(color = "green") +
+  labs( x = "% Population With Limited Healthy Food Access") +
+  theme(plot.margin = margin(t = 20, r = 20, b = 20, l = 20),
+        axis.title.x = element_text(size = 8),
+        axis.title.y = element_text(size = 8))
+
+plot3 <- ggplot(final_df, aes(x = Percent.Uninsured.Adults, y = Life.Expectancy)) +
+  geom_point(color = "blue") +
+  labs(x = "% Adult Population Without Health Insurance", y = "Life Expectancy")+ 
+  theme(plot.margin = margin(t = 20, r = 20, b = 20, l = 20),
+        axis.title.x = element_text(size = 8),
+        axis.title.y = element_text(size = 8))
+
+plot1 + plot2 + plot3 + plot_layout(ncol = 2, nrow=2) + 
+  plot_annotation(title = "Relationships Between Predictors and 
+                  Life Expectancy Outcome",
+                  theme = theme(
+      plot.title = element_text(hjust=0.5)))
+```
 

### 2024-12-15 paper.qmd
adding conclusion para

@@
 ```
+
+\newpage
 
@@
 
+In Research Question 1, we used a multiple linear regression (MLR) model to assess how socioeconomic factors affect average life expectancy.
+The analysis shows that life expectancy is influenced by segregation, uninsured rates, household income, and regional differences, with the Northeast having the highest life expectancy, followed by the Midwest, West, and South.
+In Research Question 2, we used a logistic regression model to examine how residential segregation, property taxes, and median household income influence school funding adequacy.
+The results show that region has the greatest impact, with schools in the Northeast and Midwest being significantly more likely to receive adequate funding than those in the South.
+While residential segregation, property taxes, and household income have statistically significant effects, their influence is minimal compared to regional differences.
+
 In our analysis, we observed the significant impact of the regional location of a county on both its average life expectancy (a health indicator) and school funding adequacy (an educational indicator)--two key components of societal and individual well-being.

### 2024-12-15 final_code.qmd
fixing final code

@@
 plot1 <- ggplot(completed_data, aes(x = Residential.Segregation.Index, y = Life.Expectancy, color=Region_fac)) +
-  geom_point(color='black') + geom_smooth(method="lm",se=F) + 
+  geom_point(color='black', size=.5) + geom_smooth(method="lm",se=F) + 
   scale_color_brewer(palette = "Dark2") +
-  labs(x = "Residential Segregation Index", y = "Average Life Expectancy (Years)", color='U.S. Region')+ 
+  labs(x = "Residential Segregation Index", y = "Average Life Expectancy (Yrs)", color='U.S. Region')+ 
   theme(plot.margin = margin(t = 20, r = 30, b = 20, l = 10),
@@
 , y = Life.Expectancy)) +
-  geom_point(color='black') + geom_smooth(method="lm",se=F) + 
+  geom_point(color='black', size=0.5) + geom_smooth(method="lm",se=F) + 
   scale_color_brewer(palette = "Dark2") +
-  labs( x = "% Population With Limited Healthy Food Access", y="Average Life Expectancy (Years)") +
+  labs( x = "% Population With Limited Healthy Food Access", y="Average Life Expectancy (Yrs)") +
   theme(plot.margin = margin(t = 20, r = 20, b = 20, l = 20),
@@
 plot3 <- ggplot(completed_data, aes(x = X..Uninsured.Adults, y = Life.Expectancy)) +
-  geom_point(color='black') + geom_smooth(method="lm",se=F) + 
+  geom_point(color='black', size=0.5) + geom_smooth(method="lm",se=F) + 
   scale_color_brewer(palette = "Dark2") +
-  labs(x = "% Adult Population Without Health Insurance", y = "Average Life Expectancy (Years)")+ 
+  labs(x = "% Adult Population Without Health Insurance", y = "Average Life Expectancy (Yrs)")+ 
   theme(plot.margin = margin(t = 20, r = 20, b = 20, l = 20),
@@
       plot.title = element_text(hjust=0.5)))
+```
+
+```{r}
+ggplot(completed_data, aes(x=Region_fac, y = Life.Expectancy)) +
+  geom_boxplot() +
+  labs(x = "U.S. Region", y='Average Life Expectancy (Yrs)', color='U.S. Region', title='County Average Life Expectancy Distribution, Stratified by Region') +
+  #scale_color_hue(labels = c("Inadequate", "Adequate"))+ 
+  theme(plot.margin = margin(t = 20, r = 30, b = 20, l = 10),
+axis.title.x = element_text(size = 8),
+axis.title.y = element_text(size = 8),
+title=element_text(size=9))
 ```

### 2024-12-28 200_code/final_code.qmd
structural changes

--- 2024-12-15 final_code.qmd
@@
   geom_boxplot() +
-  labs(x = "U.S. Region", y='Average Life Expectancy (Yrs)', color='U.S. Region', title='County Average Life Expectancy Distribution, Stratified by Region') +
+  labs(x = "U.S. Region", y='Average Life Expectancy (Years)', color='U.S. Region', title='County Average Life Expectancy Distribution, Stratified by Region') +
   #scale_color_hue(labels = c("Inadequate", "Adequate"))+ 
   theme(plot.margin = margin(t = 20, r = 30, b = 20, l = 10),
-axis.title.x = element_text(size = 8),
-axis.title.y = element_text(size = 8),
+axis.title.x = element_text(size = 9),
+axis.title.y = element_text(size = 9),
 title=element_text(size=9))

### 2024-12-28 paper_files/paper.tex
structural changes

--- 2024-12-15 paper.tex
@@
 
-\usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{wrapfig}
 \usepackage{float}
-\usepackage{colortbl}
-\usepackage{pdflscape}
-\usepackage{tabu}
-\usepackage{threeparttable}
-\usepackage{threeparttablex}
-\usepackage[normalem]{ulem}
-\usepackage{makecell}
-\usepackage{xcolor}
 \usepackage{tabularray}
@@
 This study examines the relationships between socioeconomic, health, and
-education indicators in U.S. counties using data from the American
-Community Survey, the USDA Food Environment Atlas, Small Area Income and
-Poverty Estimates, and other sources.We analyze how factors such as
-residential segregation, health insurance coverage, access to healthy
-foods, and region influence life expectancy, and how residential
+educational indicators in U.S. counties using data from the County
+Health Rankings \& Roadmaps program and IPUMS. We analyze how factors
+such as residential segregation, health insurance coverage, access to
+healthy foods, and region impact life expectancy, and how residential
 segregation and region affect school funding adequacy. Our findings
@@
 education, and social equity that are essential for the development of
-individuals and communities. Life expectancy is an important health
-indicator that reflects not only the overall health of a society, but
-also its stability, economic resilience, and levels of inequality. While
-access to quality healthcare and nutritious foods are drivers of
-well-being, where you live and the nature of the community you live in
-likely also have an influence on health outcomes such as life
-expectancy. According to ``The Root Causes of Health Inequity -
-Communities in Action'' report from the National Institutes of Health,
-we can infer that differences in life expectancy between regions and
-socioeconomic groups often reveal systemic inequalities, with
-marginalized communities experiencing shorter life spans due to limited
-access to health care, healthy living environments, and economic
-opportunities.\textsuperscript{1}~When it comes to educational
-indicators, adequately funded schools ensure that students have equal
-access to learning opportunities, serving as a crucial investment in
-social mobility. The Stanford Center for Opportunity Policy in Education
-highlights that ``the unequal allocation and inadequate levels of
-resources in schools and communities {[}are{]} at the heart of many gaps
-in student opportunity.''\textsuperscript{2} This emphasizes the
-importance of equitable and adequate school funding in closing
-opportunity gaps and empowering students from disadvantaged backgrounds
-to achieve academic and economic success. By better understanding the
-correlates of school funding adequacy, we can take a major step towards
-implementing policy that targets the areas and communities most affected
-by this issue.
+individuals and communities. Life expectancy is an important indicator
+that reflects not only the overall health of a society, but also its
+stability, economic resilience, and levels of inequality. While access
+to quality healthcare and nutritious foods are drivers of well-being,
+where you live and the nature of the community you live in likely also
+have an influence on health outcomes such as life expectancy. According
+to ``The Root Causes of Health Inequity - Communities in Action'' report
+from the National Institutes of Health, we can infer that differences in
+life expectancy between regions and socioeconomic groups often reveal
+systemic inequalities, with marginalized communities experiencing
+shorter life spans due to limited access to health care, healthy living
+environments, and economic opportunities.\textsuperscript{1}~When it
+comes to educational indicators, adequately funded schools ensure that
+students have equal access to learning opportunities, serving as a
+crucial investment in social mobility. The Stanford Center for
+Opportunity Policy in Education highlights that ``the unequal allocation
+and inadequate levels of resources in schools and communities {[}are{]}
+at the heart of many gaps in student opportunity.''\textsuperscript{2}
+This emphasizes the importance of equitable and adequate school funding
+in closing opportunity gaps and empowering students from disadvantaged
+backgrounds to achieve academic and economic success. By better
+understanding the correlates of school funding adequacy, we can take a
+major step towards implementing policy that targets the areas and
+communities most affected by this issue.
 
@@
 
-Variables of interest were selected a priori based on their perceived
-relevance to the two outcomes we wanted to better understand: average
-life expectancy and school funding adequacy. While nutrition and access
-to quality healthcare are inherently linked to general wellbeing, it is
-worthwhile to assess how influential the environmental and financial
-aspects of these factors are on our tangible outcome. Therefore, we
-selected two key predictors for our life expectancy model: the
-percentage of the population with limited access to healthy foods and
-the percentage of uninsured adults in a county. The former is defined by
-the US Department of Agriculture as ``the percentage of the {[}county{]}
-population that is low-income and does not live close to a grocery
-store'', where ``close'' is defined differently for rural and nonrural
-areas.\textsuperscript{6} The latter variable represents the percentage
-of adults in a county younger than 65 who lack health insurance. We also
-were interested in seeing how residential segregation would impact life
-expectancy among these predictors, given existing studies on the
-negative health effects of segregation, and whether this relationship
-would exhibit any variations across different regions. This is why we
-chose to include residential segregation index values as a predictor.
-The residential segregation index, used by the County Health Rankings \&
-Roadmaps program, measures how evenly two groups, such as Black and
-white residents, are distributed across geographic areas, with values
-ranging from 0 (complete integration) to 100 (complete
+Variables of interest were selected \emph{a priori} based on their
+perceived relevance to the two outcomes we wanted to better understand:
+average life expectancy and school funding adequacy. While nutrition and
+access to quality healthcare are inherently linked to general wellbeing,
+it is worthwhile to assess how influential the environmental and
+financial aspects of these factors are on our tangible outcome.
+Therefore, we selected two key predictors for our life expectancy model:
+the percentage of the population with limited access to healthy foods
+and the percentage of uninsured adults in a county. The former is
+defined by the US Department of Agriculture as ``the percentage of the
+{[}county{]} population that is low-income and does not live close to a
+grocery store'', where ``close'' is defined differently for rural and
+nonrural areas.\textsuperscript{6} The latter variable represents the
+percentage of adults in a county younger than 65 who lack health
+insurance. We also were interested in seeing how residential segregation
+would impact life expectancy among these predictors, given existing
+studies on the negative health effects of segregation, and whether this
+relationship would exhibit any variations across different regions. This
+is why we chose to include residential segregation index values as a
+predictor. The residential segregation index, used by the County Health
+Rankings \& Roadmaps program, measures how evenly two groups, such as
+Black and white residents, are distributed across geographic areas, with
+values ranging from 0 (complete integration) to 100 (complete
 segregation).\textsuperscript{7}~
@@
 
-\subsubsection{\texorpdfstring{\emph{Data
-Overview}}{Data Overview}}\label{data-overview}
+\subsubsection{Data Overview}\label{data-overview}
 
@@
 expectancy. Specifically, for every \$10,000 increase in median
-household income, life expectancy increases by 1.25 years on average. In
-addition, the impact of region on life expectancy was substantial: life
-expectancy is on average 1.96 years higher in the West region compared
-to the South, 2.92 years higher in the Northeast compared to the South,
-and 1.92 years higher in the Midwest compared to the South, even when
-controlling for income, uninsured population, and racial diversity. This
-suggests, income as a factor withstanding, the differing quality of and
-access to healthcare and public services found in each region can have a
-significant impact on life expectancy.
+household income, life expectancy increases by 1.25 years on average,
+holding other variables constant. In addition, the impact of region on
+life expectancy was substantial: life expectancy is on average 1.96
+years higher in the West region compared to the South, 2.92 years higher
+in the Northeast compared to the South, and 1.92 years higher in the
+Midwest compared to the South, even when controlling for income,
+uninsured population, and racial diversity. This suggests, income as a
+factor withstanding, the differing quality of and access to healthcare
+and public services found in each region can have a significant impact
+on life expectancy.
 
@@
 increase in percentage of uninsured adults in a county, a county's
-average life expectancy increases by 0.032 years. While statistically
-significant, many studies covering the relationship between an
-individual's health and having insurance coverage have proven its
-positive correlation, suggesting our results may benefit from a closer
-look. In assessing our dataset, as seen in Appendix 1A, we find that our
-dataset contains more observations for counties with low uninsured
-individuals, or equivalently, counties where most people are covered. As
-the timeframe of our dataset is 2018-2022, it is likely that the
-aforementioned observations require greater scrutiny of the significance
-of the impact of this variable on average life expectancy of a county.~
+average life expectancy increases by 0.032 years. While this was
+statistically significant, many studies covering the relationship
+between an individual's health and having insurance coverage have proven
+its positive correlation, suggesting our results may benefit from a
+closer look. In assessing our dataset, as seen in Appendix A1, we find
+that our dataset contains more observations for counties with low
+uninsured individuals, or equivalently, counties where most people are
+covered. As the timeframe of our dataset is 2018-2022, it is likely that
+the aforementioned observations require greater scrutiny of the
+significance of the impact of this variable on average life expectancy
+of a county.~
 
@@
 that the effect of this environmental factor is minor or intertwined
-with economic variables, such as income, rather than being a direct or
-standalone influence. In terms of residential segregation, for every 1
-unit increase in a county's index value (corresponding to increased
+with economic variables, such as income, rather than being a standalone
+influence. In terms of residential segregation, for every 1 unit
+increase in a county's index value (corresponding to increased
 segregation), its average life expectancy increases by 0.0071 years in
@@
 residential segregation. This did not show a statistically significant
-difference in RSS (p=0.624), indicating that the interaction between
-region and residential segregation does not add meaningful explanatory
-power to our model, Consequently, we did not find sufficient evidence to
-support regional variations in the relationship between segregation and
-life expectancy. Instead, based on residential segregation's coefficient
-estimate in the model, life expectancy does appear to increase overall
-with an increase in residential segregation, however minor this increase
-may be. One possible theory for this observation is that counties with
-more racial uniformity might experience less racial conflict and more
-social cohesion, especially for minority communities.
+difference in RSS (\emph{p}=0.624), indicating that the interaction
+between region and residential segregation does not add meaningful
+explanatory power to our model, Consequently, we did not find sufficient
+evidence to support regional variations in the relationship between
+segregation and life expectancy. Instead, based on residential
+segregation's coefficient estimate in the model, life expectancy does
+appear to increase overall with an increase in residential segregation,
+however minor this increase may be. One possible theory for this
+observation is that counties with more racial uniformity might
+experience less racial conflict and more social cohesion, especially for
+minority communities, but this should be further investigated in future
+studies.
 
@@
 \includegraphics{paper_files/figure-pdf/unnamed-chunk-10-1.pdf}
+
+\newpage
 
@@
 substantial changes in the coefficient estimates for statistically
-significant predictors; however, the p-value for residential segregation
-increased from 0.031 to 0.054. Finally, with a VIF \textless2 for all
-predictors, there was little to no multicollinearity among the
-predictors.
+significant predictors; however, the \emph{p}-value for residential
+segregation increased from 0.031 to 0.054. Finally, with a VIF
+\textless2 for all predictors, there was little to no multicollinearity
+among the predictors.
 
@@
 regression model for school funding adequacy, with the South as the
-baseline category for the categorical region variable.~While all
+baseline category for the categorical region variable. While all
 estimates are statistically significant at the 0.05 significance level,
-exponentiated regression coefficient estimates from the logistic model
-demonstrate that the level of residential segregation, the median
-household income, and median property tax of the county have very
-minimal impacts on the odds of a school being adequately funded. This
-finding is interesting as looking at Figure 2, we see that across
-regions, the average racial segregation index is consistent across
-school funding adequacy categories.~
+exponentiated regression coefficients from the logistic model
+demonstrate that a county's median household income and median property
+tax have almost no impact on the odds of its schools being adequately
+funded. Our results demonstrate that a one unit increase to the
+residential segregation index increases the odds of school funding
+adequacy by 0.02 times, allowing us to conclude that residential
+segregation has a minimal effect on the odds of the school being
+adequately funded. Figure 2 demonstrates an interesting trend---each
+region shows a different relationship between the residential
+segregation distribution of adequately and inadequately funded
+schools,~pointing to a potential interaction between region and
+segregation. ~
 
-Region, however, has a substantial impact on the odds of school funding
-adequacy.~ Being located in the Northeast or the Midwest has the
-greatest impact on how well the school is funded, with the odds of
-adequate school funding being over 200 times higher in a Northeast
-county compared to a Southern county, and the odds of adequate school
-funding being over 3 times higher in a Northeast county compared to a
-Southern county, holding all other variables constant. Such disparities
-across the country are caused by a number of factors, including: (1)
-capacity - how well off a state is based on its economy and resources,
-and (2) effort - the states willingness to provide funding for
-education. Wealthier states with a high fiscal capacity, (typically
-those in the Northeast), have more funding available to spend on
-education than states with more limited resources (typically those in
-the South and the West).~
+Region has a substantial impact on the odds of school funding
+adequacy.~Being located in the Northeast or the Midwest has the greatest
+impact on how well a school is funded, with the odds of adequate school
+funding being over 200 times higher in a Northeast county compared to a
+Southern county, and the odds of adequate school funding being over 3
+times higher in a Midwest county compared to a Southern county, holding
+all other variables constant. Such disparities across the country are
+caused by a number of factors, including: (1) capacity - how well off a
+state is based on its economy and resources, and (2) effort - the
+state's willingness to provide funding for education. Wealthier states
+with a high fiscal capacity, (typically those in the Northeast), have
+more funding available to spend on education than states with more
+limited resources (typically those in the South and the West).~
 
@@
 assumption was largely met; interestingly, different linear trends were
-observed within each plot based on region, pointing to potential
+observed within each plot based on region, further supporting potential
 interaction effects between region and the three numerical predictors.
@@
 indicator) and school funding adequacy (an educational indicator)--two
-key components of societal and an individual's well-being. Life
-expectancy is highest in the Northeast, followed by the Midwest, West,
-and South, while school funding adequacy is significantly greater in the
-Northeast and Midwest compared to the South. Our findings highlight
-substantial differences in both outcomes between regions, which are
-often caused by different economic capabilities and different
-allocations of resources among countries. This study provides empirical
-evidence that underscores the importance for equitable social policies
-and targeted investments to address regional disparities and promote
-well-being across all communities.
+key components of societal and individual well-being. Life expectancy is
+highest in the Northeast, followed by the Midwest, West, and South,
+while school funding adequacy is significantly greater in the Northeast
+and Midwest compared to the South. Our findings highlight substantial
+differences in both outcomes between regions, which are often caused by
+different economic capabilities and different allocations of resources
+among countries. This study provides empirical evidence that underscores
+the importance for equitable social policies and targeted investments to
+address regional disparities and promote well-being across all
+communities.
 
@@
 
-Future studies could incorporate additional social variables to enhance
-the analysis and explore causal relationships between region and
-outcomes. For example, integrating a dataset on employment rates and
-school performance could offer deeper insights into how socioeconomic
-factors shape life expectancy and educational funding, complementing
-existing variables in our models. Another direction for future work is
-to delve deeper into understanding the surprising findings regarding the
-effects of uninsured adult percentage and residential segregation on
-life expectancy. Expanding the dataset to include a temporal scope, and
+Future studies could look at additional interaction effects with region
+and incorporate more social variables to enhance the analysis. For
+example, integrating a dataset on employment rates and school
+performance could offer deeper insights into how socioeconomic factors
+shape life expectancy and educational funding, complementing existing
+variables in our models. Another direction for future work is to delve
+deeper into understanding the surprising findings regarding the effects
+of uninsured adult percentage and residential segregation on life
+expectancy. Expanding the dataset to include a temporal scope, and
 finding more complete sources of data on variables of interest so that
@@
 \emph{Table A1: Summary Statistics For All Variables}
+
+\subsection{\texorpdfstring{\protect\includegraphics{images/clipboard-2912764107.png}}{}}\label{section}
 
